---
title: "Bacteria Genus Analyses"
output: html_document
date: "2024-04-23"
---

-################################################################################

# Libraries
```{r}
library(tidyverse)
library(vegan)
library(ggrepel)
library(ggordiplots)
```

-################################################################################

# kraken2 & bracken commands
```{bash}
# Kraken2 classification
kraken2 \
    --db ${KRAKEN2_DB} \
    --paired \
    --gzip-compressed \
    --use-names \
    --report ${OUTDIR}/${SAMPLE}.kreport \
    --output ${OUTDIR}/${SAMPLE}.kraken \
    ${READS_1} ${READS_2}

# Bracken abundance estimation
bracken \
    -d ${KRAKEN2_DB} \
    -i ${OUTDIR}/${SAMPLE}.kreport \
    -o ${OUTDIR}/${SAMPLE}.bracken \
    -r 1000 \
    -l G \
    
combine_bracken_outputs.py \
    --files *.bracken \
    -o combined.bracken \
    --names \
    ABBA_M11,ABBA_M12,ABBA_M13,ABBA_M14,ABBA_M15,\
    ACSA_M21,ACSA_M22,ACSA_M23,ACSA_M24,ACSA_M25,\
    COCO_M01,COCO_M02,COCO_M03,COCO_M04,COCO_M05,\
    FAGR_M16,FAGR_M17,FAGR_M18,FAGR_M19,FAGR_M20,\
    THOC_M06,THOC_M07,THOC_M08,THOC_M09,THOC_M10
```


-################################################################################

# Import Data

## Bracken data
```{r}
bracken.import <- read.delim("../data/bracken_outputs/combined.bracken.txt", header = TRUE, sep = "\t", dec = ".", row.names = 1)
```

## Reads data
```{r}
reads.data <- read.delim("../data/nb_reads.txt", header = TRUE, sep = "\t", dec = ",", row.names = 1)
```

## Metadata
```{r}
metadata <- read.delim("../data/samples_metadata.txt", header = TRUE, sep = "\t", dec = ".", row.names = 1)
metadata <- metadata %>% arrange(row.names(.))
metadata <- metadata %>% mutate(across(c(plant_clade,plant_order,plant_family,host_species,tree), as.factor))
reads.data$host_species <- metadata$host_species
metadata$host_species <- factor(x = metadata$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
reads.data$host_species <- factor(x = reads.data$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
metadata$shrub_tree <- c("tree","tree","tree","tree","tree","tree","tree","tree","tree","tree","shrub","shrub","shrub","shrub","shrub","tree","tree","tree","tree","tree","tree","tree","tree","tree","tree")
metadata$shrub_tree <- factor(x = metadata$shrub_tree, levels = c("shrub","tree"))
metadata$gymno_angio_tree <- c("gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree")
metadata$gymno_angio_tree <- factor(x = metadata$gymno_angio_tree, levels = c("gymnosperm_tree","angiosperm_tree","angiosperm_shrub"))
metadata.bac.comm <- metadata
```


-################################################################################

# Taxonomy

## Export taxonomy ID number for NCBI
```{r, eval = FALSE}
#write_file(paste(bracken.import$taxonomy_id, collapse = "\n"), file = "../ncbi_taxonomy_id.txt")
# load file to: https://www.ncbi.nlm.nih.gov/Taxonomy/TaxIdentifier/tax_identifier.cgi
```

## Import NCBI taxonomy (report from online NCBI web page)
```{r}
ncbi.tax <- read_delim("../data/ncbi_tax_report.txt")
ncbi.tax <- ncbi.tax[,-c(1,2,4,6)]
ncbi.tax <- ncbi.tax %>%
  mutate(domain = case_when(
    sapply(strsplit(lineage, " "), tail, 1) == "10239" ~ "Virus",
    sapply(strsplit(lineage, " "), function(x) tail(x, 2)[1]) == "2" ~ "Bacteria",
    sapply(strsplit(lineage, " "), function(x) tail(x, 2)[1]) == "2157" ~ "Archaea",
    sapply(strsplit(lineage, " "), function(x) tail(x, 2)[1]) == "2759" ~ "Eukaryota",
    TRUE ~ NA_character_  # Keep as NA if no conditions are met
  ))
colnames(ncbi.tax) <- c("taxid","Genus","NCBI_lineage","Domain")
bracken.import <- cbind(bracken.import, ncbi.tax)
bracken.import[c(which(bracken.import$taxonomy_id != bracken.import$taxid)),]
```

-################################################################################

# Edit bracken reads dataframe

Build bacterial reads matrix
```{r}
bacteria.reads <- bracken.import[bracken.import$Domain == "Bacteria", ]
bacteria.reads <- bacteria.reads[,-c(1,2,53:56)]
bacteria.reads <- bacteria.reads[,-c(2,4,6,8,10,12,14,16,18,20,22,24,26,28,30,32,34,36,38,40,42,44,46,48,50)] # remove fractions
colnames(bacteria.reads) <- gsub("_num", "", colnames(bacteria.reads))
bacteria.reads <- t(bacteria.reads)
bacteria.reads <- data.frame(bacteria.reads)
n.reads.bacteria <- sum(rowSums(bacteria.reads)) # Total bacterial reads : 549,194,184
# Combine Methylorubrum with Methylobacterium
bacteria.reads$Methylobacterium <- bacteria.reads$Methylobacterium+bacteria.reads$Methylorubrum
bacteria.reads <- bacteria.reads[, names(bacteria.reads) != "Methylorubrum"]
bracken.reads <- bacteria.reads
```

-################################################################################

# Correlation between bracken and pangenome approach for Methylobacterium reads

```{r}
lm <- lm(2*bracken.reads[,"Methylobacterium"] ~ reads.data$mapped_reads)
plot(2*bracken.reads[,"Methylobacterium"] ~ reads.data$mapped_reads) + abline(lm[["coefficients"]][["(Intercept)"]],lm[["coefficients"]][["reads.data$mapped_reads"]])
summary(lm)
plot(lm)
hist(residuals(lm), breaks = 25)
```

-################################################################################

# Methylobacterium relative abundance in bacterial communities among hosts

For these analyses, keep all assigned bacterial genera (n = 2065), even though some of them may not be accurately assigned, because we can assume these reads belong to bacteria.

## Exploration
```{r}
# number of reads per sample (in millions)
rowSums(bracken.reads)/1000000
min(rowSums(bracken.reads))/1000000
# test correlation between assigned reads (bracken) and filtered reads
plot(rowSums(bracken.reads) ~ reads.data$filtered_pairs)
summary(lm(rowSums(bracken.reads) ~ reads.data$filtered_pairs))
```

## Rarefying : One random subsampling ("rarefying" sensu Schloss 2024)
```{r, eval=FALSE}
rrarefy.comm <- bracken.reads
length(rrarefy.comm[rrarefy.comm == 10])
rrarefy.comm[rrarefy.comm == 10] <- 1 # to make function work
rrarefy.comm <- rrarefy(rrarefy.comm, sample = min(rowSums(rrarefy.comm)))
rrarefy.comm <- data.frame(rrarefy.comm)
rowSums(bracken.reads)
rowSums(rrarefy.comm)
colnames(rrarefy.comm)[colSums(rrarefy.comm) == 0] # no lost of genera
```

Save rrarefy.comm dataframe for further code rerun. Load RDS.
```{r}
#saveRDS(rrarefy.comm, "../rds/rrarefy.comm.rds")
rrarefy.comm <- readRDS("../rds/rrarefy.comm.rds")
```


## Methylobacterium relative abundances in bacterial communities
```{r}
bracken.reads[,"Methylobacterium"]
rrarefy.comm[,"Methylobacterium"]
plot(rrarefy.comm[,"Methylobacterium"] ~ bracken.reads[,"Methylobacterium"])
metadata.bac.comm$meth_rel_abund <- rrarefy.comm[,"Methylobacterium"]/rowSums(rrarefy.comm)*100 # With rarefying
```

## Other abundant genera relative abundances in bacterial communities
```{r}
metadata.bac.comm$mycobacterium_rel_abund <- rrarefy.comm[,"Mycobacterium"]/rowSums(rrarefy.comm)*100
metadata.bac.comm$xanthomonas_rel_abund <- rrarefy.comm[,"Xanthomonas"]/rowSums(rrarefy.comm)*100
metadata.bac.comm$streptomyces_rel_abund <- rrarefy.comm[,"Streptomyces"]/rowSums(rrarefy.comm)*100
mean(metadata.bac.comm$mycobacterium_rel_abund)
sd(metadata.bac.comm$mycobacterium_rel_abund)
mean(metadata.bac.comm$xanthomonas_rel_abund)
sd(metadata.bac.comm$xanthomonas_rel_abund)
mean(metadata.bac.comm$streptomyces_rel_abund)
sd(metadata.bac.comm$streptomyces_rel_abund)
mean(metadata.bac.comm$meth_rel_abund)
sd(metadata.bac.comm$meth_rel_abund)
```


## Meth abundance ~ host

### Model
```{r}
plot(meth_rel_abund ~ host_species, data = metadata.bac.comm)
lm <- lm(meth_rel_abund ~ host_species, data = metadata.bac.comm)
plot(lm)
summary(lm)
hist(residuals(lm), breaks = 50)
shapiro.test(residuals(lm))
```

### Post-hoc
```{r}
# Shrub vs. trees

## Student t-test assumptions : normal distribution of data and equal variance
hist(metadata.bac.comm[metadata.bac.comm$shrub_tree == "tree",]$meth_rel_abund, breaks = 25)
shapiro.test(metadata.bac.comm[metadata.bac.comm$shrub_tree == "tree",]$meth_rel_abund)
hist(metadata.bac.comm[metadata.bac.comm$shrub_tree == "shrub",]$meth_rel_abund, breaks = 25)
shapiro.test(metadata.bac.comm[metadata.bac.comm$shrub_tree == "shrub",]$meth_rel_abund)
var(metadata.bac.comm[metadata.bac.comm$shrub_tree == "tree",]$meth_rel_abund)
sd(metadata.bac.comm[metadata.bac.comm$shrub_tree == "tree",]$meth_rel_abund)
var(metadata.bac.comm[metadata.bac.comm$shrub_tree == "shrub",]$meth_rel_abund)
sd(metadata.bac.comm[metadata.bac.comm$shrub_tree == "shrub",]$meth_rel_abund)
## unequal variance so welch t-test
t1 <- t.test(meth_rel_abund ~ shrub_tree, data = metadata.bac.comm, var.equal = F)
t1

# Gymno vs. angio trees

## Student t-test assumptions : normal distribution of data and equal variance
hist(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "gymnosperm_tree",]$meth_rel_abund, breaks = 25)
shapiro.test(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "gymnosperm_tree",]$meth_rel_abund)
hist(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "angiosperm_tree",]$meth_rel_abund, breaks = 25)
shapiro.test(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "angiosperm_tree",]$meth_rel_abund)
var(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "gymnosperm_tree",]$meth_rel_abund)
var(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "angiosperm_tree",]$meth_rel_abund)
## equal variance so student t-test
t2 <- t.test(meth_rel_abund ~ gymno_angio_tree, data = metadata.bac.comm[c(1:10,16:25),], var.equal = T)
t2

# p-values adjustement
p.adjust(c(t1[["p.value"]],t2[["p.value"]]), method = "bonferroni")
```

### Plot
```{r}
ggplot(metadata.bac.comm, aes(x = host_species, y = meth_rel_abund)) +
geom_boxplot(alpha = 1, aes(fill = host_species), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = host_species), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
scale_fill_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"))+
scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Host Species", 
     y = expression(bolditalic("Methylobacterium") ~ bold("relative abundance (%)")))+
scale_x_discrete(labels = c("ABBA" = "A. balsamea", "THOC" = "T. occidentalis","ACSA" = "A. saccharum", "FAGR" = "F. grandifolia", "COCO" = "C. cornuta"))+
ylim(0,20) +
theme_bw()+
theme(text = element_text(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0, face = "italic"),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/figure_1_Methylo_relative_abundance.pdf", width = 6, height = 4)
```

-################################################################################

# Community genera composition

## "Others category" = Genera that did not make up a total of 0.5% of bacterial reads across samples.
```{r}
rrarefy.comm.w_oth <- rrarefy.comm
Others.names <- names(colSums(rrarefy.comm.w_oth)[colSums(rrarefy.comm.w_oth) < (0.005 * sum(rrarefy.comm.w_oth))])
Others <- rowSums(rrarefy.comm.w_oth[, Others.names])
rrarefy.comm.w_oth <- rrarefy.comm.w_oth[, !(names(rrarefy.comm.w_oth) %in% Others.names)]
rrarefy.comm.w_oth$Others <- Others
rowSums(rrarefy.comm)
rowSums(rrarefy.comm.w_oth)
```

## List of genera names in taxonomic order
```{r}
abund_genera_taxo <- as.data.frame(read_delim("../data/bacterial_genera_taxonomy_from_NCBI.csv"))
abund_genera_taxo$Genus <- factor(abund_genera_taxo$Genus, levels = unique(abund_genera_taxo$Genus))
```


## Barplot
```{r}
abundances <- picante::matrix2sample(rrarefy.comm.w_oth)
colnames(abundances) <- c("sample","abund","genus")
abundances$sample <- factor(x = abundances$sample, levels =c ("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M23","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M18","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
abundances$genus <- droplevels(abundances$genus)
abundances$genus <- factor(x = abundances$genus, levels = abund_genera_taxo$Genus)

palette8 <- readRDS("../rds/palette8.rds")

# Composition with abundances
ggplot(abundances, aes(x = sample, y = abund, fill = genus)) +
  geom_bar(stat = "identity", position = "fill"
  ) +
  labs(x = "Sample", y = "Relative abundance") +
  scale_fill_manual(values=c(palette8, "grey80"), name = "Bacterial genera") +
  guides(fill = guide_legend(ncol = 1)) +
  theme_bw() +
  theme(text = element_text(),
      legend.justification = c(0.5, 1),
      legend.text = element_text(size = 10.5),
      legend.key.size = unit(0.43, "cm"),
      legend.position = "right", 
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 11),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/figure_S6_bacterial_genera_barplot.pdf", width = 8, height = 5.5)
```

What are the most abundant genera across samples?
```{r}
colSums(rrarefy.comm.w_oth)[order(colSums(rrarefy.comm.w_oth), decreasing = TRUE)]/sum(rrarefy.comm.w_oth)*100 
```

-################################################################################

# Beta diversity

## Remove 'noise'

Before performing diversity analyses, we will remove all presence (replace with a count of 1) of a genus that is not >= 0.1% of total bacterial reads in a given sample. The rationale is that kraken is known to be not conservative enough. So we assume these 'low abundant' genera detection are actually noise.

```{r}
bracken.reads.clean <- bracken.reads
bracken.reads.clean[bracken.reads.clean < (0.001 * rowSums(bracken.reads.clean))] <- 1 # If a genus accounts for less than 0.1% of total bacterial reads in a given sample, its read count is replaced by 1 (so the rarefaction function can work)
bracken.reads.clean <- bracken.reads.clean[, colSums(bracken.reads.clean) > 25] # remove genera which are not present >= 0.1% in at least one sample.
rowSums(bracken.reads.clean)/rowSums(bracken.reads)*100
mean(rowSums(bracken.reads.clean)/rowSums(bracken.reads)*100)
sum(bracken.reads.clean)/sum(bracken.reads)*100
length(colnames(bracken.reads.clean))
length(colnames(bracken.reads.clean))/length(colnames(bracken.reads))*100
# By getting rid of the noise, we keep approx. 82.07% of the reads, with 198 (9.59%) genera out of 2065 originally 'detected'
```


## Rarefaction
```{r, eval=FALSE}
bracken.dist <- avgdist(bracken.reads.clean, sample = min(rowSums(bracken.reads.clean)), dmethod = "bray", iteration = 100)
#saveRDS(bracken.dist, "../rds/bracken_dist_iterations_100.rds")
```

```{r}
bracken.dist <- readRDS("../rds/bracken_dist_iterations_100.rds")
```


## PERMANOVA
```{r}
bacteria.perm <- adonis2(bracken.dist ~ host_species, data = metadata)
bacteria.perm
```

## BETADISPER
```{r}
bacteria.disper <- betadisper(bracken.dist, metadata$host_species)
anova(bacteria.disper)
plot(bacteria.disper)
```
## Pairwise PERMANOVAs and BETADISPER
```{r}
shrub_tree_PERM <- adonis2(bracken.dist ~ shrub_tree, data = metadata.bac.comm)
shrub_tree_PERM
shrub_tree_DISP <- betadisper(bracken.dist, metadata.bac.comm$shrub_tree)
anova(shrub_tree_DISP)
plot(shrub_tree_DISP)

bracken.dist.noCOCO <- data.frame(as.matrix(bracken.dist))
bracken.dist.noCOCO <- bracken.dist.noCOCO[c(1:10,16:25),c(1:10,16:25)]
bracken.dist.noCOCO <- as.dist(as.matrix(bracken.dist.noCOCO), diag = FALSE, upper = FALSE)
gymno_angio_tree_PERM <- adonis2(bracken.dist.noCOCO ~ gymno_angio_tree, data = metadata.bac.comm[c(1:10,16:25),])
gymno_angio_tree_PERM
gymno_angio_tree_DISP <- betadisper(bracken.dist.noCOCO, metadata.bac.comm[c(1:10,16:25),]$gymno_angio_tree)
anova(gymno_angio_tree_DISP)
plot(gymno_angio_tree_DISP, label = F)
```

## PCoA
```{r}
bacteria.pcoa <- cmdscale(bracken.dist, k =(nrow(bracken.reads.clean) - 1), eig = TRUE)
pcoa.eig <- eigenvals(bacteria.pcoa) 
pcoa.eig <- abs(pcoa.eig) 
sum.pcoa.eig <- sum(pcoa.eig) 
pcoa.prop.exp <- pcoa.eig/sum.pcoa.eig 
(pcoa.axis1 <- pcoa.prop.exp[1])
(pcoa.axis2 <- pcoa.prop.exp[2])
site.scores.pcoa <- scores(bacteria.pcoa)
sp.scores.pcoa <- wascores(bacteria.pcoa$points, bracken.reads.clean)
site.scores.pcoa <- data.frame(site.scores.pcoa)
sp.scores.pcoa <- data.frame(sp.scores.pcoa)
host_species <- metadata.bac.comm$host_species
site.scores.pcoa <- cbind(site.scores.pcoa, host_species)
pcoa.ggordiplot <- gg_ordiplot(bacteria.pcoa, choices = c(1,2), groups = metadata.bac.comm$host_species, kind = "sd", conf = 0.95, pt.size = 3)
pcoa.ellipses <- pcoa.ggordiplot$df_ellipse # extract confidence ellipse
```

Plot
```{r}
ggplot(data = site.scores.pcoa, 
         aes(x = Dim1, y = Dim2), 
         color = host_species) + 
  expand_limits(x=c(-0.4,0.4), y=c(-0.2,0.3)) +
  geom_point(aes(shape = host_species, color = host_species), show.legend = T, 
             size=3) +
  xlab(paste("PCoA1 (",round(pcoa.axis1*100, digits = 2),"%)", sep = "")) +
  ylab(paste("PCoA2 (",round(pcoa.axis2*100, digits = 2),"%)", sep = "")) +
  scale_shape_manual(values = c(16,16,16,16,16)) +
  guides(shape = "none") +
  geom_path(data = pcoa.ellipses, 
            aes(x = x, y = y, color = Group), 
            show.legend = FALSE) + 
  scale_color_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"), name = "Host species", labels = c("ABBA" = expression(italic("A. balsamea")), "THOC" = expression(italic("T. occidentalis")),"ACSA" = expression(italic("A. saccharum")), "FAGR" = expression(italic("F. grandifolia")), "COCO" = expression(italic("C. cornuta")))) +
  theme_bw() +
  theme(text = element_text(),
      legend.justification = c(0.5, 0),
      legend.text = element_text(size = 11),
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/figure_S7_bacteria_genera_pcoa.pdf", width = 6, height = 4)
```


-################################################################################