---
title: "bracken_analyses"
output: html_document
date: "2024-04-23"
---

-################################################################################

## Load libraries
```{r}
library(tidyverse)
library(vegan)
#library(pairwiseAdonis)
library(ggrepel)
library(ggordiplots)
#library(extrafont)
#library(phylostratr)
```

-################################################################################

# Import Data

## Bracken data
```{r}
bracken.import <- read.delim("../data/bracken_outputs/combined.bracken.txt", header = TRUE, sep = "\t", dec = ".", row.names = 1)
```

## Reads data
```{r}
reads.data <- read.delim("../data/nb_reads.txt", header = TRUE, sep = "\t", dec = ",", row.names = 1)
```

## Metadata
```{r}
metadata <- read.delim("../data/samples_metadata.txt", header = TRUE, sep = "\t", dec = ".", row.names = 1)
metadata <- metadata %>% arrange(row.names(.))
metadata <- metadata %>% mutate(across(c(plant_clade,plant_order,plant_family,host_species,tree), as.factor))
reads.data$host_species <- metadata$host_species
metadata$host_species <- factor(x = metadata$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
reads.data$host_species <- factor(x = reads.data$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
metadata$shrub_tree <- c("tree","tree","tree","tree","tree","tree","tree","tree","tree","tree","shrub","shrub","shrub","shrub","shrub","tree","tree","tree","tree","tree","tree","tree","tree","tree","tree")
metadata$shrub_tree <- factor(x = metadata$shrub_tree, levels = c("shrub","tree"))
metadata$gymno_angio_tree <- c("gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree")
metadata$gymno_angio_tree <- factor(x = metadata$gymno_angio_tree, levels = c("gymnosperm_tree","angiosperm_tree","angiosperm_shrub"))
metadata.bac.comm <- metadata
```


-################################################################################

# Taxonomy

## Export taxonomy ID number for NCBI
```{r, eval = FALSE}
#write_file(paste(bracken.import$taxonomy_id, collapse = "\n"), file = "../ncbi_taxonomy_id.txt")
# load file to: https://www.ncbi.nlm.nih.gov/Taxonomy/TaxIdentifier/tax_identifier.cgi
```

## Import NCBI taxonomy (report from online NCBI web page)
```{r}
ncbi.tax <- read_delim("../data/ncbi_tax_report.txt")
ncbi.tax <- ncbi.tax[,-c(1,2,4,6)]
ncbi.tax <- ncbi.tax %>%
  mutate(domain = case_when(
    sapply(strsplit(lineage, " "), tail, 1) == "10239" ~ "Virus",
    sapply(strsplit(lineage, " "), function(x) tail(x, 2)[1]) == "2" ~ "Bacteria",
    sapply(strsplit(lineage, " "), function(x) tail(x, 2)[1]) == "2157" ~ "Archaea",
    sapply(strsplit(lineage, " "), function(x) tail(x, 2)[1]) == "2759" ~ "Eukaryota",
    TRUE ~ NA_character_  # Keep as NA if no conditions are met
  ))
colnames(ncbi.tax) <- c("taxid","Genus","NCBI_lineage","Domain")
bracken.import <- cbind(bracken.import, ncbi.tax)
bracken.import[c(which(bracken.import$taxonomy_id != bracken.import$taxid)),]
```

-################################################################################

# Edit bracken reads dataframe

Build bacterial reads matrix
```{r}
bacteria.reads <- bracken.import[bracken.import$Domain == "Bacteria", ]
bacteria.reads <- bacteria.reads[,-c(1,2,53:56)]
bacteria.reads <- bacteria.reads[,-c(2,4,6,8,10,12,14,16,18,20,22,24,26,28,30,32,34,36,38,40,42,44,46,48,50)] # remove fractions
colnames(bacteria.reads) <- gsub("_num", "", colnames(bacteria.reads))
bacteria.reads <- t(bacteria.reads)
bacteria.reads <- data.frame(bacteria.reads)
n.reads.bacteria <- sum(rowSums(bacteria.reads)) # Total bacterial reads : 549,194,184
# Combine Methylorubrum with Methylobacterium
bacteria.reads$Methylobacterium <- bacteria.reads$Methylobacterium+bacteria.reads$Methylorubrum
bacteria.reads <- bacteria.reads[, names(bacteria.reads) != "Methylorubrum"]
bracken.reads <- bacteria.reads
```

-################################################################################

# Correlation between bracken and pangenome approach for Methylobacterium reads

```{r}
lm <- lm(2*bracken.reads[,"Methylobacterium"] ~ reads.data$mapped_reads)
plot(2*bracken.reads[,"Methylobacterium"] ~ reads.data$mapped_reads) + abline(lm[["coefficients"]][["(Intercept)"]],lm[["coefficients"]][["reads.data$mapped_reads"]])
summary(lm)
plot(lm)
hist(residuals(lm), breaks = 25)
```

-################################################################################

# Methylobacterium relative abundance in bacterial communities among hosts

For these analyses, keep all assigned bacterial genera (n = 2065), even though some of them may not be accurately assigned, because we can assume these reads belong to bacteria.

## Exploration
```{r}
# number of reads per sample (in millions)
rowSums(bracken.reads)/1000000
min(rowSums(bracken.reads))/1000000
# test correlation between assigned reads (bracken) and filtered reads
plot(rowSums(bracken.reads) ~ reads.data$filtered_pairs)
summary(lm(rowSums(bracken.reads) ~ reads.data$filtered_pairs))
```

## Rarefying : One random subsampling ("rarefying" sensu Schloss 2024)
```{r, eval=FALSE}
rrarefy.comm <- bracken.reads
length(rrarefy.comm[rrarefy.comm == 10])
rrarefy.comm[rrarefy.comm == 10] <- 1 # to make function work
rrarefy.comm <- rrarefy(rrarefy.comm, sample = min(rowSums(rrarefy.comm)))
rrarefy.comm <- data.frame(rrarefy.comm)
rowSums(bracken.reads)
rowSums(rrarefy.comm)
colnames(rrarefy.comm)[colSums(rrarefy.comm) == 0] # no lost of genera
```

Save rrarefy.comm dataframe for further code rerun. Load RDS.
```{r}
#saveRDS(rrarefy.comm, "../rds/rrarefy.comm.rds")
rrarefy.comm <- readRDS("../rds/rrarefy.comm.rds")
```


## Methylobacterium relative abundances in bacterial communities
```{r}
bracken.reads[,"Methylobacterium"]
rrarefy.comm[,"Methylobacterium"]
plot(rrarefy.comm[,"Methylobacterium"] ~ bracken.reads[,"Methylobacterium"])
metadata.bac.comm$meth_rel_abund <- rrarefy.comm[,"Methylobacterium"]/rowSums(rrarefy.comm)*100 # With rarefying
```



## Other abundant genera relative abundances in bacterial communities
```{r}
metadata.bac.comm$mycobacterium_rel_abund <- rrarefy.comm[,"Mycobacterium"]/rowSums(rrarefy.comm)*100
metadata.bac.comm$xanthomonas_rel_abund <- rrarefy.comm[,"Xanthomonas"]/rowSums(rrarefy.comm)*100
metadata.bac.comm$streptomyces_rel_abund <- rrarefy.comm[,"Streptomyces"]/rowSums(rrarefy.comm)*100
mean(metadata.bac.comm$mycobacterium_rel_abund)
sd(metadata.bac.comm$mycobacterium_rel_abund)
mean(metadata.bac.comm$xanthomonas_rel_abund)
sd(metadata.bac.comm$xanthomonas_rel_abund)
mean(metadata.bac.comm$streptomyces_rel_abund)
sd(metadata.bac.comm$streptomyces_rel_abund)
mean(metadata.bac.comm$meth_rel_abund)
sd(metadata.bac.comm$meth_rel_abund)
```


## Meth abundance ~ host

### Model
```{r}
plot(meth_rel_abund ~ host_species, data = metadata.bac.comm)
lm <- lm(meth_rel_abund ~ host_species, data = metadata.bac.comm)
plot(lm)
summary(lm)
hist(residuals(lm), breaks = 50)
shapiro.test(residuals(lm))
```

### Post-hoc
```{r}
# Shrub vs. trees

## Student t-test assumptions : normal distribution of data and equal variance
hist(metadata.bac.comm[metadata.bac.comm$shrub_tree == "tree",]$meth_rel_abund, breaks = 25)
shapiro.test(metadata.bac.comm[metadata.bac.comm$shrub_tree == "tree",]$meth_rel_abund)
hist(metadata.bac.comm[metadata.bac.comm$shrub_tree == "shrub",]$meth_rel_abund, breaks = 25)
shapiro.test(metadata.bac.comm[metadata.bac.comm$shrub_tree == "shrub",]$meth_rel_abund)
var(metadata.bac.comm[metadata.bac.comm$shrub_tree == "tree",]$meth_rel_abund)
sd(metadata.bac.comm[metadata.bac.comm$shrub_tree == "tree",]$meth_rel_abund)
var(metadata.bac.comm[metadata.bac.comm$shrub_tree == "shrub",]$meth_rel_abund)
sd(metadata.bac.comm[metadata.bac.comm$shrub_tree == "shrub",]$meth_rel_abund)
## unequal variance so welch t-test
t1 <- t.test(meth_rel_abund ~ shrub_tree, data = metadata.bac.comm, var.equal = F)
t1

# Gymno vs. angio trees

## Student t-test assumptions : normal distribution of data and equal variance
hist(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "gymnosperm_tree",]$meth_rel_abund, breaks = 25)
shapiro.test(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "gymnosperm_tree",]$meth_rel_abund)
hist(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "angiosperm_tree",]$meth_rel_abund, breaks = 25)
shapiro.test(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "angiosperm_tree",]$meth_rel_abund)
var(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "gymnosperm_tree",]$meth_rel_abund)
var(metadata.bac.comm[metadata.bac.comm$gymno_angio_tree == "angiosperm_tree",]$meth_rel_abund)
## equal variance so student t-test
t2 <- t.test(meth_rel_abund ~ gymno_angio_tree, data = metadata.bac.comm[c(1:10,16:25),], var.equal = T)
t2

# p-values adjustement
p.adjust(c(t1[["p.value"]],t2[["p.value"]]), method = "bonferroni")
```

### Plot
```{r}
ggplot(metadata.bac.comm, aes(x = host_species, y = meth_rel_abund)) +
geom_boxplot(alpha = 1, aes(fill = host_species), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = host_species), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
scale_fill_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"))+
scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Host Species", 
     y = expression(bolditalic("Methylobacterium") ~ bold("relative abundance (%)")))+
scale_x_discrete(labels = c("ABBA" = "A. balsamea", "THOC" = "T. occidentalis","ACSA" = "A. saccharum", "FAGR" = "F. grandifolia", "COCO" = "C. cornuta"))+
ylim(0,20) +
theme_bw()+
theme(text = element_text(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0, face = "italic"),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/figure_1_Methylo_relative_abundance.pdf", width = 6, height = 4)
```

-################################################################################


# Community genera composition

## "Others category" = Genera that did not make up a total of 0.5% of bacterial reads across samples.
```{r}
rrarefy.comm.w_oth <- rrarefy.comm
Others.names <- names(colSums(rrarefy.comm.w_oth)[colSums(rrarefy.comm.w_oth) < (0.005 * sum(rrarefy.comm.w_oth))])
Others <- rowSums(rrarefy.comm.w_oth[, Others.names])
rrarefy.comm.w_oth <- rrarefy.comm.w_oth[, !(names(rrarefy.comm.w_oth) %in% Others.names)]
rrarefy.comm.w_oth$Others <- Others
rowSums(rrarefy.comm)
rowSums(rrarefy.comm.w_oth)
```

## List of genera names in taxonomic order
```{r}
abund_genera_taxo <- as.data.frame(read_delim("../data/bacterial_genera_taxonomy_from_NCBI.csv"))
abund_genera_taxo$Genus <- factor(abund_genera_taxo$Genus, levels = unique(abund_genera_taxo$Genus))
```


## Barplot
```{r}
abundances <- picante::matrix2sample(rrarefy.comm.w_oth)
colnames(abundances) <- c("sample","abund","genus")
abundances$sample <- factor(x = abundances$sample, levels =c ("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M23","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M18","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
abundances$genus <- droplevels(abundances$genus)
abundances$genus <- factor(x = abundances$genus, levels = abund_genera_taxo$Genus)

palette8 <- readRDS("../rds/palette8.rds")

# Composition with abundances
ggplot(abundances, aes(x = sample, y = abund, fill = genus)) +
  geom_bar(stat = "identity", position = "fill"
  ) +
  labs(x = "Sample", y = "Relative abundance") +
  scale_fill_manual(values=c(palette8, "grey80"), name = "Bacterial genera") +
  guides(fill = guide_legend(ncol = 1)) +
  theme_bw() +
  theme(text = element_text(),
      legend.justification = c(0.5, 1),
      legend.text = element_text(size = 10.5),
      legend.key.size = unit(0.43, "cm"),
      legend.position = "right", 
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 11),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/figure_S6_bacterial_genera_barplot.pdf", width = 8, height = 5.5)
```

What are the most abundant genera across samples?
```{r}
colSums(rrarefy.comm.w_oth)[order(colSums(rrarefy.comm.w_oth), decreasing = TRUE)]/sum(rrarefy.comm.w_oth)*100 
```



RENDU LA!!!! 

### Co-occurence networks

TAKE 1
```{r}
rrarefy.comm.only.abund.genera <- rrarefy.comm.w_oth[,colnames(rrarefy.comm.w_oth)!="Others"]
str(rrarefy.comm.only.abund.genera)

rrarefy.comm.only.abund.genera <- as.matrix(rrarefy.comm.only.abund.genera)

co_occurrence_matrix <- t(rrarefy.comm.only.abund.genera) %*% rrarefy.comm.only.abund.genera
co_occurrence_matrix

network <- graph.adjacency(co_occurrence_matrix, mode = "undirected", weighted = TRUE, diag = FALSE)
summary(network)

plot(network,
     vertex.size = 10, 
     vertex.label.cex = 0.5,
     edge.width = E(network)$weight / 2,  # Scale edge width based on weight
     main = "Occurrence Network")

```

TAKE 2
```{r}
rrarefy.comm.only.abund.genera <- rrarefy.comm.w_oth[,colnames(rrarefy.comm.w_oth)!="Others"]

cor_matrix <- cor(rrarefy.comm.only.abund.genera, method = "spearman")

library(reshape2)
cor_melted <- melt(cor_matrix)
ggplot(cor_melted, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, name = "Correlation") +
  theme_minimal() +
  labs(title = "Co-Abundance Heatmap", x = "Taxa", y = "Taxa") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Load igraph for network visualization
library(igraph)

# Create an igraph object from the correlation matrix (threshold to avoid too many edges)
threshold <- 0.75  # You can adjust this threshold
cor_matrix[abs(cor_matrix) < threshold] <- 0  # Remove weak correlations

# Create a graph from the correlation matrix
network <- graph.adjacency(cor_matrix, mode = "undirected", weighted = TRUE, diag = FALSE)

cor_matrix_for_layout <- abs(cor_matrix)
network_for_layout <- graph.adjacency(cor_matrix_for_layout, mode = "undirected", weighted = TRUE, diag = FALSE)


plot(network_for_layout, 
     vertex.size = 10, 
     vertex.label.cex = 0.8,
     edge.width = E(network)$weight * 5,  # Scale edge width by correlation strength
     edge.color = ifelse(E(network)$weight > 0, "red", "blue"),  # Red for positive, blue for negative
     layout = layout_with_fr(network_for_layout),  # Force-directed layout
     main = "Co-Abundance Network (Positive and Negative Correlations)")
```





################################################################################

# Diversity analyses

## Remove 'noise'

Before performing diversity analyses, we will remove all presence (replace with a count of 1) of a genus that is not >= 0.1% of total bacterial reads in a given sample. The rationale is that kraken is known to be not conservative enough. So we assume these 'low abundant' genera detection are actually noise.

```{r}
bracken.reads.clean <- bracken.reads
bracken.reads.clean[bracken.reads.clean < (0.001 * rowSums(bracken.reads.clean))] <- 1 # If a genus accounts for less than 0.1% of total bacterial reads in a given sample, its read count is replaced by 1 (so the rarefaction functions can work)
bracken.reads.clean <- bracken.reads.clean[, colSums(bracken.reads.clean) > 25] # remove genera which are not present >= 0.1% in at least one sample.

## IMPORTANT NOTE : whatever the noise threshold I use (0.01, 0.001, 0.0001), my PERMANOVA and PCoA results are almost exactly the same! (tests with 2 iterations)

# I tested bringing back all values for genera that we kept in the community matrix, but this is not a good idea. First, rarefaction functions of vegan need counts of 1, otherwise they don't work. So we would have to insert a dummy variable. # I tried for alpha diversity, see NOTE after alpha diversity plot.
# For beta diversity, adding a dummy variable in this case did not change the PERMANOVA and PCoA significantly.
# AGAIN, in any cases, my PERMANOVA and PCoA results are consistent and robust to the details of my method in this script chunk.
#bracken.reads.clean.2 <- bracken.reads[, colnames(bracken.reads.clean)]
#bracken.reads.clean <- bracken.reads.clean.2
#bracken.reads.clean$dummy <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)

# Statistics
# On 'bracken.reads.clean'
rowSums(bracken.reads.clean)/rowSums(bracken.reads)*100
mean(rowSums(bracken.reads.clean)/rowSums(bracken.reads)*100)
sum(bracken.reads.clean)/sum(bracken.reads)*100
length(colnames(bracken.reads.clean))/length(colnames(bracken.reads))*100
# By getting rid of the noise, we keep approx. 82.07% of the reads, with 198 (9.59%) genera out of 2065 originally 'detected'
```


## Alpha diversity

### Rarefaction
```{r, eval=FALSE}
rowSums(bracken.reads.clean)
rarefy.alpha <- rarefy(bracken.reads.clean, sample = min(rowSums(bracken.reads.clean)), se = T)
rarefy.alpha <- data.frame(as.matrix(t(rarefy.alpha)))
```

Save rarefy.alpha dataframe for further code rerun. Load RDS.
```{r}
#saveRDS(rarefy.alpha, "../rds/rarefy.alpha.rds")
rarefy.alpha <- readRDS("../rds/rarefy.alpha.rds")
```

```{r}
rarefy.alpha$non.absence <- 198 - rowSums(bracken.reads.clean == 1)
rarefy.alpha$nb.reads.clean <- rowSums(bracken.reads.clean)
```

```{r}
lm <- lm(rarefy.alpha$S ~ metadata.bac.comm$host_species)
summary(lm)
hist(residuals(lm))
```

```{r}
plot(rarefy.alpha$S ~ metadata.bac.comm$host_species)
```

# SUPER DUPER IMPORTANT NOTE
The way the data is structured, if I keep '1' values for low abundant genera that are kept in my analysis (dataframe 'bracken.reads.clean'), the S (expected richness) values I am not sure they make total sense and if I can trust them. If I add a dummy variable just so the function can work, than the genera richness is the same across all samples.
Therefore, I don't feel like I can confidently analyze alpha diversity with bracken output, at least for now.

BUT anyhow, is it really important to evaluate genera richness among hosts? not really. Compositional differences are most interesting!



## Beta diversity

### Rarefaction
```{r, eval=FALSE}
bracken.dist <- avgdist(bracken.reads.clean, sample = min(rowSums(bracken.reads.clean)), dmethod = "bray", iteration = 100)
#saveRDS(bracken.dist, "../rds/bracken_dist_iterations_100.rds")
```

```{r}
bracken.dist <- readRDS("../rds/bracken_dist_iterations_100.rds")
```


## PERMANOVA
```{r}
bacteria.perm <- adonis2(bracken.dist ~ host_species, data = metadata)
bacteria.perm
```

## BETADISPER
```{r}
bacteria.disper <- betadisper(bracken.dist, metadata$host_species)
anova(bacteria.disper)
plot(bacteria.disper)
```
## Pairwise PERMANOVAs
```{r}
shrub_tree_PERM <- adonis2(bracken.dist ~ shrub_tree, data = metadata.bac.comm)
shrub_tree_PERM
shrub_tree_DISP <- betadisper(bracken.dist, metadata.bac.comm$shrub_tree)
anova(shrub_tree_DISP)
plot(shrub_tree_DISP)

#bracken.dist.noCOCO <- avgdist(bracken.reads.clean[c(1:10,16:25),], sample = min(rowSums(bracken.reads.clean[c(1:10,16:25),])), dmethod = "bray", iteration = 100)
#saveRDS(bracken.dist, "../rds/bracken_dist_iterations_100.rds")
#bracken.dist.noCOCO <- readRDS("../rds/bracken_dist_iterations_100.rds")
bracken.dist.noCOCO <- data.frame(as.matrix(bracken.dist))
bracken.dist.noCOCO <- bracken.dist.noCOCO[c(1:10,16:25),c(1:10,16:25)]
bracken.dist.noCOCO <- as.dist(as.matrix(bracken.dist.noCOCO), diag = FALSE, upper = FALSE)
gymno_angio_tree_PERM <- adonis2(bracken.dist.noCOCO ~ gymno_angio_tree, data = metadata.bac.comm[c(1:10,16:25),])
gymno_angio_tree_PERM
gymno_angio_tree_DISP <- betadisper(bracken.dist.noCOCO, metadata.bac.comm[c(1:10,16:25),]$gymno_angio_tree)
anova(gymno_angio_tree_DISP)
plot(gymno_angio_tree_DISP, label = F)


bracken.dist.gymno <- data.frame(as.matrix(bracken.dist))
bracken.dist.gymno <- bracken.dist.gymno[c(1:5,21:25),c(1:5,21:25)]
bracken.dist.gymno <- as.dist(as.matrix(bracken.dist.gymno), diag = FALSE, upper = FALSE)
gymno_PERM <- adonis2(bracken.dist.gymno ~ host_species, data = metadata.bac.comm[c(1:5,21:25),])
gymno_PERM
gymno_DISP <- betadisper(bracken.dist.gymno, metadata.bac.comm[c(1:5,21:25),]$host_species)
anova(gymno_DISP)
plot(gymno_DISP, label = F)

bracken.dist.angio <- data.frame(as.matrix(bracken.dist))
bracken.dist.angio <- bracken.dist.angio[c(6:10,16:20),c(6:10,16:20)]
bracken.dist.angio <- as.dist(as.matrix(bracken.dist.angio), diag = FALSE, upper = FALSE)
angio_PERM <- adonis2(bracken.dist.angio ~ host_species, data = metadata.bac.comm[c(6:10,16:20),])
angio_PERM
angio_DISP <- betadisper(bracken.dist.angio, metadata.bac.comm[c(6:10,16:20),]$host_species)
anova(angio_DISP)
plot(angio_DISP, label = F)


```

## PCoA
```{r}
bacteria.pcoa <- cmdscale(bracken.dist, k =(nrow(bracken.reads.clean) - 1), eig = TRUE)
pcoa.eig <- eigenvals(bacteria.pcoa) # extract eigenvalues
pcoa.eig <- abs(pcoa.eig) # absolute values
sum.pcoa.eig <- sum(pcoa.eig) # total of eigenvalues
pcoa.prop.exp <- pcoa.eig/sum.pcoa.eig # proportion explained by each axis
(pcoa.axis1 <- pcoa.prop.exp[1])
(pcoa.axis2 <- pcoa.prop.exp[2])
site.scores.pcoa <- scores(bacteria.pcoa) # site scores
sp.scores.pcoa <- wascores(bacteria.pcoa$points, bracken.reads.clean) # species scores 
site.scores.pcoa <- data.frame(site.scores.pcoa)
sp.scores.pcoa <- data.frame(sp.scores.pcoa)
host_species <- metadata.bac.comm$host_species
site.scores.pcoa <- cbind(site.scores.pcoa, host_species)
pcoa.ggordiplot <- gg_ordiplot(bacteria.pcoa, choices = c(1,2), groups = metadata.bac.comm$host_species, kind = "sd", conf = 0.95, pt.size = 3)
pcoa.ellipses <- pcoa.ggordiplot$df_ellipse # extract confidence ellipse
```

Plot
```{r}
ggplot(data = site.scores.pcoa, 
         aes(x = Dim1, y = Dim2), 
         color = host_species) + 
  expand_limits(x=c(-0.4,0.4), y=c(-0.2,0.3)) +
  geom_point(aes(shape = host_species, color = host_species), show.legend = T, 
             size=3) +
  xlab(paste("PCoA1 (",round(pcoa.axis1*100, digits = 2),"%)", sep = "")) +
  ylab(paste("PCoA2 (",round(pcoa.axis2*100, digits = 2),"%)", sep = "")) +
  #geom_text_repel(data = sp.scores.pcoa, 
  #                aes(x = X1, y = X2), 
  #                label=rownames(sp.scores.pcoa), 
  #                color = "black") +
  scale_shape_manual(values = c(16,16,16,16,16)) +
  guides(shape = "none") +
  geom_path(data = pcoa.ellipses, 
            aes(x = x, y = y, color = Group), 
            show.legend = FALSE) + 
  scale_color_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"), name = "Host species", labels = c("ABBA" = expression(italic("A. balsamea")), "THOC" = expression(italic("T. occidentalis")),"ACSA" = expression(italic("A. Saccharum")), "FAGR" = expression(italic("F. grandifolia")), "COCO" = expression(italic("C. cornuta")))) +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.text = element_text(size = 11),
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/bacteria_pcoa.tiff", width = 6, height = 4)
```

################################################################################

dbRDA tests

### dbRDA
```{r}
dbrda <- dbrda(bracken.dist ~ host_species, data = metadata.bac.comm)
summary(dbrda)
plot(dbrda)
anova.cca(dbrda, by = "terms", permutations = 999)
RsquareAdj(dbrda)
```

Extract proportions explained by axes 1 and 2
```{r}
dbrda.eig <- eigenvals(dbrda, model = "constrained") # extract eigenvalues
dbrda.eig <- abs(dbrda.eig) # calculate eigenvalues absolute values
dbrda.eig.sum <- sum(dbrda.eig) # total of all eigenvalues
dbrda.prop.exp <- dbrda.eig/dbrda.eig.sum # proportion explained by every axes
(dbrda.axis1 <- dbrda.prop.exp[1])
(dbrda.axis2 <- dbrda.prop.exp[2])
```

Extract dbRDA scores
```{r}
dbrda.site.scores <- data.frame(scores(dbrda, display = "sites", scaling = "sites")) # site scores
#sppscores(dbrda) <- GC.agg.acc.count.rar
#dbrda.sp.scores <- data.frame(scores(dbrda, display = "species", scaling = "sites")) # species scores 
#dbrda.sp.scores.DA <- dbrda.sp.scores[rownames(DA.KM),] # Keep only KEGG Module which are differentially abundant in DESeq2 analysis
```

Add host_species factor in site scores data frame for ggplot 
```{r}
host_species <- metadata.bac.comm$host_species
dbrda.site.scores <- cbind(dbrda.site.scores,host_species)
rm(host_species)
```

dbRDA plot
```{r}
dbrda.plot <- 
  ggplot(data = dbrda.site.scores, 
         aes(x = dbRDA1, y = dbRDA2), 
         color = host_species) + 
  expand_limits(x=c(-0.5,0.5), y=c(-0.3,0.4)) +
  geom_point(aes(shape = host_species, color = host_species), 
             size=4) +
  xlab("dbRDA 1 (%)") + 
  ylab("dbRDA 2 (%)") +
  #geom_text(data = dbrda.sp.scores.DA, 
  #                aes(x = CAP1, y = CAP2), 
  #                label=rownames(dbrda.sp.scores.DA), 
  #                color = "black") +
  #geom_segment(data = vect.pcoa.env, 
  #             aes(x = 0, xend = Dim1, y = 0, yend = Dim3),
  #             arrow = arrow(length = unit(0.75, "cm")), 
  #             color = "black") +
  #geom_text_repel(data = vect.pcoa.env, 
  #                aes(x = Dim1, y = Dim3, label = variable.env, size = 3), 
  #                color = "firebrick1", 
  #                fontface = "bold", 
  #                show.legend = FALSE) +
  scale_shape_manual(values = c(17,16,16,16,17)) +
  #geom_path(data = pcoa1.ellipses, 
  #          aes(x = x, y = y, color = Group), 
  #          show.legend = FALSE) + 
  scale_color_manual(values=c("turquoise1","red1", "gold1","orchid1","limegreen")) +
  dark_theme_linedraw() +
  theme(axis.text.x = element_text(color = "grey80", size = 10, angle = 0, hjust = 0.5, vjust = 0),
        axis.text.y = element_text(color = "grey80", size = 10, angle = 0, hjust = 0, vjust = 0.5),  
        axis.title.x = element_text(color = "grey80", size = 14, angle = 0, hjust = 0.5, vjust = -1),
        axis.title.y = element_text(color = "grey80", size = 14, angle = 90, hjust = 0.5, vjust = 1),
        plot.background = element_rect(fill = "grey15"),
        panel.background = element_blank(),
        legend.background = element_blank(),
        panel.grid.major = element_line(size = 0, linetype = 'solid', colour = "grey90"), 
        panel.grid.minor = element_line(size = 0, linetype = 'solid', colour = "grey90"))
dbrda.plot
#ggsave("../dbrda.plot.tiff")
```











################################################################################
# --------- BACKUP BELOW
BELOW NOT USED FOR NOW, KEPT AS BACKUP




```{r}




# In summary, we kept only genera (and their read counts) that represented a relative abundance > 0.5% in at least one sample.

rowSums(bracken.reads)
rowSums(bracken.reads.pm)
rowSums(bracken.reads.2)

# % reads lost 
(1 - (rowSums(bracken.reads.2)/rowSums(bracken.reads)))*100

# NOW, IN NEED TO LUMP LOST GENERA INTO "OTHER"CATEGORY TO USE TO VISUALIZE RELATIVE ABUNDANCES (BARPLOT) AND TO COMPARE RELATIVE ABUNDANCE OF METHYLO AMONG HOST SPECIES. THEN DELETE THIS OTHER COLUMN FOR ALPHA AND BETA DIVERSITY.

bracken.reads.others <- bracken.reads[,setdiff(colnames(bracken.reads), colnames(bracken.reads.2))]
braken.comm.r.others.2 <- data.frame(rowSums(bracken.reads.others))
colnames(braken.comm.r.others.2) <- "Others"
bracken.reads.3 <- bracken.reads.2
bracken.reads.3$Others <- braken.comm.r.others.2$Others


bracken.comm <- (bracken.reads.2/rowSums(bracken.reads.2))*100 # values are now relative abundances of kept genera, expressed in percentage
rowSums(bracken.comm)
colSums(bracken.comm)

braken.comm.filt <- bracken.comm
braken.comm.filt[braken.comm.filt < 0.05] <- 0 # put zero for all values < 0.05.

bracken.comm.others <- (bracken.reads.3/rowSums(bracken.reads.3))*100 # values are now relative abundances of kept genera, expressed in percentage



```



# Edit bracken df - fractions
```{r}
bracken.frac <- bracken.import[,-c(1,2)]
bracken.frac <- bracken.frac[,-c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29,31,33,35,37,39,41,43,45,47,49)]
colnames(bracken.frac) <- gsub("_frac", "", colnames(bracken.frac))
```

Transpose and filter data
```{r}
options(scipen = 999) # to reset sci. not. : options(scipen = 0)
bracken.comm <- t(bracken.frac)
bracken.comm <- as.data.frame(bracken.comm)
rowSums(bracken.comm)
bracken.comm <- bracken.comm[, !grepl("virus", colnames(bracken.comm))] # remove virus
bracken.comm <- bracken.comm[, colSums(bracken.comm)>0.01] # remove genera with less than 1% additive abundance across all samples
bracken.comm <- (bracken.comm/rowSums(bracken.comm))*100 # values are now relative abundances of kept genera, expressed in percentage
rowSums(bracken.comm)
colSums(bracken.comm)
```

ABOVE NOT USED FOR NOW, KEPT AS BACKUP
################################################################################











# Exploratory analysis of Methylobacterium relative abundance (%) in samples
```{r}
plot(bracken.comm[,"Methylobacterium"] ~ metadata.bac.comm$host_species)
```








################################################################################

# Alpha diversity

## Indices calculation

### Richness and Shannon Index
```{r}
bacteria.PA.comm <- data.frame(ifelse(bacteria.comm > 0, 1, 0))
metadata.bac.comm$richness <- rowSums(bacteria.PA.comm)
metadata.bac.comm$shannon <- diversity(bacteria.comm, index = "shannon")
hist(metadata.bac.comm$richness, breaks = 25)
shapiro.test(metadata.bac.comm$richness)
hist(metadata.bac.comm$shannon, breaks = 25)
shapiro.test(metadata.bac.comm$shannon)
```
Richness and shannon values are normally distributed.

### Pielou Index
```{r}
metadata.bac.comm$pielou <- metadata.bac.comm$shannon/log(metadata.bac.comm$richness)
hist(metadata.bac.comm$pielou, breaks = 25)
shapiro.test(metadata.bac.comm$pielou)
```
Pielou values are normally distributed.

## Models

### Richness
```{r}
lm <- lm(richness ~ host_species, data = metadata.bac.comm)
plot(richness ~ host_species, data = metadata.bac.comm)
summary(lm)
anova(lm)
plot(lm)
hist(residuals(lm))
shapiro.test(residuals(lm))
```
Linear model's assumptions are met.


### Shannon
```{r}
lm <- lm(shannon ~ host_species, data = metadata.bac.comm)
plot(shannon ~ host_species, data = metadata.bac.comm)
summary(lm)
anova(lm)
plot(lm)
hist(residuals(lm))
shapiro.test(residuals(lm))
```
Linear model's assumptions are met.

Plot
```{r}
ggplot(metadata.bac.comm, aes(x = host_species, y = shannon)) +
geom_boxplot(alpha = 1, aes(fill = host_species), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = host_species), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
scale_fill_manual(values=c("#63AFF9","#CA0E78","#6104BB","#F39635","#1EA670"))+
scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Host Species", 
     y = "Shannon Index")+
ylim(0.5,2.5) +
theme_bw()+
theme(text = element_text(family = "Arial"),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/Bacterial_genera_shannon.tiff", width = 6, height = 4)
```



### Evenness
```{r}
lm <- lm(pielou ~ host_species, data = metadata.bac.comm)
plot(pielou ~ host_species, data = metadata.bac.comm)
summary(lm)
anova(lm)
plot(lm) 
hist(residuals(lm))
shapiro.test(residuals(lm))
```


################################################################################

# Beta diversity

With Methylobacterium 
Genera composition by host species (including Methylobacterium)

## Distance matrix
```{r}
bacteria.dist <- vegdist(bacteria.comm, method='bray')
```

## PERMANOVA
```{r}
bacteria.perm <- adonis2(bacteria.dist ~ host_species, data = metadata)
bacteria.perm
```

## BETADISPER
```{r}
bacteria.disper <- betadisper(bacteria.dist, metadata$host_species)
anova(bacteria.disper)
plot(bacteria.disper)
```
## Pairwise adonis
```{r}
pair.adonis.bac.bray <- pairwise.adonis(bacteria.dist, metadata.bac.comm$host_species)
pair.adonis.bac.bray
```


## PCoA
```{r}
bacteria.pcoa <- cmdscale(bacteria.dist, k =(nrow(bacteria.comm) - 1), eig = TRUE)
pcoa.eig <- eigenvals(bacteria.pcoa) # extract eigenvalues
pcoa.eig <- abs(pcoa.eig) # absolute values
sum.pcoa.eig <- sum(pcoa.eig) # total of eigenvalues
pcoa.prop.exp <- pcoa.eig/sum.pcoa.eig # proportion explained by each axis
(pcoa.axis1 <- pcoa.prop.exp[1])
(pcoa.axis2 <- pcoa.prop.exp[2])
site.scores.pcoa <- scores(bacteria.pcoa) # site scores
sp.scores.pcoa <- wascores(bacteria.pcoa$points, bacteria.comm) # species scores 
site.scores.pcoa <- data.frame(site.scores.pcoa)
sp.scores.pcoa <- data.frame(sp.scores.pcoa)
host_species <- metadata.bac.comm$host_species
site.scores.pcoa <- cbind(site.scores.pcoa, host_species)
pcoa.ggordiplot <- gg_ordiplot(bacteria.pcoa, choices = c(1,2), groups = metadata.bac.comm$host_species, kind = "sd", conf = 0.95, pt.size = 3)
pcoa.ellipses <- pcoa.ggordiplot$df_ellipse # extract confidence ellipse
```
Plot
```{r}
ggplot(data = site.scores.pcoa, 
         aes(x = Dim1, y = Dim2), 
         color = host_species) + 
  expand_limits(x=c(-0.8,0.8), y=c(-0.8,0.8)) +
  geom_point(aes(shape = host_species, color = host_species), show.legend = T, 
             size=3) +
  xlab(paste("PCoA1 (",round(pcoa.axis1*100, digits = 2),"%)", sep = "")) +
  ylab(paste("PCoA2 (",round(pcoa.axis2*100, digits = 2),"%)", sep = "")) +
  #geom_text_repel(data = sp.scores.pcoa, 
  #                aes(x = X1, y = X2), 
  #                label=rownames(sp.scores.pcoa), 
  #                color = "black") +
  scale_shape_manual(values = c(16,16,16,16,16)) +
  guides(shape = "none") +
  geom_path(data = pcoa.ellipses, 
            aes(x = x, y = y, color = Group), 
            show.legend = FALSE) + 
  scale_color_manual(values=c("#63AFF9","#CA0E78","#6104BB","#F39635","#1EA670"), name = "Host species") +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.text = element_text(size = 11),
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/bacteria_pcoa.tiff", width = 6, height = 4.5)
```

### Barplot
```{r}
abundances.bac <- picante::matrix2sample(bracken.comm)
colnames(abundances.bac) <- c("sample","abund","genus")

ggplot(abundances.bac, aes(x = sample, y = abund, fill = genus)) +
  geom_bar(stat = "identity",
#          position = "fill" # add to get relative abundances
  ) +
  labs(x = "Sample", y = "Relative abundance") +
  guides(fill = guide_legend(ncol = 1)) +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      legend.justification = c(0.5, 1),
      legend.text = element_text(size = 10.5),
      legend.key.size = unit(0.43, "cm"),
      legend.position = "right",  
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 11),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/bacteria_barplot.tiff", width = 8, height = 5.5)





  scale_fill_manual(values=c(
  "#CC3366", # Lichenibacterium
  "#FFB2B2", "#B3C2FF", "#FFE1B2", 
  "#0098A1", # Methylobacterium
  "#FFD4A3", "#FFC8A1", 
  "#FFDBA3", "#FFA07A", 
  "#FF9CDA", # Pseudomonas
  "#99C1FF",# Sphingomonas
  "#E7EBA3", "#A2B89D","#FF8B5C", "#F9AFA3", "#F6D2A6", "#D1D1B5",  "#E7C6A0", 
  "#A7C6A7", "#9BC6C8", "#B5C6D9", "#A4D8E1",  
  "#5BA2D6", "#7BAFFF", "#3A8CBB", "#A1E1D5", 
  "#E1D5B2", "#D9B3A0", "#E9B7A6", "#EAB0B5", "#E7C5A8", 
  "#C8B4A0", "#C2C8A1", "#B7D1A7",  "#B4D6B8", "#FF9E6D",
  "#B2D8E1", "#C2E6D5", "#C7D5E1", "#C4B2E1", "#B0C4E1", 
  "#9CB2E1", "#9DC6E1", "#A8D9D0", "#A6D2E1", "#D4A4D2", 
  "#D1A4D9", "#D2B2FF", "#FFCCB2", "#D5C2E1", "#E1B2C2", 
  "#E1A2B4", "#D6A1B8", "#EAB8C9", "#FFC9F5", "#FF9CDA", 
  "#FFA1B1", "#FFC5B8", "#FFC9E0", "#FFD7B3", "#D9C9A6",
  "#A1E1D5", "#E1D5B2", "#D9B3A0", "#E9B7A6", "#EAB0B5", 
  "#E7C5A8", "#C8B4A0", "#C2C8A1", "#B7D1A7", "#E7EBA3", 
  "#B4D6B8", "#B2D8E1", "#C2E6D5", "#C7D5E1", "#C4B2E1", 
  "#B0C4E1", "#9CB2E1", "#9DC6E1", "#A8D9D0", "#A6D2E1",  "#FFAA90",
  "#D4A4D2", "#D1A4D9", "#D2B2FF", "#B3C2FF", "#A1C6E1", 
  "#D5C2E1", "#E1B2C2", "#E1A2B4", "#D6A1B8", "#EAB8C9", 
  "#FFC9F5",  "#FFA1B1", "#FFC5B8", "#FFC9E0"
), name = "Bacterial genera") +
```




















################################################################################
CODE BELOW KEPT IN CASE OF....


Without Methylobacterium 
Other genera composition by host species (excluding Methylobacterium)

```{r}
bacteria.comm.others <- bacteria.comm[, colnames(bacteria.comm)!="Methylobacterium"]
```

Create distance matrix
```{r}
bacteria.dist.nometh <- vegdist(bacteria.comm.others, method='bray')
```

PERMANOVA
```{r}
bacteria.perm.nometh <- adonis2(bacteria.dist.nometh ~ host_species + plot, data = metadata)
bacteria.perm.nometh
```

BETADISPER
```{r}
bacteria.disper.nometh <- betadisper(bacteria.dist.nometh, metadata$host_species)
anova(bacteria.disper.nometh)
plot(bacteria.disper.nometh)
```

Conclusion

Whether I include or exclude Methylobacterium of the bacterial composition, I still get similar results : Host species significantly explains 36% of the variation in bacterial genera composition.

Extracting PCoA site scores
For testing the effect of total bacterial community (excluding Methylobacterium) on Methylobacterium's genetic/functional composition (analyses in other scripts).

PCoA
```{r}
bacteria.pcoa <- cmdscale(bacteria.dist.nometh, k =(nrow(bacteria.comm.others) - 1), eig = TRUE)
```

Extract PCoA scores
```{r}
bacteria.pcoa.site.scores <- data.frame(scores(bacteria.pcoa, display = "sites", choices = c(1,2))) # site scores for axes 1 and 2
```

Save scores as RDS file.
```{r}
saveRDS(bacteria.pcoa.site.scores, "../bacteria.pcoa.site.scores.rds")
```

