---
title: "Combine variability profiles"
output: html_document
date: "2024-11-14"
---

-################################################################################

# Libraries
```{r}
library(tidyverse)
```

# Unzipping files

Four files must be unzipped before running this script.
var_cdn_sp_018.zip
var_cdn_sp_021.zip
var_cdn_sp_022.zip
var_cdn_sp_024.zip
These files are stored in '../data/anvio_outputs/'

-################################################################################

# M_sp_018

## Import data from anvio
```{r}
var.prof.ABBA <- read.table("../data/anvio_outputs/var_cdn_sp_018/var_cdn_sp_018_ABBA", header = T)
var.prof.ACSA <- read.table("../data/anvio_outputs/var_cdn_sp_018/var_cdn_sp_018_ACSA", header = T)
var.prof.COCO <- read.table("../data/anvio_outputs/var_cdn_sp_018/var_cdn_sp_018_COCO", header = T)
var.prof.FAGR <- read.table("../data/anvio_outputs/var_cdn_sp_018/var_cdn_sp_018_FAGR", header = T)
var.prof.THOC <- read.table("../data/anvio_outputs/var_cdn_sp_018/var_cdn_sp_018_THOC", header = T)
var.prof.all <- rbind(var.prof.ABBA,var.prof.ACSA,var.prof.COCO,var.prof.FAGR,var.prof.THOC)
var.prof.all$sample_id <- as.factor(var.prof.all$sample_id)
var.prof.all <- var.prof.all %>% arrange(corresponding_gene_call, sample_id)
var.prof.all$corresponding_gene_call <- as.factor(var.prof.all$corresponding_gene_call)
str(var.prof.all)
```

## Export data to anvio
```{r}
#write_delim(var.prof.all, file = "../data/anvio_inputs/variability_profile_all_samples_M_sp_018.txt", delim = "\t")
```

-################################################################################

# M_sp_021

## Import data from anvio
```{r}
var.prof.ABBA <- read.table("../data/anvio_outputs/var_cdn_sp_021/var_cdn_sp_021_ABBA", header = T)
var.prof.ACSA <- read.table("../data/anvio_outputs/var_cdn_sp_021/var_cdn_sp_021_ACSA", header = T)
var.prof.COCO <- read.table("../data/anvio_outputs/var_cdn_sp_021/var_cdn_sp_021_COCO", header = T)
var.prof.FAGR <- read.table("../data/anvio_outputs/var_cdn_sp_021/var_cdn_sp_021_FAGR", header = T)
var.prof.THOC <- read.table("../data/anvio_outputs/var_cdn_sp_021/var_cdn_sp_021_THOC", header = T)
var.prof.all <- rbind(var.prof.ABBA,var.prof.ACSA,var.prof.COCO,var.prof.FAGR,var.prof.THOC)
var.prof.all$sample_id <- as.factor(var.prof.all$sample_id)
var.prof.all <- var.prof.all %>% arrange(corresponding_gene_call, sample_id)
var.prof.all$corresponding_gene_call <- as.factor(var.prof.all$corresponding_gene_call)
str(var.prof.all)
```

## Export data to anvio
```{r}
#write_delim(var.prof.all, file = "../data/anvio_inputs/variability_profile_all_samples_M_sp_021.txt", delim = "\t")
```

-################################################################################

# M_sp_022

## Import data from anvio
```{r}
var.prof.ABBA <- read.table("../data/anvio_outputs/var_cdn_sp_022/var_cdn_sp_022_ABBA", header = T)
var.prof.ACSA <- read.table("../data/anvio_outputs/var_cdn_sp_022/var_cdn_sp_022_ACSA", header = T)
var.prof.COCO <- read.table("../data/anvio_outputs/var_cdn_sp_022/var_cdn_sp_022_COCO", header = T)
var.prof.FAGR <- read.table("../data/anvio_outputs/var_cdn_sp_022/var_cdn_sp_022_FAGR", header = T)
var.prof.THOC <- read.table("../data/anvio_outputs/var_cdn_sp_022/var_cdn_sp_022_THOC", header = T)
var.prof.all <- rbind(var.prof.ABBA,var.prof.ACSA,var.prof.COCO,var.prof.FAGR,var.prof.THOC)
var.prof.all$sample_id <- as.factor(var.prof.all$sample_id)
var.prof.all <- var.prof.all %>% arrange(corresponding_gene_call, sample_id)
var.prof.all$corresponding_gene_call <- as.factor(var.prof.all$corresponding_gene_call)
str(var.prof.all)
```

## Export data to anvio
```{r}
#write_delim(var.prof.all, file = "../data/anvio_inputs/variability_profile_all_samples_M_sp_022.txt", delim = "\t")
```

-################################################################################

# M_sp_024

## Import data from anvio
```{r}
var.prof.ABBA <- read.table("../data/anvio_outputs/var_cdn_sp_024/var_cdn_sp_024_ABBA", header = T)
var.prof.ACSA <- read.table("../data/anvio_outputs/var_cdn_sp_024/var_cdn_sp_024_ACSA", header = T)
var.prof.COCO <- read.table("../data/anvio_outputs/var_cdn_sp_024/var_cdn_sp_024_COCO", header = T)
var.prof.FAGR <- read.table("../data/anvio_outputs/var_cdn_sp_024/var_cdn_sp_024_FAGR", header = T)
var.prof.THOC <- read.table("../data/anvio_outputs/var_cdn_sp_024/var_cdn_sp_024_THOC", header = T)
var.prof.all <- rbind(var.prof.ABBA,var.prof.ACSA,var.prof.COCO,var.prof.FAGR,var.prof.THOC)
var.prof.all$sample_id <- as.factor(var.prof.all$sample_id)
var.prof.all <- var.prof.all %>% arrange(corresponding_gene_call, sample_id)
var.prof.all$corresponding_gene_call <- as.factor(var.prof.all$corresponding_gene_call)
str(var.prof.all)
```

## Export data to anvio
```{r}
#write_delim(var.prof.all, file = "../data/anvio_inputs/variability_profile_all_samples_M_sp_024.txt", delim = "\t")
```

-################################################################################