---
title: "data_preparation"
output: html_document
date: "2024-04-16"
---

-################################################################################

# Librairies
```{r}
library(tidyverse)
library(data.table)
```

-################################################################################

# Import Data

## Metapangenomes

Import individual tree species metapangenomes
```{r}
# ABBA
ABBA_summary <- read.delim("../data/anvio_outputs/Methylobacterium-ABBA-PAN/Methylobacterium-ABBA-METAPAN-SUMMARY/Methylobacterium-ABBA-PAN_gene_clusters_summary.txt.gz", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
ABBA_summary <- ABBA_summary[,-1]
# ACSA
ACSA_summary <- read.delim("../data/anvio_outputs/Methylobacterium-ACSA-PAN/Methylobacterium-ACSA-METAPAN-SUMMARY/Methylobacterium-ACSA-PAN_gene_clusters_summary.txt.gz", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
ACSA_summary <- ACSA_summary[,-1]
# COCO
COCO_summary <- read.delim("../data/anvio_outputs/Methylobacterium-COCO-PAN/Methylobacterium-COCO-METAPAN-SUMMARY/Methylobacterium-COCO-PAN_gene_clusters_summary.txt.gz", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
COCO_summary <- COCO_summary[,-1]
# FAGR
FAGR_summary <- read.delim("../data/anvio_outputs/Methylobacterium-FAGR-PAN/Methylobacterium-FAGR-METAPAN-SUMMARY/Methylobacterium-FAGR-PAN_gene_clusters_summary.txt.gz", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
FAGR_summary <- FAGR_summary[,-1]
# THOC
THOC_summary <- read.delim("../data/anvio_outputs/Methylobacterium-THOC-PAN/Methylobacterium-THOC-METAPAN-SUMMARY/Methylobacterium-THOC-PAN_gene_clusters_summary.txt.gz", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
THOC_summary <- THOC_summary[,-1]
```

Since these 5 metagenomes files are similar, we will work with only one from nom on, but merge gene type (core vs. accessory), gene coverage depth, and gene horizontal coverage data.

Remove unnecessary metagenomes files
```{r}
rm(ACSA_summary, COCO_summary, FAGR_summary, THOC_summary)
```

Edit ABBA_summary
```{r}
data <- ABBA_summary
# Remove unwanted columns
data <- subset(data, select = -c(2,12:15))
# Change vectors to factors
data <- data %>% mutate(across(c(gene_cluster_id, genome_name, gene_callers_id, KEGG_Class_ACC, KEGG_Class, KEGG_BRITE_ACC, KEGG_BRITE, KEGG_Module_ACC, KEGG_Module, KOfam_ACC, KOfam, Pfam_ACC, Pfam, CAZyme_ACC, CAZyme, COG20_FUNCTION_ACC, COG20_FUNCTION, COG20_PATHWAY_ACC, COG20_PATHWAY, COG20_CATEGORY_ACC, COG20_CATEGORY), as.factor))
# Change vectors to numeric
data <- data %>% mutate(across(c(num_genomes_gene_cluster_has_hits, num_genes_in_gene_cluster, max_num_paralogs, SCG), as.numeric))
str(data)
```

## mOTUpan data
```{r}
mOTUpan.import <- read.delim("../data/anvio_outputs/core_pan_genome_table", header = TRUE, sep = "\t", dec = ".", quote="")
colnames(mOTUpan.import)[1] <- "gene_cluster_id"
mOTUpan <- mOTUpan.import[,-7]
mOTUpan <- mOTUpan %>% mutate(across(c(gene_cluster_id, type), as.factor))
str(mOTUpan)
```

## Gene 'detection' (horizontal coverage)

Create new factorial vector with 104 genome names
```{r}
genomes <- unique(data$genome_name)
```

Import gene detection data
```{r}
# ABBA
ABBA.txtFileName <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/ABBA/ABBA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
ABBA.gene.detect <- rbindlist(sapply(ABBA.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ABBA.gene.detect$gene_callers_id <- as.factor(ABBA.gene.detect$gene_callers_id)

# ACSA
ACSA.txtFileName <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/ACSA/ACSA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
ACSA.gene.detect <- rbindlist(sapply(ACSA.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ACSA.gene.detect$gene_callers_id <- as.factor(ACSA.gene.detect$gene_callers_id)

# COCO
COCO.txtFileName <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/COCO/COCO_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
COCO.gene.detect <- rbindlist(sapply(COCO.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
COCO.gene.detect$gene_callers_id <- as.factor(COCO.gene.detect$gene_callers_id)

# FAGR
FAGR.txtFileName <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/FAGR/FAGR_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
FAGR.gene.detect <- rbindlist(sapply(FAGR.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
FAGR.gene.detect$gene_callers_id <- as.factor(FAGR.gene.detect$gene_callers_id)

# THOC
THOC.txtFileName <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/THOC/THOC_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
THOC.gene.detect <- rbindlist(sapply(THOC.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
THOC.gene.detect$gene_callers_id <- as.factor(THOC.gene.detect$gene_callers_id)

rm(ABBA.txtFileName,ACSA.txtFileName,COCO.txtFileName,FAGR.txtFileName,THOC.txtFileName)
```



-################################################################################

# Edit dataframes

## Merge 'mOTUpan' with 'data'
```{r}
data <- merge(data, mOTUpan, by = "gene_cluster_id")








#data <- data[c(1:3,50:54,4:49)] # rearrange column order
```

## Create data.pangenome RDS file
```{r}
data.pangenome <- data[c(1:3,30:34,4:10)]
#saveRDS(data.pangenome, "../rds/data.pangenome.rds")
data.pangenome <- readRDS("../rds/data.pangenome.rds")
```













## Functions and gene coverage stats df
```{r}
data.fct.cov <- subset(data, select = c(1:5,36:54))
data.fct.cov <- data.fct.cov[,c(1:5,10:13,22:23,8:9,6:7,14:17,18:21,24)] # rearrange column order
# Import eggNOG functions 
data.fct.cov <- merge(data.fct.cov, eggNOG, by = "gene_callers_id", all.x = TRUE)
data.fct.cov <- data.fct.cov[order(match(data.fct.cov$gene_callers_id, data$gene_callers_id)), ]
rownames(data.fct.cov) <- NULL
data.fct.cov <- data.fct.cov[,c(2,1,3,4:41)] # rearrange column order
```


Save dataframes as RDS files for subsequent analyses.
```{r}
#saveRDS(data.fct.cov, "../rds/data.fct.cov.rds")
```





################################################################################

# Gene detection

Load 'data.fct.cov'
```{r}
data.fct.cov <- readRDS("../rds/data.fct.cov.rds")
```

Create new factorial vector with 104 genome names
```{r}
genomes <- unique(data.fct.cov$genome_name)
```

Import gene detection data
```{r}
# ABBA
ABBA.txtFileName <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/ABBA/ABBA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
ABBA.gene.detect <- rbindlist(sapply(ABBA.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ABBA.gene.detect$gene_callers_id <- as.factor(ABBA.gene.detect$gene_callers_id)

# ACSA
ACSA.txtFileName <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/ACSA/ACSA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
ACSA.gene.detect <- rbindlist(sapply(ACSA.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ACSA.gene.detect$gene_callers_id <- as.factor(ACSA.gene.detect$gene_callers_id)

# COCO
COCO.txtFileName <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/COCO/COCO_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
COCO.gene.detect <- rbindlist(sapply(COCO.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
COCO.gene.detect$gene_callers_id <- as.factor(COCO.gene.detect$gene_callers_id)

# FAGR
FAGR.txtFileName <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/FAGR/FAGR_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
FAGR.gene.detect <- rbindlist(sapply(FAGR.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
FAGR.gene.detect$gene_callers_id <- as.factor(FAGR.gene.detect$gene_callers_id)

# THOC
THOC.txtFileName <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/THOC/THOC_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
THOC.gene.detect <- rbindlist(sapply(THOC.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
THOC.gene.detect$gene_callers_id <- as.factor(THOC.gene.detect$gene_callers_id)

rm(ABBA.txtFileName,ACSA.txtFileName,COCO.txtFileName,FAGR.txtFileName,THOC.txtFileName)
```

Create gene.detection dataframe
```{r}
ABBA.gene.detect <- ABBA.gene.detect[,-1]
ACSA.gene.detect <- ACSA.gene.detect[,-1]
COCO.gene.detect <- COCO.gene.detect[,-1]
FAGR.gene.detect <- FAGR.gene.detect[,-1]
THOC.gene.detect <- THOC.gene.detect[,-1]
gene.detection <- merge(ABBA.gene.detect,ACSA.gene.detect, by= "gene_callers_id")
gene.detection <- merge(gene.detection,COCO.gene.detect, by= "gene_callers_id")
gene.detection <- merge(gene.detection,FAGR.gene.detect, by= "gene_callers_id")
gene.detection <- merge(gene.detection,THOC.gene.detect, by= "gene_callers_id")
```

Merge gene.detection and data
```{r}
data.fct.det <- merge(data.fct.cov, gene.detection, by= "gene_callers_id")
data.fct.det <- data.fct.det[order(match(data.fct.det$gene_callers_id, data.fct.cov$gene_callers_id)), ]
rownames(data.fct.det) <- NULL
data.fct.det <- data.fct.det[,c(2,1,3,4:66)] # rearrange column order
```

Remove
```{r}
rm(ABBA.gene.detect,ACSA.gene.detect,COCO.gene.detect,FAGR.gene.detect,THOC.gene.detect,gene.detection)
```

Save dataframe
```{r}
#saveRDS(data.fct.det, file = "../rds/data.fct.det.rds")
```

################################################################################

# Gene coverage

Import gene coverage data
```{r}
# ABBA
ABBA.txtFileName2 <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/ABBA/ABBA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_coverages.txt",sep="")
ABBA.gene.cov <- rbindlist(sapply(ABBA.txtFileName2, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ABBA.gene.cov$gene_callers_id <- as.factor(ABBA.gene.cov$gene_callers_id)

# ACSA
ACSA.txtFileName2 <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/ACSA/ACSA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_coverages.txt",sep="")
ACSA.gene.cov <- rbindlist(sapply(ACSA.txtFileName2, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ACSA.gene.cov$gene_callers_id <- as.factor(ACSA.gene.cov$gene_callers_id)

# COCO
COCO.txtFileName2 <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/COCO/COCO_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_coverages.txt",sep="")
COCO.gene.cov <- rbindlist(sapply(COCO.txtFileName2, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
COCO.gene.cov$gene_callers_id <- as.factor(COCO.gene.cov$gene_callers_id)

# FAGR
FAGR.txtFileName2 <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/FAGR/FAGR_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_coverages.txt",sep="")
FAGR.gene.cov <- rbindlist(sapply(FAGR.txtFileName2, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
FAGR.gene.cov$gene_callers_id <- as.factor(FAGR.gene.cov$gene_callers_id)

# THOC
THOC.txtFileName2 <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/THOC/THOC_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_coverages.txt",sep="")
THOC.gene.cov <- rbindlist(sapply(THOC.txtFileName2, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
THOC.gene.cov$gene_callers_id <- as.factor(THOC.gene.cov$gene_callers_id)

rm(ABBA.txtFileName2,ACSA.txtFileName2,COCO.txtFileName2,FAGR.txtFileName2,THOC.txtFileName2)
```

Create gene.coverage dataframe
```{r}
ABBA.gene.cov <- ABBA.gene.cov[,-1]
ACSA.gene.cov <- ACSA.gene.cov[,-1]
COCO.gene.cov <- COCO.gene.cov[,-1]
FAGR.gene.cov <- FAGR.gene.cov[,-1]
THOC.gene.cov <- THOC.gene.cov[,-1]
gene.coverage <- merge(ABBA.gene.cov,ACSA.gene.cov, by= "gene_callers_id")
gene.coverage <- merge(gene.coverage,COCO.gene.cov, by= "gene_callers_id")
gene.coverage <- merge(gene.coverage,FAGR.gene.cov, by= "gene_callers_id")
gene.coverage <- merge(gene.coverage,THOC.gene.cov, by= "gene_callers_id")
```

Merge gene.coverage and data
```{r}
data.fct.cov <- merge(data.fct.cov, gene.coverage, by= "gene_callers_id")
data.fct.cov <- data.fct.cov[order(match(data.fct.cov$gene_callers_id, data.fct.det$gene_callers_id)), ]
rownames(data.fct.cov) <- NULL
data.fct.cov <- data.fct.cov[,c(2,1,3,4:66)] # rearrange column order
```

Remove
```{r}
rm(ABBA.gene.cov,ACSA.gene.cov,COCO.gene.cov,FAGR.gene.cov,THOC.gene.cov,gene.coverage)
```

Save dataframe
```{r}
#saveRDS(data.fct.cov, file = "../rds/data.fct.cov.rds")
```


################################################################################

# Annotations statistics
```{r}
# anvio intergrated annotations
sum(data.fct.cov$COG20_FUNCTION_ACC!="")
sum(data.fct.cov$KOfam_ACC!="")
sum(data.fct.cov$CAZyme_ACC!="")
sum(data.fct.cov$Pfam_ACC!="")

# eggNOG annotations
test <- data.fct.cov
test[test=="-"] <- ""
sum(!is.na(test$COG_category))
sum(!is.na(test$KEGG_ko))
sum(!is.na(test$CAZy))
sum(!is.na(test$PFAMs))
```




################################################################################

# Missing gene calls

Verification of missing gene calls

See important note below chunk.

In 'TREE.gene.detect' objects (tables created after merging TREE profiles with anvi-merge), there is 594478 gene calls.
In 'TREE_summary' and 'data' objects (tables created after pangenomes construction with anvi-pan-genome), there is 594053 gene calls. 401 gene calls have been dropped because they are rRNA gene calls, and 24 gene calls have been lost for a yet unknown reason.
Below, I inspect what are those 24 missing gene calls.
```{r}
# Create table with 425 gene calls not included in metapangenome data
missing.gene.calls <- anti_join(ABBA.gene.detect, data.fct.cov, by = "gene_callers_id")
# Import all gene calls from contigs database that have functional annotation
all.functions <- read.delim("/mnt/storage/jocelyn/msc/anvio_artifacts/all_functions.txt", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
all.functions$gene_callers_id <- as.factor(all.functions$gene_callers_id)
# How many unique gene calls have an assigned function?
all.functions %>% distinct(gene_callers_id) # 472 447 unique gene calls are functionaly annotated
# Merge missing gene calls with functionaly annotated gene calls to see if missing gene calls have an annotated function
missing.gene.calls.functions <- merge(missing.gene.calls,all.functions, by = "gene_callers_id")

# Second analysis similar to the one above but comparing missing gene calls with all gene calls (functionaly annotated or not)

# Import all gene calls from contigs database
ABBA.txtFileNameX <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/ABBA/ABBA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_calls.txt",sep="")
gene.calls <- rbindlist(sapply(ABBA.txtFileNameX, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
gene.calls$gene_callers_id <- as.factor(gene.calls$gene_callers_id)
# Merge missing gene calls with all gene calls to see if missing gene calls have an annotated function
missing.gene.calls.functions.2 <- merge(missing.gene.calls,gene.calls, by = "gene_callers_id")
```
Conclusion : No missing gene calls in metapangenome have any annotated functions.

Remove
```{r}
rm(ABBA.gene.detect,ACSA.gene.detect,COCO.gene.detect,FAGR.gene.detect,THOC.gene.detect)
rm(missing.gene.calls,missing.gene.calls.functions,missing.gene.calls.functions.2,ABBA.txtFileNameX)
```

#IMPORTANT NOTE

Gene callers ID tracking through anvio pipeline.

-- After contigs database (anvi-gen-contigs-database) and HMM (anvi-run-hmms)

* 'prodigal' (597,777 gene calls)
* 'Ribosomal_RNA_23S' (202 gene calls)
* 'Ribosomal_RNA_16S' (199 gene calls)

For a total of 598 178 gene calls --> gene_callers_id from 0 to 598177
Prodigal genes (ORF) --> gene_callers_id from 0 to 597776
Ribosomal_RNA_23S --> gene_callers_id from 597777 to 597978
Ribosomal_RNA_16S --> gene_callers_id from 597979 to 598177


-- After profiling TREE mapping results (anvi-profile), so consequently after merging TREE profiles (anvi-merge)

Because choice of minimum contig length (-M) = 1000,
We went from 598178 to 594478 gene calls (this include rRNA genes) (loss of 3700 gene calls, 0.6% of total gene calls)


-- After summarizing (anvi-summarize)

Loss of 401 gene calls (rRNA gene calls)
We went from 594478 to 594077 gene calls


-- Gene clustering (anvi-pan-genome)

Loss of 24 gene calls 
We went from 594077 to 594053 gene calls
I don't know why those gene calls were lost BUT
Those 24 gene calls have no functions assigned (verified in the chunk above).



################################################################################













################################################################################

SECTION BELOW NOT DONE, BECAUSE USELESS.

Import gene non outlier coverage data
```{r}
# ABBA
ABBA.txtFileName3 <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/ABBA/ABBA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_non_outlier_coverages.txt",sep="")
ABBA.gene.no.cov <- rbindlist(sapply(ABBA.txtFileName3, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ABBA.gene.no.cov$gene_callers_id <- as.factor(ABBA.gene.no.cov$gene_callers_id)

# ACSA
ACSA.txtFileName3 <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/ACSA/ACSA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_non_outlier_coverages.txt",sep="")
ACSA.gene.no.cov <- rbindlist(sapply(ACSA.txtFileName3, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ACSA.gene.no.cov$gene_callers_id <- as.factor(ACSA.gene.no.cov$gene_callers_id)

# COCO
COCO.txtFileName3 <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/COCO/COCO_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_non_outlier_coverages.txt",sep="")
COCO.gene.no.cov <- rbindlist(sapply(COCO.txtFileName3, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
COCO.gene.no.cov$gene_callers_id <- as.factor(COCO.gene.no.cov$gene_callers_id)

# FAGR
FAGR.txtFileName3 <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/FAGR/FAGR_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_non_outlier_coverages.txt",sep="")
FAGR.gene.no.cov <- rbindlist(sapply(FAGR.txtFileName3, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
FAGR.gene.no.cov$gene_callers_id <- as.factor(FAGR.gene.no.cov$gene_callers_id)

# THOC
THOC.txtFileName3 <- paste("/mnt/storage/jocelyn/msc/anvio_analysis/THOC/THOC_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_non_outlier_coverages.txt",sep="")
THOC.gene.no.cov <- rbindlist(sapply(THOC.txtFileName3, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
THOC.gene.no.cov$gene_callers_id <- as.factor(THOC.gene.no.cov$gene_callers_id)

rm(ABBA.txtFileName3,ACSA.txtFileName3,COCO.txtFileName3,FAGR.txtFileName3,THOC.txtFileName3)
```

Create gene.no.coverage dataframe
```{r}
ABBA.gene.no.cov <- ABBA.gene.no.cov[,-1]
ACSA.gene.no.cov <- ACSA.gene.no.cov[,-1]
COCO.gene.no.cov <- COCO.gene.no.cov[,-1]
FAGR.gene.no.cov <- FAGR.gene.no.cov[,-1]
THOC.gene.no.cov <- THOC.gene.no.cov[,-1]
gene.no.coverage <- merge(ABBA.gene.no.cov,ACSA.gene.no.cov, by= "gene_callers_id")
gene.no.coverage <- merge(gene.no.coverage,COCO.gene.no.cov, by= "gene_callers_id")
gene.no.coverage <- merge(gene.no.coverage,FAGR.gene.no.cov, by= "gene_callers_id")
gene.no.coverage <- merge(gene.no.coverage,THOC.gene.no.cov, by= "gene_callers_id")
```

Merge gene.no.coverage and data
```{r}
data.gene.no.cov <- merge(data,gene.no.coverage, by= "gene_callers_id")
```

Save dataframe
```{r}
#saveRDS(data.gene.no.cov, file = "data.gene.no.cov.rds")
```
