---
title: "Genes diversity"
output: html_document
date: "2024-10-29"
---

-################################################################################

# Libraries
```{r}
library(tidyverse)
library(vegan)
library(picante)
library(ggordiplots)
#library(gridExtra)
#library(cowplot)
library(viridis)
#library(DESeq2)
#library(patchwork)
#library(EnhancedVolcano)
```

# Import data

## Horizontal coverage ('detection')
```{r}
data.fct.det <- readRDS("../rds/data.fct.det.rds")
rownames(data.fct.det) <- data.fct.det$gene_callers_id
```

## Coverage depth
```{r}
data.fct.cov <- readRDS("../rds/data.fct.cov.rds")
rownames(data.fct.cov) <- data.fct.cov$gene_callers_id
```

## Metadata
```{r}
metadata <- read.delim("../data/samples_metadata.txt", header = TRUE, sep = "\t", dec = ".", row.names = 1)
metadata <- metadata %>% arrange(row.names(.))
metadata <- metadata %>% mutate(across(c(plant_clade,plant_order,plant_family,host_species,tree), as.factor))
metadata$host_species <- factor(x = metadata$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
metadata$shrub_tree <- c("tree","tree","tree","tree","tree","tree","tree","tree","tree","tree","shrub","shrub","shrub","shrub","shrub","tree","tree","tree","tree","tree","tree","tree","tree","tree","tree")
metadata$shrub_tree <- factor(x = metadata$shrub_tree, levels = c("shrub","tree"))
metadata$gymno_angio_tree <- c("gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree")
metadata$gymno_angio_tree <- factor(x = metadata$gymno_angio_tree, levels = c("gymnosperm_tree","angiosperm_tree","angiosperm_shrub"))
```

-################################################################################

# Prepare data

Sanity check
```{r}
all(rownames(data.fct.cov)==rownames(data.fct.det))
all(colnames(data.fct.cov)==colnames(data.fct.det))
```

## Detection

For analyses, keep only genes that are horizontally covered at least 1X for 50% of nucleotides. Those genes are considered 'present'.

If det<0.5 = 0, if det>=0.5 = 1
```{r}
data.det.PA <- data.frame(ifelse(data.fct.det[,c(25:49)] < 0.5, 0, 1))
```

## Coverage

For analyses, keep coverage values only for genes that are 'present', then round the coverage values to integers (counts of genes).

If det=0, cov=0; if det=1, cov=cov
```{r}
data.cov.pre <- data.fct.cov[,c(25:49)]
data.cov <- data.frame(mapply(function(x, y) ifelse(x == 1, y, 0), data.det.PA, data.cov.pre))
rownames(data.cov) <- rownames(data.fct.cov)
```

## Round
```{r}
data.cov.c <- round(data.cov, digits=0)
```

## Merge coverage data
Merge coverage counts dataframe with functions
```{r}
all(rownames(data.fct.cov)==rownames(data.cov.c))
data.fct.cov.c <- cbind(data.fct.cov[,c(1:24)], data.cov.c)
saveRDS(data.fct.cov.c, "../rds/data.fct.cov.c.rds")
```

## Aggregate gene counts by GC

Aggregate (sum) count values by GC
```{r}
data.fct.cov.c %>% distinct(gene_cluster_id) # There is 70,695 gene clusters in Methylobacterium pangenome
data.cov.c.agg <- aggregate(cbind(data.fct.cov.c$ABBA_M11,data.fct.cov.c$ABBA_M12,data.fct.cov.c$ABBA_M13,data.fct.cov.c$ABBA_M14,data.fct.cov.c$ABBA_M15,data.fct.cov.c$ACSA_M21,data.fct.cov.c$ACSA_M22,data.fct.cov.c$ACSA_M23,data.fct.cov.c$ACSA_M24,data.fct.cov.c$ACSA_M25,data.fct.cov.c$COCO_M01,data.fct.cov.c$COCO_M02,data.fct.cov.c$COCO_M03,data.fct.cov.c$COCO_M04,data.fct.cov.c$COCO_M05,data.fct.cov.c$FAGR_M16,data.fct.cov.c$FAGR_M17,data.fct.cov.c$FAGR_M18,data.fct.cov.c$FAGR_M19,data.fct.cov.c$FAGR_M20,data.fct.cov.c$THOC_M06,data.fct.cov.c$THOC_M07,data.fct.cov.c$THOC_M08,data.fct.cov.c$THOC_M09,data.fct.cov.c$THOC_M10),  by = list(data.fct.cov.c$gene_cluster_id), sum)
colnames(data.cov.c.agg) <- c("gene_cluster_id","ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","ACSA_M21","ACSA_M22","ACSA_M23","ACSA_M24","ACSA_M25","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05","FAGR_M16","FAGR_M17","FAGR_M18","FAGR_M19","FAGR_M20","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10")
```

## GC community matrix
```{r}
rownames(data.cov.c.agg) <- data.cov.c.agg$gene_cluster_id
data.cov.c.agg <- data.cov.c.agg[,-1]
data.cov.c.agg.comm <- as.data.frame(t(data.cov.c.agg))
data.cov.c.agg.comm <- data.cov.c.agg.comm[,which(colSums(data.cov.c.agg.comm)!=0)] # 22,190 gene clusters with non zero coverage values
sum(data.cov.c.agg.comm) # 7,476,568 total GC counts
rowSums(data.cov.c.agg.comm)
min(rowSums(data.cov.c.agg.comm))
max(rowSums(data.cov.c.agg.comm))
```

-################################################################################

# Rarefying

Rarefying at 44,221
```{r}
data.cov.c.agg.comm.rar <- data.cov.c.agg.comm[which(rowSums(data.cov.c.agg.comm) >= 44221),]
data.cov.c.agg.comm.rar <- as.data.frame(rrarefy(data.cov.c.agg.comm.rar, sample = 44221))
```

Save 'data.cov.c.agg.comm.rar' as RDS file for consistency in rerunning the code.
```{r}
#saveRDS(data.cov.c.agg.comm.rar, "../rds/data.cov.c.agg.comm.rar.rds")
data.cov.c.agg.comm.rar <- readRDS("../rds/data.cov.c.agg.comm.rar.rds")
```

## Adjust metadata table
```{r}
metadata.rar <- metadata[rownames(data.cov.c.agg.comm.rar),]
#saveRDS(metadata.rar, "../rds/metadata.rar.rds")
```

## Verify the effect of rarefying
```{r}
data.cov.c.agg.comm.rar <- data.cov.c.agg.comm.rar[,which(colSums(data.cov.c.agg.comm.rar)!=0)] # 17,488 gene clusters left after rarefaction 
ncol(data.cov.c.agg.comm)-ncol(data.cov.c.agg.comm.rar) # loss of 4702 GC after rarefaction
1-(ncol(data.cov.c.agg.comm.rar)/ncol(data.cov.c.agg.comm)) # loss of 21.19% of GC
# loss of 2 samples (1 ACSA, 1 FAGR)

# rarefied vs. raw GC richness
rowSums(data.cov.c.agg.comm.rar)
richness_raw <- rowSums((data.cov.c.agg.comm[rownames(data.cov.c.agg.comm.rar),]>0)*1)
richness_rar <- rowSums((data.cov.c.agg.comm.rar>0)*1)
plot(richness_rar~richness_raw,xlab='number of unique GC in raw data',ylab='number of unique GC in rarefied data') + abline(0,1)
rm(richness_raw,richness_rar)
```

## Filtering of rarefied data for analyses (Remove low abundant GC)
```{r}
data.cov.c.agg.comm.rar.filt <- data.cov.c.agg.comm.rar[,colSums(data.cov.c.agg.comm.rar) >= 0.00001*sum(colSums(data.cov.c.agg.comm.rar))] # Remove GC that did not make up a total of 0.001% of total GC counts across samples. (i.e. GC that have less than 1 count for every 100,000 counts)
```

```{r}
length(colnames(data.cov.c.agg.comm.rar.filt)) # 10,091 GC
length(colnames(data.cov.c.agg.comm.rar.filt))/length(colnames(data.cov.c.agg.comm.rar)) # Keep 57.7% of rarefied GC
length(colnames(data.cov.c.agg.comm.rar.filt))/length(colnames(data.cov.c.agg.comm)) # Keep 45.5% of all detected GC
sum(data.cov.c.agg.comm.rar.filt)/sum(data.cov.c.agg.comm.rar) # Keep 97.4% of rarefied counts
sum(data.cov.c.agg.comm.rar.filt)/sum(data.cov.c.agg.comm) # Keep 13.3% of total counts
```

## Simplify dataframe name for 'data.cov.c.agg.comm.rar.filt'
```{r}
GC.rar.filt <- data.cov.c.agg.comm.rar.filt
```

Save 'GC.rar.filt' as RDS file for simplicity in recalling the dataframe for DACGs analyses.
```{r}
#saveRDS(GC.rar.filt, "../rds/GC.rar.filt.rds")
GC.rar.filt <- readRDS("../rds/GC.rar.filt.rds")
```

-################################################################################

# Alpha diversity

## Calculate GC richness and shannon index
```{r}
metadata.rar$richness.GC <- rowSums(GC.rar.filt>0)
metadata.rar$shannon.GC <- vegan::diversity(GC.rar.filt, index = "shannon")
```

## Richness
```{r}
lm <- lm(richness.GC ~ host_species, data = metadata.rar)
plot(richness.GC ~ host_species, data = metadata.rar)
summary(lm)
plot(lm)
hist(residuals(lm), breaks = 25)
shapiro.test(residuals(lm))
```

## Shannon index
```{r}
lm <- lm(shannon.GC ~ host_species, data = metadata.rar)
plot(shannon.GC ~ host_species, data = metadata.rar)
summary(lm)
plot(lm)
hist(residuals(lm), breaks = 50)
shapiro.test(residuals(lm))
# residuals are non normal

# Kruskall-Wallis
kruskal.test(shannon.GC ~ host_species, data = metadata.rar)
rstatix::kruskal_effsize(shannon.GC ~ host_species, data = metadata.rar)
```

-################################################################################

# Beta diversity

## Rarefaction only (new rarefaction; no filtering)

```{r}
rowSums(data.cov.c.agg.comm)
```

```{r, eval=FALSE}
GC.dist.rar <- avgdist(data.cov.c.agg.comm, sample = 44221, dmethod = "bray", iteration = 100)
#saveRDS(GC.dist.rar, "../rds/GC.dist.rar.rds")
```

```{r}
GC.dist.rar <- readRDS("../rds/GC.dist.rar.rds")
```

## PERMANOVA
```{r}
GC.perm <- adonis2(GC.dist.rar ~ host_species, data = metadata.rar)
GC.perm
```

## BETADISPER
```{r}
GC.disper <- betadisper(GC.dist.rar, metadata.rar$host_species)
anova(GC.disper)
plot(GC.disper)
```

## Pairwise PERMANOVAs
```{r}
shrub_tree_PERM <- adonis2(GC.dist.rar ~ shrub_tree, data = metadata.rar)
shrub_tree_PERM
shrub_tree_DISP <- betadisper(GC.dist.rar, metadata.rar$shrub_tree)
anova(shrub_tree_DISP)
plot(shrub_tree_DISP)

GC.dist.rar.noCOCO <- data.frame(as.matrix(GC.dist.rar))
GC.dist.rar.noCOCO <- GC.dist.rar.noCOCO[c(1:9,15:23),c(1:9,15:23)]
GC.dist.rar.noCOCO <- as.dist(as.matrix(GC.dist.rar.noCOCO), diag = FALSE, upper = FALSE)
gymno_angio_tree_PERM <- adonis2(GC.dist.rar.noCOCO ~ gymno_angio_tree, data = metadata.rar[c(1:9,15:23),])
gymno_angio_tree_PERM
gymno_angio_tree_DISP <- betadisper(GC.dist.rar.noCOCO, metadata.rar[c(1:9,15:23),]$gymno_angio_tree)
anova(gymno_angio_tree_DISP)
plot(gymno_angio_tree_DISP, label = F)
```

## PCoA

```{r}
GC.pcoa <- cmdscale(GC.dist.rar, k = (nrow(metadata.rar) - 1), eig = TRUE)
pcoa.eig <- eigenvals(GC.pcoa)
pcoa.eig <- abs(pcoa.eig)
sum.pcoa.eig <- sum(pcoa.eig)
pcoa.prop.exp <- pcoa.eig/sum.pcoa.eig 
(pcoa.axis1 <- pcoa.prop.exp[1])
(pcoa.axis2 <- pcoa.prop.exp[2])
site.scores.pcoa <- scores(GC.pcoa) 
sp.scores.pcoa <- wascores(GC.pcoa$points, data.cov.c.agg.comm[c(1:5, 6:7,9:10, 11:15, 16:17,19:20, 21:25),]) 
site.scores.pcoa <- data.frame(site.scores.pcoa)
sp.scores.pcoa <- data.frame(sp.scores.pcoa)
host_species <- metadata.rar$host_species
site.scores.pcoa <- cbind(site.scores.pcoa, host_species)
pcoa.ggordiplot <- gg_ordiplot(GC.pcoa, choices = c(1,2), groups = metadata.rar$host_species, kind = "sd", conf = 0.95, pt.size = 3)
pcoa.ellipses <- pcoa.ggordiplot$df_ellipse # extract confidence ellipse
```

Plot
```{r}
ggplot(data = site.scores.pcoa, 
         aes(x = Dim1, y = Dim2), 
         color = host_species) + 
  expand_limits(x=c(-0.2,0.4), y=c(-0.2,0.3)) +
  geom_point(aes(shape = host_species, color = host_species), show.legend = T, 
             size=3) +
  xlab(paste("PCoA1 (",round(pcoa.axis1*100, digits = 2),"%)", sep = "")) +
  ylab(paste("PCoA2 (",round(pcoa.axis2*100, digits = 2),"%)", sep = "")) +
  #geom_text_repel(data = sp.scores.pcoa, 
  #                aes(x = X1, y = X2), 
  #                label=rownames(sp.scores.pcoa), 
  #                color = "black") +
  scale_shape_manual(values = c(16,16,16,16,16)) +
  guides(shape = "none") +
  geom_path(data = pcoa.ellipses, 
            aes(x = x, y = y, color = Group), 
            show.legend = FALSE) + 
  scale_color_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"), name = "Host species", labels = c("ABBA" = expression(italic("A. balsamea")), "THOC" = expression(italic("T. occidentalis")),"ACSA" = expression(italic("A. saccharum")), "FAGR" = expression(italic("F. grandifolia")), "COCO" = expression(italic("C. cornuta")))) +
  theme_bw() +
  theme(text = element_text(),
      legend.justification = c(0.5, 0),
      legend.text = element_text(size = 11),
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold"),
      plot.margin = margin(10, 10, 10, 10)
      )
ggsave("../figures/figure_3B_gene_clusters_pcoa.pdf", width = 6, height = 4)
```

-################################################################################