---
title: "Genes data preparation"
output: html_document
date: "2024-04-16"
---

-################################################################################

# Librairies
```{r}
library(tidyverse)
library(data.table)
```

-################################################################################

# Import Data

## Metapangenomes

Import individual tree species metapangenomes
```{r}
# ABBA
ABBA_summary <- read.delim("../data/anvio_outputs/Methylobacterium-ABBA-PAN/Methylobacterium-ABBA-METAPAN-SUMMARY/Methylobacterium-ABBA-PAN_gene_clusters_summary.txt.gz", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
ABBA_summary <- ABBA_summary[,-1]
# ACSA
ACSA_summary <- read.delim("../data/anvio_outputs/Methylobacterium-ACSA-PAN/Methylobacterium-ACSA-METAPAN-SUMMARY/Methylobacterium-ACSA-PAN_gene_clusters_summary.txt.gz", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
ACSA_summary <- ACSA_summary[,-1]
# COCO
COCO_summary <- read.delim("../data/anvio_outputs/Methylobacterium-COCO-PAN/Methylobacterium-COCO-METAPAN-SUMMARY/Methylobacterium-COCO-PAN_gene_clusters_summary.txt.gz", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
COCO_summary <- COCO_summary[,-1]
# FAGR
FAGR_summary <- read.delim("../data/anvio_outputs/Methylobacterium-FAGR-PAN/Methylobacterium-FAGR-METAPAN-SUMMARY/Methylobacterium-FAGR-PAN_gene_clusters_summary.txt.gz", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
FAGR_summary <- FAGR_summary[,-1]
# THOC
THOC_summary <- read.delim("../data/anvio_outputs/Methylobacterium-THOC-PAN/Methylobacterium-THOC-METAPAN-SUMMARY/Methylobacterium-THOC-PAN_gene_clusters_summary.txt.gz", header = TRUE, sep = "\t", dec = ".", quote="", fill=FALSE)
THOC_summary <- THOC_summary[,-1]
```

Since these 5 metapangenomes files are similar, we will work with only one from nom on, but merge gene type (core vs. accessory), gene coverage depth, and gene horizontal coverage data for all samples.

Remove unnecessary metagenomes files
```{r}
rm(ACSA_summary, COCO_summary, FAGR_summary, THOC_summary)
```

Edit ABBA_summary
```{r}
data <- ABBA_summary
# Remove unwanted columns
data <- subset(data, select = -c(2,12:15))
# Change vectors to factors
data <- data %>% mutate(across(c(gene_cluster_id, genome_name, gene_callers_id, KEGG_Class_ACC, KEGG_Class, KEGG_BRITE_ACC, KEGG_BRITE, KEGG_Module_ACC, KEGG_Module, KOfam_ACC, KOfam, Pfam_ACC, Pfam, CAZyme_ACC, CAZyme, COG20_FUNCTION_ACC, COG20_FUNCTION, COG20_PATHWAY_ACC, COG20_PATHWAY, COG20_CATEGORY_ACC, COG20_CATEGORY), as.factor))
# Change vectors to numeric
data <- data %>% mutate(across(c(num_genomes_gene_cluster_has_hits, num_genes_in_gene_cluster, max_num_paralogs, SCG), as.numeric))
```

## mOTUpan data
```{r}
mOTUpan.import <- read.delim("../data/anvio_outputs/core_pan_genome_table", header = TRUE, sep = "\t", dec = ".", quote="")
colnames(mOTUpan.import)[1] <- "gene_cluster_id"
mOTUpan <- mOTUpan.import[,-7]
mOTUpan <- mOTUpan %>% mutate(across(c(gene_cluster_id, type), as.factor))
```

## Gene 'detection' (horizontal coverage)

Create new factorial vector with 104 genome names
```{r}
genomes <- unique(data$genome_name)
```

Import gene detection data
```{r}
# ABBA
ABBA.txtFileName <- paste("../data/anvio_outputs/ABBA/ABBA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
ABBA.gene.detect <- rbindlist(sapply(ABBA.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ABBA.gene.detect$gene_callers_id <- as.factor(ABBA.gene.detect$gene_callers_id)

# ACSA
ACSA.txtFileName <- paste("../data/anvio_outputs/ACSA/ACSA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
ACSA.gene.detect <- rbindlist(sapply(ACSA.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ACSA.gene.detect$gene_callers_id <- as.factor(ACSA.gene.detect$gene_callers_id)

# COCO
COCO.txtFileName <- paste("../data/anvio_outputs/COCO/COCO_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
COCO.gene.detect <- rbindlist(sapply(COCO.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
COCO.gene.detect$gene_callers_id <- as.factor(COCO.gene.detect$gene_callers_id)

# FAGR
FAGR.txtFileName <- paste("../data/anvio_outputs/FAGR/FAGR_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
FAGR.gene.detect <- rbindlist(sapply(FAGR.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
FAGR.gene.detect$gene_callers_id <- as.factor(FAGR.gene.detect$gene_callers_id)

# THOC
THOC.txtFileName <- paste("../data/anvio_outputs/THOC/THOC_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_detection.txt",sep="")
THOC.gene.detect <- rbindlist(sapply(THOC.txtFileName, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
THOC.gene.detect$gene_callers_id <- as.factor(THOC.gene.detect$gene_callers_id)

rm(ABBA.txtFileName,ACSA.txtFileName,COCO.txtFileName,FAGR.txtFileName,THOC.txtFileName)
```

Create gene.detection dataframe
```{r}
ABBA.gene.detect <- ABBA.gene.detect[,-1]
ACSA.gene.detect <- ACSA.gene.detect[,-1]
COCO.gene.detect <- COCO.gene.detect[,-1]
FAGR.gene.detect <- FAGR.gene.detect[,-1]
THOC.gene.detect <- THOC.gene.detect[,-1]
gene.detection <- merge(ABBA.gene.detect,ACSA.gene.detect, by= "gene_callers_id")
gene.detection <- merge(gene.detection,COCO.gene.detect, by= "gene_callers_id")
gene.detection <- merge(gene.detection,FAGR.gene.detect, by= "gene_callers_id")
gene.detection <- merge(gene.detection,THOC.gene.detect, by= "gene_callers_id")
rm(ABBA.gene.detect,ACSA.gene.detect,COCO.gene.detect,FAGR.gene.detect,THOC.gene.detect)
```

## Gene coverage (coverage depth)

Import gene coverage data
```{r}
# ABBA
ABBA.txtFileName2 <- paste("../data/anvio_outputs/ABBA/ABBA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_coverages.txt",sep="")
ABBA.gene.cov <- rbindlist(sapply(ABBA.txtFileName2, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ABBA.gene.cov$gene_callers_id <- as.factor(ABBA.gene.cov$gene_callers_id)

# ACSA
ACSA.txtFileName2 <- paste("../data/anvio_outputs/ACSA/ACSA_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_coverages.txt",sep="")
ACSA.gene.cov <- rbindlist(sapply(ACSA.txtFileName2, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
ACSA.gene.cov$gene_callers_id <- as.factor(ACSA.gene.cov$gene_callers_id)

# COCO
COCO.txtFileName2 <- paste("../data/anvio_outputs/COCO/COCO_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_coverages.txt",sep="")
COCO.gene.cov <- rbindlist(sapply(COCO.txtFileName2, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
COCO.gene.cov$gene_callers_id <- as.factor(COCO.gene.cov$gene_callers_id)

# FAGR
FAGR.txtFileName2 <- paste("../data/anvio_outputs/FAGR/FAGR_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_coverages.txt",sep="")
FAGR.gene.cov <- rbindlist(sapply(FAGR.txtFileName2, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
FAGR.gene.cov$gene_callers_id <- as.factor(FAGR.gene.cov$gene_callers_id)

# THOC
THOC.txtFileName2 <- paste("../data/anvio_outputs/THOC/THOC_PROFILE-SUMMARY/bin_by_bin/",genomes,"/",genomes,"-gene_coverages.txt",sep="")
THOC.gene.cov <- rbindlist(sapply(THOC.txtFileName2, fread, simplify = FALSE),
                use.names = TRUE, idcol = "FileName")
THOC.gene.cov$gene_callers_id <- as.factor(THOC.gene.cov$gene_callers_id)

rm(ABBA.txtFileName2,ACSA.txtFileName2,COCO.txtFileName2,FAGR.txtFileName2,THOC.txtFileName2)
```

Create gene.coverage dataframe
```{r}
ABBA.gene.cov <- ABBA.gene.cov[,-1]
ACSA.gene.cov <- ACSA.gene.cov[,-1]
COCO.gene.cov <- COCO.gene.cov[,-1]
FAGR.gene.cov <- FAGR.gene.cov[,-1]
THOC.gene.cov <- THOC.gene.cov[,-1]
gene.coverage <- merge(ABBA.gene.cov,ACSA.gene.cov, by= "gene_callers_id")
gene.coverage <- merge(gene.coverage,COCO.gene.cov, by= "gene_callers_id")
gene.coverage <- merge(gene.coverage,FAGR.gene.cov, by= "gene_callers_id")
gene.coverage <- merge(gene.coverage,THOC.gene.cov, by= "gene_callers_id")
rm(ABBA.gene.cov,ACSA.gene.cov,COCO.gene.cov,FAGR.gene.cov,THOC.gene.cov)
```

-################################################################################

# Edit dataframes and create RDS files for subsequent analyses

## Merge 'mOTUpan' with 'data'
```{r}
data <- merge(data, mOTUpan, by = "gene_cluster_id")
data.pangenome <- data[c(1:3,30:34,4:10)]
```

## Create data.pangenome RDS file
```{r}
#saveRDS(data.pangenome, "../rds/data.pangenome.rds")
data.pangenome <- readRDS("../rds/data.pangenome.rds")
```

## Merge 'gene.detection' with 'data'
```{r}
data.fct.det <- merge(data, gene.detection, by = "gene_callers_id")
data.fct.det <- data.fct.det[order(match(data.fct.det$gene_callers_id, data$gene_callers_id)), ]
rownames(data.fct.det) <- NULL
data.fct.det <- data.fct.det[,c(2,1,3,30,31,17,18,11,12,19,20,25,26,23,24,27,28,13,14,21,22,15,16,29,35:59)]
```

## Create data.fct.det RDS file
```{r}
#saveRDS(data.fct.det, "../rds/data.fct.det.rds")
data.fct.det <- readRDS("../rds/data.fct.det.rds")
```

## Merge 'gene.coverage' with 'data'
```{r}
data.fct.cov <- merge(data, gene.coverage, by = "gene_callers_id")
data.fct.cov <- data.fct.cov[order(match(data.fct.cov$gene_callers_id, data$gene_callers_id)), ]
rownames(data.fct.cov) <- NULL
data.fct.cov <- data.fct.cov[,c(2,1,3,30,31,17,18,11,12,19,20,25,26,23,24,27,28,13,14,21,22,15,16,29,35:59)]
```

## Create data.fct.cov RDS file
```{r}
#saveRDS(data.fct.cov, "../rds/data.fct.cov.rds")
data.fct.cov <- readRDS("../rds/data.fct.cov.rds")
```

-################################################################################

# Descriptive statistics : number of individual genes and functional annotations from 3 databases
```{r}
length(unique(data.pangenome$gene_callers_id))
sum(data.fct.cov$COG20_FUNCTION_ACC!="")
sum(data.fct.cov$KOfam_ACC!="")
sum(data.fct.cov$Pfam_ACC!="")
```

-################################################################################