---
title: "Genes diversity"
output: html_document
date: "2024-10-29"
---

-################################################################################

# Libraries
```{r}
library(tidyverse)
library(vegan)
library(picante)
library(ggordiplots)
#library(gridExtra)
library(cowplot)
library(viridis)
library(DESeq2)
#library(patchwork)
#library(EnhancedVolcano)
```

# Import data

## Horizontal coverage ('detection')
```{r}
data.fct.det <- readRDS("../rds/data.fct.det.rds")
rownames(data.fct.det) <- data.fct.det$gene_callers_id
```

## Coverage depth
```{r}
data.fct.cov <- readRDS("../rds/data.fct.cov.rds")
rownames(data.fct.cov) <- data.fct.cov$gene_callers_id
```

## Metadata
```{r}
metadata <- read.delim("../data/samples_metadata.txt", header = TRUE, sep = "\t", dec = ".", row.names = 1)
metadata <- metadata %>% arrange(row.names(.))
metadata <- metadata %>% mutate(across(c(plant_clade,plant_order,plant_family,host_species,tree), as.factor))
metadata$host_species <- factor(x = metadata$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
metadata$shrub_tree <- c("tree","tree","tree","tree","tree","tree","tree","tree","tree","tree","shrub","shrub","shrub","shrub","shrub","tree","tree","tree","tree","tree","tree","tree","tree","tree","tree")
metadata$shrub_tree <- factor(x = metadata$shrub_tree, levels = c("shrub","tree"))
metadata$gymno_angio_tree <- c("gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree")
metadata$gymno_angio_tree <- factor(x = metadata$gymno_angio_tree, levels = c("gymnosperm_tree","angiosperm_tree","angiosperm_shrub"))
```

-################################################################################

# Prepare data

Sanity check
```{r}
all(rownames(data.fct.cov)==rownames(data.fct.det))
all(colnames(data.fct.cov)==colnames(data.fct.det))
```

## Detection

For analyses, keep only genes that are horizontally covered at least 1X for 50% of nucleotides. Those genes are considered 'present'.

If det<0.5 = 0, if det>=0.5 = 1
```{r}
data.det.PA <- data.frame(ifelse(data.fct.det[,c(25:49)] < 0.5, 0, 1))
```

## Coverage

For analyses, keep coverage values only for genes that are 'present', then round the coverage values to integers (counts of genes).

If det=0, cov=0; if det=1, cov=cov
```{r}
data.cov.pre <- data.fct.cov[,c(25:49)]
data.cov <- data.frame(mapply(function(x, y) ifelse(x == 1, y, 0), data.det.PA, data.cov.pre))
rownames(data.cov) <- rownames(data.fct.cov)
```

## Round
```{r}
data.cov.c <- round(data.cov, digits=0)
```

## Merge coverage data
Merge coverage counts dataframe with functions
```{r}
all(rownames(data.fct.cov)==rownames(data.cov.c))
data.fct.cov.c <- cbind(data.fct.cov[,c(1:24)], data.cov.c)
```

## Aggregate gene counts by GC

Aggregate (sum) count values by GC
```{r}
data.fct.cov.c %>% distinct(gene_cluster_id) # There is 70,695 gene clusters in Methylobacterium pangenome
data.cov.c.agg <- aggregate(cbind(data.fct.cov.c$ABBA_M11,data.fct.cov.c$ABBA_M12,data.fct.cov.c$ABBA_M13,data.fct.cov.c$ABBA_M14,data.fct.cov.c$ABBA_M15,data.fct.cov.c$ACSA_M21,data.fct.cov.c$ACSA_M22,data.fct.cov.c$ACSA_M23,data.fct.cov.c$ACSA_M24,data.fct.cov.c$ACSA_M25,data.fct.cov.c$COCO_M01,data.fct.cov.c$COCO_M02,data.fct.cov.c$COCO_M03,data.fct.cov.c$COCO_M04,data.fct.cov.c$COCO_M05,data.fct.cov.c$FAGR_M16,data.fct.cov.c$FAGR_M17,data.fct.cov.c$FAGR_M18,data.fct.cov.c$FAGR_M19,data.fct.cov.c$FAGR_M20,data.fct.cov.c$THOC_M06,data.fct.cov.c$THOC_M07,data.fct.cov.c$THOC_M08,data.fct.cov.c$THOC_M09,data.fct.cov.c$THOC_M10),  by = list(data.fct.cov.c$gene_cluster_id), sum)
colnames(data.cov.c.agg) <- c("gene_cluster_id","ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","ACSA_M21","ACSA_M22","ACSA_M23","ACSA_M24","ACSA_M25","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05","FAGR_M16","FAGR_M17","FAGR_M18","FAGR_M19","FAGR_M20","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10")
```

## GC community matrix
```{r}
rownames(data.cov.c.agg) <- data.cov.c.agg$gene_cluster_id
data.cov.c.agg <- data.cov.c.agg[,-1]
data.cov.c.agg.comm <- as.data.frame(t(data.cov.c.agg))
data.cov.c.agg.comm <- data.cov.c.agg.comm[,which(colSums(data.cov.c.agg.comm)!=0)] # 22,190 gene clusters with non zero coverage values
sum(data.cov.c.agg.comm) # 7,476,568 total GC counts
rowSums(data.cov.c.agg.comm)
min(rowSums(data.cov.c.agg.comm))
max(rowSums(data.cov.c.agg.comm))
```

-################################################################################

# Rarefying

Rarefying at 44,221
```{r}
data.cov.c.agg.comm.rar <- data.cov.c.agg.comm[which(rowSums(data.cov.c.agg.comm) >= 44221),]
data.cov.c.agg.comm.rar <- as.data.frame(rrarefy(data.cov.c.agg.comm.rar, sample = 44221))
```

Save 'data.cov.c.agg.comm.rar' as RDS file for consistency in rerunning the code.
```{r}
#saveRDS(data.cov.c.agg.comm.rar, "../rds/data.cov.c.agg.comm.rar.rds")
data.cov.c.agg.comm.rar <- readRDS("../rds/data.cov.c.agg.comm.rar.rds")
```

## Adjust metadata table
```{r}
metadata.rar <- metadata[rownames(data.cov.c.agg.comm.rar),]
```

## Verify the effect of rarefying
```{r}
data.cov.c.agg.comm.rar <- data.cov.c.agg.comm.rar[,which(colSums(data.cov.c.agg.comm.rar)!=0)] # 17,488 gene clusters left after rarefaction 
ncol(data.cov.c.agg.comm)-ncol(data.cov.c.agg.comm.rar) # loss of 4702 GC after rarefaction
1-(ncol(data.cov.c.agg.comm.rar)/ncol(data.cov.c.agg.comm)) # loss of 21.19% of GC
# loss of 2 samples (1 ACSA, 1 FAGR)

# rarefied vs. raw GC richness
rowSums(data.cov.c.agg.comm.rar)
richness_raw <- rowSums((data.cov.c.agg.comm[rownames(data.cov.c.agg.comm.rar),]>0)*1)
richness_rar <- rowSums((data.cov.c.agg.comm.rar>0)*1)
plot(richness_rar~richness_raw,xlab='number of unique GC in raw data',ylab='number of unique GC in rarefied data') + abline(0,1)
rm(richness_raw,richness_rar)
```

## Filtering of rarefied data for analyses (Remove low abundant GC)
```{r}
data.cov.c.agg.comm.rar.filt <- data.cov.c.agg.comm.rar[,colSums(data.cov.c.agg.comm.rar) >= 0.00001*sum(colSums(data.cov.c.agg.comm.rar))] # Remove GC that did not make up a total of 0.001% of total GC counts across samples. (i.e. GC that have less than 1 count for every 100,000 counts)
```

```{r}
length(colnames(data.cov.c.agg.comm.rar.filt)) # 10,091 GC
length(colnames(data.cov.c.agg.comm.rar.filt))/length(colnames(data.cov.c.agg.comm.rar)) # Keep 57.7% of rarefied GC
length(colnames(data.cov.c.agg.comm.rar.filt))/length(colnames(data.cov.c.agg.comm)) # Keep 45.5% of all detected GC
sum(data.cov.c.agg.comm.rar.filt)/sum(data.cov.c.agg.comm.rar) # Keep 97.4% of rarefied counts
sum(data.cov.c.agg.comm.rar.filt)/sum(data.cov.c.agg.comm) # Keep 13.3% of total counts
```

## Simplify dataframe name for 'data.cov.c.agg.comm.rar.filt'
```{r}
GC.rar.filt <- data.cov.c.agg.comm.rar.filt
```

Save 'GC.rar.filt' as RDS file for simplicity in recalling the dataframe for DACGs analyses.
```{r}
#saveRDS(GC.rar.filt, "../rds/GC.rar.filt.rds")
GC.rar.filt <- readRDS("../rds/GC.rar.filt.rds")
```


-################################################################################

# Alpha diversity

## Calculate GC richness and shannon index
```{r}
metadata.rar$richness.GC <- rowSums(GC.rar.filt>0)
metadata.rar$shannon.GC <- vegan::diversity(GC.rar.filt, index = "shannon")
```

## Richness
```{r}
lm <- lm(richness.GC ~ host_species, data = metadata.rar)
plot(richness.GC ~ host_species, data = metadata.rar)
summary(lm)
plot(lm)
hist(residuals(lm), breaks = 25)
shapiro.test(residuals(lm))
```

## Shannon index
```{r}
lm <- lm(shannon.GC ~ host_species, data = metadata.rar)
plot(shannon.GC ~ host_species, data = metadata.rar)
summary(lm)
plot(lm)
hist(residuals(lm), breaks = 50)
shapiro.test(residuals(lm))
# residuals are non normal

# Kruskall-Wallis
kruskal.test(shannon.GC ~ host_species, data = metadata.rar)
rstatix::kruskal_effsize(shannon.GC ~ host_species, data = metadata.rar)
```

-################################################################################

# Beta diversity

## Rarefaction only (new rarefaction; no filtering)

```{r}
rowSums(data.cov.c.agg.comm)
```

```{r, eval=FALSE}
GC.dist.rar <- avgdist(data.cov.c.agg.comm, sample = 44221, dmethod = "bray", iteration = 100)
#saveRDS(GC.dist.rar, "../rds/GC.dist.rar.rds")
```

```{r}
GC.dist.rar <- readRDS("../rds/GC.dist.rar.rds")
```

## PERMANOVA
```{r}
GC.perm <- adonis2(GC.dist.rar ~ host_species, data = metadata.rar)
GC.perm
```

## BETADISPER
```{r}
GC.disper <- betadisper(GC.dist.rar, metadata.rar$host_species)
anova(GC.disper)
plot(GC.disper)
```

## Pairwise PERMANOVAs
```{r}
shrub_tree_PERM <- adonis2(GC.dist.rar ~ shrub_tree, data = metadata.rar)
shrub_tree_PERM
shrub_tree_DISP <- betadisper(GC.dist.rar, metadata.rar$shrub_tree)
anova(shrub_tree_DISP)
plot(shrub_tree_DISP)

GC.dist.rar.noCOCO <- data.frame(as.matrix(GC.dist.rar))
GC.dist.rar.noCOCO <- GC.dist.rar.noCOCO[c(1:9,15:23),c(1:9,15:23)]
GC.dist.rar.noCOCO <- as.dist(as.matrix(GC.dist.rar.noCOCO), diag = FALSE, upper = FALSE)
gymno_angio_tree_PERM <- adonis2(GC.dist.rar.noCOCO ~ gymno_angio_tree, data = metadata.rar[c(1:9,15:23),])
gymno_angio_tree_PERM
gymno_angio_tree_DISP <- betadisper(GC.dist.rar.noCOCO, metadata.rar[c(1:9,15:23),]$gymno_angio_tree)
anova(gymno_angio_tree_DISP)
plot(gymno_angio_tree_DISP, label = F)
```

## PCoA

```{r}
GC.pcoa <- cmdscale(GC.dist.rar, k = (nrow(metadata.rar) - 1), eig = TRUE)
pcoa.eig <- eigenvals(GC.pcoa)
pcoa.eig <- abs(pcoa.eig)
sum.pcoa.eig <- sum(pcoa.eig)
pcoa.prop.exp <- pcoa.eig/sum.pcoa.eig 
(pcoa.axis1 <- pcoa.prop.exp[1])
(pcoa.axis2 <- pcoa.prop.exp[2])
site.scores.pcoa <- scores(GC.pcoa) 
sp.scores.pcoa <- wascores(GC.pcoa$points, data.cov.c.agg.comm[c(1:5, 6:7,9:10, 11:15, 16:17,19:20, 21:25),]) 
site.scores.pcoa <- data.frame(site.scores.pcoa)
sp.scores.pcoa <- data.frame(sp.scores.pcoa)
host_species <- metadata.rar$host_species
site.scores.pcoa <- cbind(site.scores.pcoa, host_species)
pcoa.ggordiplot <- gg_ordiplot(GC.pcoa, choices = c(1,2), groups = metadata.rar$host_species, kind = "sd", conf = 0.95, pt.size = 3)
pcoa.ellipses <- pcoa.ggordiplot$df_ellipse # extract confidence ellipse
```

Plot
```{r}
ggplot(data = site.scores.pcoa, 
         aes(x = Dim1, y = Dim2), 
         color = host_species) + 
  expand_limits(x=c(-0.2,0.4), y=c(-0.2,0.3)) +
  geom_point(aes(shape = host_species, color = host_species), show.legend = T, 
             size=3) +
  xlab(paste("PCoA1 (",round(pcoa.axis1*100, digits = 2),"%)", sep = "")) +
  ylab(paste("PCoA2 (",round(pcoa.axis2*100, digits = 2),"%)", sep = "")) +
  #geom_text_repel(data = sp.scores.pcoa, 
  #                aes(x = X1, y = X2), 
  #                label=rownames(sp.scores.pcoa), 
  #                color = "black") +
  scale_shape_manual(values = c(16,16,16,16,16)) +
  guides(shape = "none") +
  geom_path(data = pcoa.ellipses, 
            aes(x = x, y = y, color = Group), 
            show.legend = FALSE) + 
  scale_color_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"), name = "Host species", labels = c("ABBA" = expression(italic("A. balsamea")), "THOC" = expression(italic("T. occidentalis")),"ACSA" = expression(italic("A. saccharum")), "FAGR" = expression(italic("F. grandifolia")), "COCO" = expression(italic("C. cornuta")))) +
  theme_bw() +
  theme(text = element_text(),
      legend.justification = c(0.5, 0),
      legend.text = element_text(size = 11),
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold"),
      plot.margin = margin(10, 10, 10, 10)
      )
ggsave("../figures/figure_3B_gene_clusters_pcoa.pdf", width = 6, height = 4)
```

-################################################################################

# DACGs analyses 

```{r}
options(scipen = 999)
```




RENDU LÀ!!!! 260714 22h49



## Prepare data for DESeq

Rarefied and filtered data
Use data after 'rarefying' at 44,221 and after filtering low abundant GC
```{r}
countData <- as.data.frame(t(GC.rar.filt))
sum(rowSums(countData)==0)
colSums(countData)
```

Make sure count dataframe's columns and metadata's rows are in the same order
```{r}
all(rownames(metadata.rar) == colnames(countData))
```

Prepare dataframe containing rarefied and filtered GC and their (rarefied) coverage values, and functional annotations.
```{r}
#GC.rar.filt <- readRDS("../rds/GC.rar.filt.rds")

# Edit rarefied and filtered GC dataframe
GC.rar.filt.t <- as.data.frame(t(GC.rar.filt))
GC.rar.filt.t$gene_cluster_id <- rownames(GC.rar.filt.t)
row.names(GC.rar.filt.t) <- NULL

# Edit 'data.fct.cov.c' to remove coverage values
data.fct <- data.fct.cov.c[,c(1:41)]

# Merge dataframes to have GC that were kept after rarefaction and filtering, with their coverage values (rarefied values)
GC.rar.fct.cov <- merge(data.fct, GC.rar.filt.t, by = "gene_cluster_id")
GC.rar.fct.cov <- GC.rar.fct.cov[order(GC.rar.fct.cov$gene_cluster_id, GC.rar.fct.cov$genome_name), ]
GC.rar.fct.cov <- droplevels(GC.rar.fct.cov)

#saveRDS(GC.rar.fct.cov, "../rds/GC.rar.fct.cov.rds")
GC.rar.fct.cov <- readRDS("../rds/GC.rar.fct.cov.rds")
```

## Trees & Shrub

### Analysis
```{r}
GC.dds.2 <- DESeqDataSetFromMatrix(countData = countData,
                              colData = metadata.rar, 
                              design = ~ shrub_tree)
GC.dds.2$host_species <- relevel(GC.dds.2$shrub_tree, ref = "shrub")
GC.dds.2 <- DESeq(GC.dds.2)
resultsNames(GC.dds.2) # lists the coefficients
```

Save and load GC.dds as RDS file
```{r}
#saveRDS(GC.dds.2, "../rds/GC.dds.2_v5.rds")
GC.dds.2 <- readRDS("../rds/GC.dds.2_v5.rds")
```

### Exploration
```{r}
GC.res.tree.shrub <- results(GC.dds.2, contrast=c("shrub_tree","tree","shrub"), alpha = 0.05)
# test with shrinkage
#GC.res.tree.shrub <- lfcShrink(GC.dds.2, coef="shrub_tree_tree_vs_shrub", type="apeglm")
# test with shrinkage and lfc hypothesis
#GC.res.tree.shrub <- lfcShrink(GC.dds.2, coef="shrub_tree_tree_vs_shrub", type="apeglm")
GC.sig.tree.shrub <- GC.res.tree.shrub[is.na(GC.res.tree.shrub$padj)==F&GC.res.tree.shrub$padj<0.05,] # 271 DAGs
```
### Volcano plot

Plot
```{r}
keyvals <- ifelse(
    GC.res.tree.shrub$padj >= 0.05, 'lightgray',
      ifelse(GC.res.tree.shrub$padj < 0.01 & abs(GC.res.tree.shrub$log2FoldChange) >= 2, 'limegreen',
        'black'))
keyvals[is.na(keyvals)] <- 'black'
names(keyvals)[keyvals == 'lightgray'] <- 'n.s. (p ≥ 0.05)'
names(keyvals)[keyvals == 'black'] <- 'p < 0.05'
names(keyvals)[keyvals == 'limegreen'] <- 'p < 0.01 & log(2)FC ≥ 2 '
#keyvals <- factor(x = keyvals, levels = c('lightgray','black','limegreen'))
#levels(keyvals) <- c('lightgray','black','limegreen')

#keyvals <- keyvals[!is.na(keyvals)]

volcano.TS <-  EnhancedVolcano(GC.res.tree.shrub,
    lab = rownames(GC.res.tree.shrub),
    x = 'log2FoldChange',
    y = 'padj',
    title = "",
    pCutoff = 0.01,
    FCcutoff = 2,
    pointSize = 1.5,
    labSize = 6.0,
    selectLab = "",
    subtitle = "",
    caption = "",
    shape = 16,
    colCustom = keyvals,
    colAlpha = 1,
    cutoffLineCol = 'forestgreen',
    cutoffLineWidth = 0.5,
    hline = 0.05,
    hlineCol = "red",
    hlineType = "dashed",
    hlineWidth = 0.6,
    #vline = c(-1,1),
    #vlineCol = c("firebrick","firebrick"),
    #vlineType = c("dashed","dashed"),
    #vlineWidth = c(0.5,0.5),
    legendPosition = 'right',
    legendLabSize = 13,
    legendIconSize = 4.0,
    drawConnectors = F,
    #widthConnectors = 1.0,
    #colConnectors = 'black',
    #arrowheads = FALSE,
    gridlines.major = TRUE,
    gridlines.minor = FALSE
    #border = 'partial',
    #borderWidth = 1,
    #borderColour = 'black'
    ) +
    coord_cartesian(xlim=c(-5, 5)) +
    scale_x_continuous(breaks=seq(-5,5, 1)) + 
    scale_y_continuous(breaks=seq(0, 9)) + 
    theme_bw() +
    theme(text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_blank(),
      legend.text = element_text(size = 11),
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
volcano.TS

ggsave("../figures/volcano.TS.tiff", width = 6, height = 4)
```

Data
```{r}
volc.TS.data <- volcano.TS[["data"]][volcano.TS [["data"]]$Sig == "FC_P",]
volc.TS.data$gene_cluster_id <- rownames(volc.TS.data)
volc.TS.data <- volc.TS.data[,c(11,1:10)]
rownames(volc.TS.data) <- NULL

#saveRDS(volc.TS.data, "../rds/volc.TS.data.rds")
volc.TS.data <- readRDS("../rds/volc.TS.data.rds")

DAGs.functions.all.info <- read_delim("../DAGs.functions.all.info.csv")
DAGs.functions.all.info <- DAGs.functions.all.info[,c(5,1:4,6:13)]

volc.TS.data.fct <- merge(volc.TS.data, DAGs.functions.all.info[,c(2:4,10:13)], by = "gene_cluster_id")
volc.TS.data.fct <- volc.TS.data.fct[!grepl("^GC_\\d{8}$", volc.TS.data.fct$GC.function), ]
volc.TS.data.fct <- volc.TS.data.fct[order(-volc.TS.data.fct$log2FoldChange), ]
rownames(volc.TS.data.fct) <- NULL

#saveRDS(volc.TS.data.fct, "../rds/volc.TS.data.fct.rds")
volc.TS.data.fct <- readRDS("../rds/volc.TS.data.fct.rds")
```


### Trees vs COCO
```{r}
GC.sig.Trees.COCO <- as.data.frame(GC.res.tree.shrub[is.na(GC.res.tree.shrub$padj)==F&GC.res.tree.shrub$padj<0.05&GC.res.tree.shrub$log2FoldChange>0,]) # 231 DAGs
```

```{r}
# Arrange the results dataframe in decreasing log2FoldChange
GC.sig.Trees.COCO$gene_cluster_id <- as.factor(rownames(GC.sig.Trees.COCO))
GC.sig.Trees.COCO <- GC.sig.Trees.COCO[order(-GC.sig.Trees.COCO$log2FoldChange), ]
row.names(GC.sig.Trees.COCO) <- NULL

DAG.Trees.COCO <- GC.sig.Trees.COCO

# Add comparisons
DAG.Trees.COCO$ABBA <- rep(1,nrow(DAG.Trees.COCO))
DAG.Trees.COCO$THOC <- rep(1,nrow(DAG.Trees.COCO))
DAG.Trees.COCO$ACSA <- rep(1,nrow(DAG.Trees.COCO))
DAG.Trees.COCO$FAGR <- rep(1,nrow(DAG.Trees.COCO))
DAG.Trees.COCO$COCO <- rep(-1,nrow(DAG.Trees.COCO))

# Merge DAGs with function and coverage df
DAG.Trees.COCO.fct <- merge(GC.rar.fct.cov, DAG.Trees.COCO, by = "gene_cluster_id")

# Keep rows with unique functional annotation
DAG.Trees.COCO.fct.distinct <- DAG.Trees.COCO.fct %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, CAZyme, Description, .keep_all = TRUE)
DAG.Trees.COCO.fct.distinct <- DAG.Trees.COCO.fct.distinct[,-c(2:3)] # remove gene caller id and genome name because irrelevant
DAG.Trees.COCO.fct.distinct <- DAG.Trees.COCO.fct.distinct[c(1:3,63:64,4:62,65:73)] # rearrange column order
DAG.Trees.COCO.fct.distinct <- DAG.Trees.COCO.fct.distinct[order(-DAG.Trees.COCO.fct.distinct$log2FoldChange), ] # place rows in order of decreasing mean_log2FC
row.names(DAG.Trees.COCO.fct.distinct) <- NULL
DAG.Trees.COCO.fct.distinct <- droplevels(DAG.Trees.COCO.fct.distinct)
str(DAG.Trees.COCO.fct.distinct)
```

### COCO vs Trees
```{r}
GC.res.shrub.tree <- results(GC.dds.2, contrast=c("shrub_tree","shrub","tree"), alpha = 0.05)
GC.sig.COCO.Trees <- as.data.frame(GC.res.shrub.tree[is.na(GC.res.shrub.tree$padj)==F&GC.res.shrub.tree$padj<0.05&GC.res.shrub.tree$log2FoldChange>0,]) # 40 DAGs
```

```{r}
# Arrange the results dataframe in decreasing log2FoldChange
GC.sig.COCO.Trees$gene_cluster_id <- as.factor(rownames(GC.sig.COCO.Trees))
GC.sig.COCO.Trees <- GC.sig.COCO.Trees[order(-GC.sig.COCO.Trees$log2FoldChange), ]
row.names(GC.sig.COCO.Trees) <- NULL

DAG.COCO.Trees <- GC.sig.COCO.Trees

# Add comparisons
DAG.COCO.Trees$ABBA <- rep(-1,nrow(DAG.COCO.Trees))
DAG.COCO.Trees$THOC <- rep(-1,nrow(DAG.COCO.Trees))
DAG.COCO.Trees$ACSA <- rep(-1,nrow(DAG.COCO.Trees))
DAG.COCO.Trees$FAGR <- rep(-1,nrow(DAG.COCO.Trees))
DAG.COCO.Trees$COCO <- rep(1,nrow(DAG.COCO.Trees))

# Merge DAGs with function and coverage df
DAG.COCO.Trees.fct <- merge(GC.rar.fct.cov, DAG.COCO.Trees, by = "gene_cluster_id")

# Keep rows with unique functional annotation
DAG.COCO.Trees.fct.distinct <- DAG.COCO.Trees.fct %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, CAZyme, Description, .keep_all = TRUE)
DAG.COCO.Trees.fct.distinct <- DAG.COCO.Trees.fct.distinct[,-c(2:3)] # remove gene caller id and genome name because irrelevant
DAG.COCO.Trees.fct.distinct <- DAG.COCO.Trees.fct.distinct[c(1:3,63:64,4:62,65:73)] # rearrange column order
DAG.COCO.Trees.fct.distinct <- DAG.COCO.Trees.fct.distinct[order(-DAG.COCO.Trees.fct.distinct$log2FoldChange), ] # place rows in order of decreasing mean_log2FC
row.names(DAG.COCO.Trees.fct.distinct) <- NULL
DAG.COCO.Trees.fct.distinct <- droplevels(DAG.COCO.Trees.fct.distinct)
str(DAG.COCO.Trees.fct.distinct)
```


## Angio vs Gymno

### Analysis
```{r}
GC.dds.3 <- DESeqDataSetFromMatrix(countData = countData,
                              colData = metadata.rar, 
                              design = ~ plant_clade)
GC.dds.3$plant_clade <- relevel(GC.dds.3$plant_clade, ref = "Gymnospermae")
GC.dds.3 <- DESeq(GC.dds.3)
resultsNames(GC.dds.3) # lists the coefficients
```

Save and load GC.dds as RDS file
```{r}
#saveRDS(GC.dds.3, "../rds/GC.dds.3_v5.rds")
GC.dds.3 <- readRDS("../rds/GC.dds.3_v5.rds")
```

### Exploration
```{r}
GC.res.angio.gymno <- results(GC.dds.3, contrast=c("plant_clade","Angiospermae","Gymnospermae"), alpha = 0.05)
# test with shrinkage
#GC.res.angio.gymno <- lfcShrink(GC.dds.3, coef="plant_clade_Angiospermae_vs_Gymnospermae", type="apeglm")
GC.sig.angio.gymno <- GC.res.angio.gymno[is.na(GC.res.angio.gymno$padj)==F&GC.res.angio.gymno$padj<0.05,] # 738 DAGs
```

### Volcano plot

Plot
```{r}
keyvals <- ifelse(
    GC.res.angio.gymno$padj >= 0.05, 'lightgray',
      ifelse(GC.res.angio.gymno$padj < 0.01 & abs(GC.res.angio.gymno$log2FoldChange) >= 2, 'limegreen',
        'black'))
keyvals[is.na(keyvals)] <- 'black'
names(keyvals)[keyvals == 'lightgray'] <- 'n.s. (p ≥ 0.05)'
names(keyvals)[keyvals == 'black'] <- 'p < 0.05'
names(keyvals)[keyvals == 'limegreen'] <- 'p < 0.01 & log(2)FC ≥ 2 '
#keyvals <- factor(x = keyvals, levels = c('lightgray','black','limegreen'))
#levels(keyvals) <- c('lightgray','black','limegreen')

#keyvals <- keyvals[!is.na(keyvals)]

volcano.AG <-  EnhancedVolcano(GC.res.angio.gymno,
    lab = rownames(GC.res.angio.gymno),
    x = 'log2FoldChange',
    y = 'padj',
    title = "",
    pCutoff = 0.01,
    FCcutoff = 2,
    pointSize = 1.5,
    labSize = 6.0,
    selectLab = "",
    subtitle = "",
    caption = "",
    shape = 16,
    colCustom = keyvals,
    colAlpha = 1,
    cutoffLineCol = 'forestgreen',
    cutoffLineWidth = 0.5,
    hline = 0.05,
    hlineCol = "red",
    hlineType = "dashed",
    hlineWidth = 0.6,
    #vline = c(-1,1),
    #vlineCol = c("firebrick","firebrick"),
    #vlineType = c("dashed","dashed"),
    #vlineWidth = c(0.5,0.5),
    legendPosition = 'right',
    legendLabSize = 13,
    legendIconSize = 4.0,
    drawConnectors = F,
    #widthConnectors = 1.0,
    #colConnectors = 'black',
    #arrowheads = FALSE,
    gridlines.major = TRUE,
    gridlines.minor = FALSE
    #border = 'partial',
    #borderWidth = 1,
    #borderColour = 'black'
    ) +
    coord_cartesian(xlim=c(-4, 4)) +
    scale_x_continuous(breaks=seq(-4,4, 1)) + 
    scale_y_continuous(breaks=seq(0, 9)) + 
    theme_bw() +
    theme(text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_blank(),
      legend.text = element_text(size = 11),
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
volcano.AG

ggsave("../figures/volcano.AG.tiff", width = 6, height = 4)
```

Data
```{r}
volc.AG.data <- volcano.AG[["data"]][volcano.AG [["data"]]$Sig == "FC_P",]
volc.AG.data$gene_cluster_id <- rownames(volc.AG.data)
volc.AG.data <- volc.AG.data[,c(11,1:10)]
rownames(volc.AG.data) <- NULL

#saveRDS(volc.AG.data, "../rds/volc.AG.data.rds")
volc.AG.data <- readRDS("../rds/volc.AG.data.rds")

DAGs.functions.all.info <- read_delim("../DAGs.functions.all.info.csv")
DAGs.functions.all.info <- DAGs.functions.all.info[,c(5,1:4,6:13)]

volc.AG.data.fct <- merge(volc.AG.data, DAGs.functions.all.info[,c(2:4,10:13)], by = "gene_cluster_id")
volc.AG.data.fct <- volc.AG.data.fct[!grepl("^GC_\\d{8}$", volc.AG.data.fct$GC.function), ]
volc.AG.data.fct <- volc.AG.data.fct[order(-volc.AG.data.fct$log2FoldChange), ]
rownames(volc.AG.data.fct) <- NULL

#saveRDS(volc.AG.data.fct, "../rds/volc.AG.data.fct.rds")
volc.AG.data.fct <- readRDS("../rds/volc.AG.data.fct.rds")
```


### Angio vs Gymno
```{r}
GC.sig.Angio.Gymno <- as.data.frame(GC.res.angio.gymno[is.na(GC.res.angio.gymno$padj)==F&GC.res.angio.gymno$padj<0.05&GC.res.angio.gymno$log2FoldChange>0,]) # 468 DAGs
```

```{r}
# Arrange the results dataframe in decreasing log2FoldChange
GC.sig.Angio.Gymno$gene_cluster_id <- as.factor(rownames(GC.sig.Angio.Gymno))
GC.sig.Angio.Gymno <- GC.sig.Angio.Gymno[order(-GC.sig.Angio.Gymno$log2FoldChange), ]
row.names(GC.sig.Angio.Gymno) <- NULL

DAG.Angio.Gymno <- GC.sig.Angio.Gymno

# Add comparisons
DAG.Angio.Gymno$ABBA <- rep(-1,nrow(DAG.Angio.Gymno))
DAG.Angio.Gymno$THOC <- rep(-1,nrow(DAG.Angio.Gymno))
DAG.Angio.Gymno$ACSA <- rep(1,nrow(DAG.Angio.Gymno))
DAG.Angio.Gymno$FAGR <- rep(1,nrow(DAG.Angio.Gymno))
DAG.Angio.Gymno$COCO <- rep(1,nrow(DAG.Angio.Gymno))

# Merge DAGs with function and coverage df
DAG.Angio.Gymno.fct <- merge(GC.rar.fct.cov, DAG.Angio.Gymno, by = "gene_cluster_id")

# Keep rows with unique functional annotation
DAG.Angio.Gymno.fct.distinct <- DAG.Angio.Gymno.fct %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, CAZyme, Description, .keep_all = TRUE)
DAG.Angio.Gymno.fct.distinct <- DAG.Angio.Gymno.fct.distinct[,-c(2:3)] # remove gene caller id and genome name because irrelevant
DAG.Angio.Gymno.fct.distinct <- DAG.Angio.Gymno.fct.distinct[c(1:3,63:64,4:62,65:73)] # rearrange column order
DAG.Angio.Gymno.fct.distinct <- DAG.Angio.Gymno.fct.distinct[order(-DAG.Angio.Gymno.fct.distinct$log2FoldChange), ] # place rows in order of decreasing mean_log2FC
row.names(DAG.Angio.Gymno.fct.distinct) <- NULL
DAG.Angio.Gymno.fct.distinct <- droplevels(DAG.Angio.Gymno.fct.distinct)
str(DAG.Angio.Gymno.fct.distinct)

# Remove from 'DAG.Angio.Gymno.fct.distinct' DAGs that are also in 'DAG.COCO.Trees.fct.distinct'
which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.Angio.Gymno.fct.distinct$gene_cluster_id)
GC.levels.temp <- levels(DAG.COCO.Trees.fct.distinct$gene_cluster_id)
DAG.Angio.Gymno.fct.distinct.only <- subset(DAG.Angio.Gymno.fct.distinct, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.Angio.Gymno.fct.distinct.only) <- NULL
DAG.Angio.Gymno.fct.distinct.only <- droplevels(DAG.Angio.Gymno.fct.distinct.only)
str(DAG.Angio.Gymno.fct.distinct.only)
rm(GC.levels.temp)
```


### Gymno vs Angio
```{r}
GC.res.gymno.angio <- results(GC.dds.3, contrast=c("plant_clade","Gymnospermae","Angiospermae"), alpha = 0.05)
GC.sig.Gymno.Angio <- as.data.frame(GC.res.gymno.angio[is.na(GC.res.gymno.angio$padj)==F&GC.res.gymno.angio$padj<0.05&GC.res.gymno.angio$log2FoldChange>0,]) # 270 DAGs
```

```{r}
# Arrange the results dataframe in decreasing log2FoldChange
GC.sig.Gymno.Angio$gene_cluster_id <- as.factor(rownames(GC.sig.Gymno.Angio))
GC.sig.Gymno.Angio <- GC.sig.Gymno.Angio[order(-GC.sig.Gymno.Angio$log2FoldChange), ]
row.names(GC.sig.Gymno.Angio) <- NULL

DAG.Gymno.Angio <- GC.sig.Gymno.Angio

# Add comparisons
DAG.Gymno.Angio$ABBA <- rep(1,nrow(DAG.Gymno.Angio))
DAG.Gymno.Angio$THOC <- rep(1,nrow(DAG.Gymno.Angio))
DAG.Gymno.Angio$ACSA <- rep(-1,nrow(DAG.Gymno.Angio))
DAG.Gymno.Angio$FAGR <- rep(-1,nrow(DAG.Gymno.Angio))
DAG.Gymno.Angio$COCO <- rep(-1,nrow(DAG.Gymno.Angio))

# Merge DAGs with function and coverage df
DAG.Gymno.Angio.fct <- merge(GC.rar.fct.cov, DAG.Gymno.Angio, by = "gene_cluster_id")

# Keep rows with unique functional annotation
DAG.Gymno.Angio.fct.distinct <- DAG.Gymno.Angio.fct %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, CAZyme, Description, .keep_all = TRUE)
DAG.Gymno.Angio.fct.distinct <- DAG.Gymno.Angio.fct.distinct[,-c(2:3)] # remove gene caller id and genome name because irrelevant
DAG.Gymno.Angio.fct.distinct <- DAG.Gymno.Angio.fct.distinct[c(1:3,63:64,4:62,65:73)] # rearrange column order
DAG.Gymno.Angio.fct.distinct <- DAG.Gymno.Angio.fct.distinct[order(-DAG.Gymno.Angio.fct.distinct$log2FoldChange), ] # place rows in order of decreasing mean_log2FC
row.names(DAG.Gymno.Angio.fct.distinct) <- NULL
DAG.Gymno.Angio.fct.distinct <- droplevels(DAG.Gymno.Angio.fct.distinct)
str(DAG.Gymno.Angio.fct.distinct)

# Remove from 'DAG.Trees.COCO.fct.distinct' DAGs that are also in 'DAG.Gymno.Angio.fct.distinct'
which(DAG.Trees.COCO.fct.distinct$gene_cluster_id %in% DAG.Gymno.Angio.fct.distinct$gene_cluster_id)
GC.levels.temp <- levels(DAG.Gymno.Angio.fct.distinct$gene_cluster_id)
DAG.Trees.COCO.fct.distinct.only <- subset(DAG.Trees.COCO.fct.distinct, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.Trees.COCO.fct.distinct.only) <- NULL
DAG.Trees.COCO.fct.distinct.only <- droplevels(DAG.Trees.COCO.fct.distinct.only)
str(DAG.Trees.COCO.fct.distinct.only)
rm(GC.levels.temp)
```



## Angio Trees vs Gymno Trees

### Analysis
```{r}
GC.dds.4 <- DESeqDataSetFromMatrix(countData = countData[,-c(10:14)],
                              colData = metadata.rar[-c(10:14),], 
                              design = ~ plant_clade)
GC.dds.4$plant_clade <- relevel(GC.dds.4$plant_clade, ref = "Gymnospermae")
GC.dds.4 <- DESeq(GC.dds.4)
resultsNames(GC.dds.4) # lists the coefficients
```

Save and load GC.dds as RDS file
```{r}
#saveRDS(GC.dds.4, "../rds/GC.dds.4_v5.rds")
GC.dds.4 <- readRDS("../rds/GC.dds.4_v5.rds")
```

### Exploration
```{r}
GC.res.angioT.gymnoT <- results(GC.dds.4, contrast=c("plant_clade","Angiospermae","Gymnospermae"), alpha = 0.05)
GC.sig.angioT.gymnoT <- GC.res.angioT.gymnoT[is.na(GC.res.angioT.gymnoT$padj)==F&GC.res.angioT.gymnoT$padj<0.05,] # 228 DAGs
```

### Angio Trees vs Gymno Trees

```{r}
GC.sig.AngioT.GymnoT <- as.data.frame(GC.res.angioT.gymnoT[is.na(GC.res.angioT.gymnoT$padj)==F&GC.res.angioT.gymnoT$padj<0.05&GC.res.angioT.gymnoT$log2FoldChange>0,]) # 189 DAGs
```

```{r}
# Arrange the results dataframe in decreasing log2FoldChange
GC.sig.AngioT.GymnoT$gene_cluster_id <- as.factor(rownames(GC.sig.AngioT.GymnoT))
GC.sig.AngioT.GymnoT <- GC.sig.AngioT.GymnoT[order(-GC.sig.AngioT.GymnoT$log2FoldChange), ]
row.names(GC.sig.AngioT.GymnoT) <- NULL

DAG.AngioT.GymnoT <- GC.sig.AngioT.GymnoT

# Add comparisons
DAG.AngioT.GymnoT$ABBA <- rep(-1,nrow(DAG.AngioT.GymnoT))
DAG.AngioT.GymnoT$THOC <- rep(-1,nrow(DAG.AngioT.GymnoT))
DAG.AngioT.GymnoT$ACSA <- rep(1,nrow(DAG.AngioT.GymnoT))
DAG.AngioT.GymnoT$FAGR <- rep(1,nrow(DAG.AngioT.GymnoT))
DAG.AngioT.GymnoT$COCO <- rep(0,nrow(DAG.AngioT.GymnoT))

# Merge DAGs with function and coverage df
DAG.AngioT.GymnoT.fct <- merge(GC.rar.fct.cov, DAG.AngioT.GymnoT, by = "gene_cluster_id")

# Keep rows with unique functional annotation
DAG.AngioT.GymnoT.fct.distinct <- DAG.AngioT.GymnoT.fct %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, CAZyme, Description, .keep_all = TRUE)
DAG.AngioT.GymnoT.fct.distinct <- DAG.AngioT.GymnoT.fct.distinct[,-c(2:3)] # remove gene caller id and genome name because irrelevant
DAG.AngioT.GymnoT.fct.distinct <- DAG.AngioT.GymnoT.fct.distinct[c(1:3,63:64,4:62,65:73)] # rearrange column order
DAG.AngioT.GymnoT.fct.distinct <- DAG.AngioT.GymnoT.fct.distinct[order(-DAG.AngioT.GymnoT.fct.distinct$log2FoldChange), ] # place rows in order of decreasing mean_log2FC
row.names(DAG.AngioT.GymnoT.fct.distinct) <- NULL
DAG.AngioT.GymnoT.fct.distinct <- droplevels(DAG.AngioT.GymnoT.fct.distinct)
str(DAG.AngioT.GymnoT.fct.distinct)

# Remove from 'DAG.AngioT.GymnoT.fct.distinct' DAGs that are also in 'DAG.Angio.Gymno.fct.distinct.only'
which(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.GymnoT.fct.distinct$gene_cluster_id) #
GC.levels.temp <- levels(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id)
DAG.AngioT.GymnoT.fct.distinct.only <- subset(DAG.AngioT.GymnoT.fct.distinct, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.AngioT.GymnoT.fct.distinct.only) <- NULL
DAG.AngioT.GymnoT.fct.distinct.only <- droplevels(DAG.AngioT.GymnoT.fct.distinct.only)
str(DAG.AngioT.GymnoT.fct.distinct.only)
rm(GC.levels.temp)
```


### Gymno Trees vs Angio Trees 

```{r}
GC.res.gymnoT.angioT <- results(GC.dds.4, contrast=c("plant_clade","Gymnospermae","Angiospermae"), alpha = 0.05)
GC.sig.GymnoT.AngioT <- as.data.frame(GC.res.gymnoT.angioT[is.na(GC.res.gymnoT.angioT$padj)==F&GC.res.gymnoT.angioT$padj<0.05&GC.res.gymnoT.angioT$log2FoldChange>0,]) # 39 DAGs
```

```{r}
# Arrange the results dataframe in decreasing log2FoldChange
GC.sig.GymnoT.AngioT$gene_cluster_id <- as.factor(rownames(GC.sig.GymnoT.AngioT))
GC.sig.GymnoT.AngioT <- GC.sig.GymnoT.AngioT[order(-GC.sig.GymnoT.AngioT$log2FoldChange), ]
row.names(GC.sig.GymnoT.AngioT) <- NULL

DAG.GymnoT.AngioT <- GC.sig.GymnoT.AngioT

# Add comparisons
DAG.GymnoT.AngioT$ABBA <- rep(1,nrow(DAG.GymnoT.AngioT))
DAG.GymnoT.AngioT$THOC <- rep(1,nrow(DAG.GymnoT.AngioT))
DAG.GymnoT.AngioT$ACSA <- rep(-1,nrow(DAG.GymnoT.AngioT))
DAG.GymnoT.AngioT$FAGR <- rep(-1,nrow(DAG.GymnoT.AngioT))
DAG.GymnoT.AngioT$COCO <- rep(0,nrow(DAG.GymnoT.AngioT))

# Merge DAGs with function and coverage df
DAG.GymnoT.AngioT.fct <- merge(GC.rar.fct.cov, DAG.GymnoT.AngioT, by = "gene_cluster_id")

# Keep rows with unique functional annotation
DAG.GymnoT.AngioT.fct.distinct <- DAG.GymnoT.AngioT.fct %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, CAZyme, Description, .keep_all = TRUE)
DAG.GymnoT.AngioT.fct.distinct <- DAG.GymnoT.AngioT.fct.distinct[,-c(2:3)] # remove gene caller id and genome name because irrelevant
DAG.GymnoT.AngioT.fct.distinct <- DAG.GymnoT.AngioT.fct.distinct[c(1:3,63:64,4:62,65:73)] # rearrange column order
DAG.GymnoT.AngioT.fct.distinct <- DAG.GymnoT.AngioT.fct.distinct[order(-DAG.GymnoT.AngioT.fct.distinct$log2FoldChange), ] # place rows in order of decreasing mean_log2FC
row.names(DAG.GymnoT.AngioT.fct.distinct) <- NULL
DAG.GymnoT.AngioT.fct.distinct <- droplevels(DAG.GymnoT.AngioT.fct.distinct)
str(DAG.GymnoT.AngioT.fct.distinct)

# Remove from 'DAG.GymnoT.AngioT.fct.distinct' DAGs that are also in 'DAG.Gymno.Angio.fct.distinct'
which(DAG.Gymno.Angio.fct.distinct$gene_cluster_id %in% DAG.GymnoT.AngioT.fct.distinct$gene_cluster_id) #
GC.levels.temp <- levels(DAG.Gymno.Angio.fct.distinct$gene_cluster_id)
DAG.GymnoT.AngioT.fct.distinct.only <- subset(DAG.GymnoT.AngioT.fct.distinct, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.GymnoT.AngioT.fct.distinct.only) <- NULL
DAG.GymnoT.AngioT.fct.distinct.only <- droplevels(DAG.GymnoT.AngioT.fct.distinct.only)
str(DAG.GymnoT.AngioT.fct.distinct.only)
rm(GC.levels.temp)
```



## ABBA vs THOC & ACSA vs FAGR

### Analysis
```{r}
GC.dds.5 <- DESeqDataSetFromMatrix(countData = countData,
                              colData = metadata.rar, 
                              design = ~ host_species)
GC.dds.5$host_species <- relevel(GC.dds.5$host_species, ref = "COCO")
GC.dds.5 <- DESeq(GC.dds.5)
resultsNames(GC.dds.5) # lists the coefficients
```

Save and load GC.dds as RDS file
```{r}
#saveRDS(GC.dds.5, "../rds/GC.dds.5_v5.rds")
GC.dds.5 <- readRDS("../rds/GC.dds.5_v5.rds")
```

### Exploration
```{r}
GC.res.ABBA.THOC <- results(GC.dds.5, contrast=c("host_species","ABBA","THOC"), alpha = 0.05)
GC.sig.ABBA.THOC <- GC.res.ABBA.THOC[is.na(GC.res.ABBA.THOC$padj)==F&GC.res.ABBA.THOC$padj<0.05,] # 0 DAG

GC.res.ACSA.FAGR <- results(GC.dds.5, contrast=c("host_species","ACSA","FAGR"), alpha = 0.05)
GC.sig.ACSA.FAGR <- GC.res.ACSA.FAGR[is.na(GC.res.ACSA.FAGR$padj)==F&GC.res.ACSA.FAGR$padj<0.05,] # 0 DAG
```



## Angio OR Gymno Trees vs COCO

### Analysis
```{r}
GC.dds.6 <- DESeqDataSetFromMatrix(countData = countData,
                              colData = metadata.rar, 
                              design = ~ gymno_angio_tree)
GC.dds.6$gymno_angio_tree <- relevel(GC.dds.6$gymno_angio_tree, ref = "angiosperm_shrub")
GC.dds.6 <- DESeq(GC.dds.6)
resultsNames(GC.dds.6) # lists the coefficients
```

Save and load GC.dds as RDS file
```{r}
#saveRDS(GC.dds.6, "../rds/GC.dds.6_v5.rds")
GC.dds.6 <- readRDS("../rds/GC.dds.6_v5.rds")
```

### Exploration
```{r}
GC.res.AngioT.COCO <- results(GC.dds.6, contrast=c("gymno_angio_tree","angiosperm_tree","angiosperm_shrub"), alpha = 0.05)
GC.res.COCO.AngioT <- results(GC.dds.6, contrast=c("gymno_angio_tree","angiosperm_shrub","angiosperm_tree"), alpha = 0.05)
GC.res.GymnoT.COCO <- results(GC.dds.6, contrast=c("gymno_angio_tree","gymnosperm_tree","angiosperm_shrub"), alpha = 0.05)
GC.res.COCO.GymnoT <- results(GC.dds.6, contrast=c("gymno_angio_tree","angiosperm_shrub","gymnosperm_tree"), alpha = 0.05)
```

### Angio Trees vs COCO

```{r}
GC.sig.AngioT.COCO <- as.data.frame(GC.res.AngioT.COCO[is.na(GC.res.AngioT.COCO$padj)==F&GC.res.AngioT.COCO$padj<0.05&GC.res.AngioT.COCO$log2FoldChange>0,]) # 118 DAGs
```

```{r}
# Arrange the results dataframe in decreasing log2FoldChange
GC.sig.AngioT.COCO$gene_cluster_id <- as.factor(rownames(GC.sig.AngioT.COCO))
GC.sig.AngioT.COCO <- GC.sig.AngioT.COCO[order(-GC.sig.AngioT.COCO$log2FoldChange), ]
row.names(GC.sig.AngioT.COCO) <- NULL

DAG.AngioT.COCO <- GC.sig.AngioT.COCO

# Add comparisons
DAG.AngioT.COCO$ABBA <- rep(0,nrow(DAG.AngioT.COCO))
DAG.AngioT.COCO$THOC <- rep(0,nrow(DAG.AngioT.COCO))
DAG.AngioT.COCO$ACSA <- rep(1,nrow(DAG.AngioT.COCO))
DAG.AngioT.COCO$FAGR <- rep(1,nrow(DAG.AngioT.COCO))
DAG.AngioT.COCO$COCO <- rep(-1,nrow(DAG.AngioT.COCO))

# Merge DAGs with function and coverage df
DAG.AngioT.COCO.fct <- merge(GC.rar.fct.cov, DAG.AngioT.COCO, by = "gene_cluster_id")

# Keep rows with unique functional annotation
DAG.AngioT.COCO.fct.distinct <- DAG.AngioT.COCO.fct %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, CAZyme, Description, .keep_all = TRUE)
DAG.AngioT.COCO.fct.distinct <- DAG.AngioT.COCO.fct.distinct[,-c(2:3)] # remove gene caller id and genome name because irrelevant
DAG.AngioT.COCO.fct.distinct <- DAG.AngioT.COCO.fct.distinct[c(1:3,63:64,4:62,65:73)] # rearrange column order
DAG.AngioT.COCO.fct.distinct <- DAG.AngioT.COCO.fct.distinct[order(-DAG.AngioT.COCO.fct.distinct$log2FoldChange), ] # place rows in order of decreasing mean_log2FC
row.names(DAG.AngioT.COCO.fct.distinct) <- NULL
DAG.AngioT.COCO.fct.distinct <- droplevels(DAG.AngioT.COCO.fct.distinct)
str(DAG.AngioT.COCO.fct.distinct)

# Remove from 'DAG.AngioT.COCO.fct.distinct' DAGs that are also in 'DAG.Trees.COCO.fct.distinct.only'
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct$gene_cluster_id) #
GC.levels.temp <- levels(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id)
DAG.AngioT.COCO.fct.distinct.only <- subset(DAG.AngioT.COCO.fct.distinct, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.AngioT.COCO.fct.distinct.only) <- NULL
DAG.AngioT.COCO.fct.distinct.only <- droplevels(DAG.AngioT.COCO.fct.distinct.only)
str(DAG.AngioT.COCO.fct.distinct.only)
rm(GC.levels.temp)

# Remove from 'DAG.AngioT.COCO.fct.distinct.only' DAGs that are also in 'DAG.Angio.Gymno.fct.distinct.only'
which(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id) #
GC.levels.temp <- levels(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id)
DAG.AngioT.COCO.fct.distinct.only <- subset(DAG.AngioT.COCO.fct.distinct.only, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.AngioT.COCO.fct.distinct.only) <- NULL
DAG.AngioT.COCO.fct.distinct.only <- droplevels(DAG.AngioT.COCO.fct.distinct.only)
str(DAG.AngioT.COCO.fct.distinct.only)
rm(GC.levels.temp)

# Remove from 'DAG.AngioT.COCO.fct.distinct.only' DAGs that are also in 'DAG.Gymno.Angio.fct.distinct'
which(DAG.Gymno.Angio.fct.distinct$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id) #
GC.levels.temp <- levels(DAG.Gymno.Angio.fct.distinct$gene_cluster_id)
DAG.AngioT.COCO.fct.distinct.only <- subset(DAG.AngioT.COCO.fct.distinct.only, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.AngioT.COCO.fct.distinct.only) <- NULL
DAG.AngioT.COCO.fct.distinct.only <- droplevels(DAG.AngioT.COCO.fct.distinct.only)
str(DAG.AngioT.COCO.fct.distinct.only)
rm(GC.levels.temp)

# Modify 'DAG.AngioT.GymnoT.fct.distinct.only' to include '-1' to COCO column to take into account the results from 'DAG.AngioT.COCO.fct.distinct.only'
which(DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id) #
# MANUAL ENTER, VERIFY IF SOMETHING CHANGE EN AMONT!!
DAG.AngioT.GymnoT.fct.distinct.only$COCO <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,-1,-1,-1,-1)
# Then
# Remove from 'DAG.AngioT.COCO.fct.distinct.only' DAGs that are also in 'DAG.AngioT.GymnoT.fct.distinct.only'
which(DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id) #
GC.levels.temp <- levels(DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id)
DAG.AngioT.COCO.fct.distinct.only <- subset(DAG.AngioT.COCO.fct.distinct.only, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.AngioT.COCO.fct.distinct.only) <- NULL
DAG.AngioT.COCO.fct.distinct.only <- droplevels(DAG.AngioT.COCO.fct.distinct.only)
str(DAG.AngioT.COCO.fct.distinct.only)
rm(GC.levels.temp)
```


### COCO vs Angio Trees

```{r}
GC.sig.COCO.AngioT <- as.data.frame(GC.res.COCO.AngioT[is.na(GC.res.COCO.AngioT$padj)==F&GC.res.COCO.AngioT$padj<0.05&GC.res.COCO.AngioT$log2FoldChange>0,]) # 6 DAGs
```

```{r}
# Arrange the results dataframe in decreasing log2FoldChange
GC.sig.COCO.AngioT$gene_cluster_id <- as.factor(rownames(GC.sig.COCO.AngioT))
GC.sig.COCO.AngioT <- GC.sig.COCO.AngioT[order(-GC.sig.COCO.AngioT$log2FoldChange), ]
row.names(GC.sig.COCO.AngioT) <- NULL

DAG.COCO.AngioT <- GC.sig.COCO.AngioT

# Add comparisons
DAG.COCO.AngioT$ABBA <- rep(0,nrow(DAG.COCO.AngioT))
DAG.COCO.AngioT$THOC <- rep(0,nrow(DAG.COCO.AngioT))
DAG.COCO.AngioT$ACSA <- rep(-1,nrow(DAG.COCO.AngioT))
DAG.COCO.AngioT$FAGR <- rep(-1,nrow(DAG.COCO.AngioT))
DAG.COCO.AngioT$COCO <- rep(1,nrow(DAG.COCO.AngioT))

# Merge DAGs with function and coverage df
DAG.COCO.AngioT.fct <- merge(GC.rar.fct.cov, DAG.COCO.AngioT, by = "gene_cluster_id")

# Keep rows with unique functional annotation
DAG.COCO.AngioT.fct.distinct <- DAG.COCO.AngioT.fct %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, CAZyme, Description, .keep_all = TRUE)
DAG.COCO.AngioT.fct.distinct <- DAG.COCO.AngioT.fct.distinct[,-c(2:3)] # remove gene caller id and genome name because irrelevant
DAG.COCO.AngioT.fct.distinct <- DAG.COCO.AngioT.fct.distinct[c(1:3,63:64,4:62,65:73)] # rearrange column order
DAG.COCO.AngioT.fct.distinct <- DAG.COCO.AngioT.fct.distinct[order(-DAG.COCO.AngioT.fct.distinct$log2FoldChange), ] # place rows in order of decreasing mean_log2FC
row.names(DAG.COCO.AngioT.fct.distinct) <- NULL
DAG.COCO.AngioT.fct.distinct <- droplevels(DAG.COCO.AngioT.fct.distinct)
str(DAG.COCO.AngioT.fct.distinct)

# Remove from 'DAG.COCO.AngioT.fct.distinct' DAGs that are also in 'DAG.COCO.Trees.fct.distinct'
which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.COCO.AngioT.fct.distinct$gene_cluster_id) #
GC.levels.temp <- levels(DAG.COCO.Trees.fct.distinct$gene_cluster_id)
DAG.COCO.AngioT.fct.distinct.only <- subset(DAG.COCO.AngioT.fct.distinct, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.COCO.AngioT.fct.distinct.only) <- NULL
DAG.COCO.AngioT.fct.distinct.only <- droplevels(DAG.COCO.AngioT.fct.distinct.only)
str(DAG.COCO.AngioT.fct.distinct.only)
rm(GC.levels.temp)
```


### Gymno Trees vs COCO

```{r}
GC.sig.GymnoT.COCO <- as.data.frame(GC.res.GymnoT.COCO[is.na(GC.res.GymnoT.COCO$padj)==F&GC.res.GymnoT.COCO$padj<0.05&GC.res.GymnoT.COCO$log2FoldChange>0,]) # 435 DAGs
```

```{r}
# Arrange the results dataframe in decreasing log2FoldChange
GC.sig.GymnoT.COCO$gene_cluster_id <- as.factor(rownames(GC.sig.GymnoT.COCO))
GC.sig.GymnoT.COCO <- GC.sig.GymnoT.COCO[order(-GC.sig.GymnoT.COCO$log2FoldChange), ]
row.names(GC.sig.GymnoT.COCO) <- NULL

DAG.GymnoT.COCO <- GC.sig.GymnoT.COCO

# Add comparisons
DAG.GymnoT.COCO$ABBA <- rep(1,nrow(DAG.GymnoT.COCO))
DAG.GymnoT.COCO$THOC <- rep(1,nrow(DAG.GymnoT.COCO))
DAG.GymnoT.COCO$ACSA <- rep(0,nrow(DAG.GymnoT.COCO))
DAG.GymnoT.COCO$FAGR <- rep(0,nrow(DAG.GymnoT.COCO))
DAG.GymnoT.COCO$COCO <- rep(-1,nrow(DAG.GymnoT.COCO))

# Merge DAGs with function and coverage df
DAG.GymnoT.COCO.fct <- merge(GC.rar.fct.cov, DAG.GymnoT.COCO, by = "gene_cluster_id")

# Keep rows with unique functional annotation
DAG.GymnoT.COCO.fct.distinct <- DAG.GymnoT.COCO.fct %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, CAZyme, Description, .keep_all = TRUE)
DAG.GymnoT.COCO.fct.distinct <- DAG.GymnoT.COCO.fct.distinct[,-c(2:3)] # remove gene caller id and genome name because irrelevant
DAG.GymnoT.COCO.fct.distinct <- DAG.GymnoT.COCO.fct.distinct[c(1:3,63:64,4:62,65:73)] # rearrange column order
DAG.GymnoT.COCO.fct.distinct <- DAG.GymnoT.COCO.fct.distinct[order(-DAG.GymnoT.COCO.fct.distinct$log2FoldChange), ] # place rows in order of decreasing mean_log2FC
row.names(DAG.GymnoT.COCO.fct.distinct) <- NULL
DAG.GymnoT.COCO.fct.distinct <- droplevels(DAG.GymnoT.COCO.fct.distinct)
str(DAG.GymnoT.COCO.fct.distinct)

# Remove from 'DAG.GymnoT.COCO.fct.distinct' DAGs that are also in 'DAG.Trees.COCO.fct.distinct.only'
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.GymnoT.COCO.fct.distinct$gene_cluster_id) #
GC.levels.temp <- levels(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id)
DAG.GymnoT.COCO.fct.distinct.only <- subset(DAG.GymnoT.COCO.fct.distinct, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.GymnoT.COCO.fct.distinct.only) <- NULL
DAG.GymnoT.COCO.fct.distinct.only <- droplevels(DAG.GymnoT.COCO.fct.distinct.only)
str(DAG.GymnoT.COCO.fct.distinct.only)
rm(GC.levels.temp)

# Remove from 'DAG.GymnoT.COCO.fct.distinct.only' DAGs that are also in 'DAG.Gymno.Angio.fct.distinct'
which(DAG.Gymno.Angio.fct.distinct$gene_cluster_id %in% DAG.GymnoT.COCO.fct.distinct.only$gene_cluster_id) #
GC.levels.temp <- levels(DAG.Gymno.Angio.fct.distinct$gene_cluster_id)
DAG.GymnoT.COCO.fct.distinct.only <- subset(DAG.GymnoT.COCO.fct.distinct.only, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.GymnoT.COCO.fct.distinct.only) <- NULL
DAG.GymnoT.COCO.fct.distinct.only <- droplevels(DAG.GymnoT.COCO.fct.distinct.only)
str(DAG.GymnoT.COCO.fct.distinct.only)
rm(GC.levels.temp)
```


### COCO vs Gymno Trees

```{r}
GC.sig.COCO.GymnoT <- as.data.frame(GC.res.COCO.GymnoT[is.na(GC.res.COCO.GymnoT$padj)==F&GC.res.COCO.GymnoT$padj<0.05&GC.res.COCO.GymnoT$log2FoldChange>0,]) # 341 DAGs
```

```{r}
# Arrange the results dataframe in decreasing log2FoldChange
GC.sig.COCO.GymnoT$gene_cluster_id <- as.factor(rownames(GC.sig.COCO.GymnoT))
GC.sig.COCO.GymnoT <- GC.sig.COCO.GymnoT[order(-GC.sig.COCO.GymnoT$log2FoldChange), ]
row.names(GC.sig.COCO.GymnoT) <- NULL

DAG.COCO.GymnoT <- GC.sig.COCO.GymnoT

# Add comparisons
DAG.COCO.GymnoT$ABBA <- rep(-1,nrow(DAG.COCO.GymnoT))
DAG.COCO.GymnoT$THOC <- rep(-1,nrow(DAG.COCO.GymnoT))
DAG.COCO.GymnoT$ACSA <- rep(0,nrow(DAG.COCO.GymnoT))
DAG.COCO.GymnoT$FAGR <- rep(0,nrow(DAG.COCO.GymnoT))
DAG.COCO.GymnoT$COCO <- rep(1,nrow(DAG.COCO.GymnoT))

# Merge DAGs with function and coverage df
DAG.COCO.GymnoT.fct <- merge(GC.rar.fct.cov, DAG.COCO.GymnoT, by = "gene_cluster_id")

# Keep rows with unique functional annotation
DAG.COCO.GymnoT.fct.distinct <- DAG.COCO.GymnoT.fct %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, CAZyme, Description, .keep_all = TRUE)
DAG.COCO.GymnoT.fct.distinct <- DAG.COCO.GymnoT.fct.distinct[,-c(2:3)] # remove gene caller id and genome name because irrelevant
DAG.COCO.GymnoT.fct.distinct <- DAG.COCO.GymnoT.fct.distinct[c(1:3,63:64,4:62,65:73)] # rearrange column order
DAG.COCO.GymnoT.fct.distinct <- DAG.COCO.GymnoT.fct.distinct[order(-DAG.COCO.GymnoT.fct.distinct$log2FoldChange), ] # place rows in order of decreasing mean_log2FC
row.names(DAG.COCO.GymnoT.fct.distinct) <- NULL
DAG.COCO.GymnoT.fct.distinct <- droplevels(DAG.COCO.GymnoT.fct.distinct)
str(DAG.COCO.GymnoT.fct.distinct)

# Remove from 'DAG.COCO.GymnoT.fct.distinct' DAGs that are also in 'DAG.COCO.Trees.fct.distinct'
which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct$gene_cluster_id) #
GC.levels.temp <- levels(DAG.COCO.Trees.fct.distinct$gene_cluster_id)
DAG.COCO.GymnoT.fct.distinct.only <- subset(DAG.COCO.GymnoT.fct.distinct, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.COCO.GymnoT.fct.distinct.only) <- NULL
DAG.COCO.GymnoT.fct.distinct.only <- droplevels(DAG.COCO.GymnoT.fct.distinct.only)
str(DAG.COCO.GymnoT.fct.distinct.only)
rm(GC.levels.temp)

# Remove from 'DAG.COCO.GymnoT.fct.distinct' DAGs that are also in 'DAG.Angio.Gymno.fct.distinct.only'
which(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct.only$gene_cluster_id) #
GC.levels.temp <- levels(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id)
DAG.COCO.GymnoT.fct.distinct.only <- subset(DAG.COCO.GymnoT.fct.distinct.only, !(gene_cluster_id %in% GC.levels.temp))
row.names(DAG.COCO.GymnoT.fct.distinct.only) <- NULL
DAG.COCO.GymnoT.fct.distinct.only <- droplevels(DAG.COCO.GymnoT.fct.distinct.only)
str(DAG.COCO.GymnoT.fct.distinct.only)
rm(GC.levels.temp)
```



## Plots DAGs

### Preparation - Combine all DAGs GC dataframes

List of DAGs dataframes
```{r}
str(DAG.Trees.COCO.fct.distinct.only)
str(DAG.COCO.Trees.fct.distinct)
str(DAG.Angio.Gymno.fct.distinct.only)
str(DAG.Gymno.Angio.fct.distinct)
str(DAG.AngioT.GymnoT.fct.distinct.only)
str(DAG.GymnoT.AngioT.fct.distinct.only)
str(DAG.AngioT.COCO.fct.distinct.only)
str(DAG.COCO.AngioT.fct.distinct.only)
str(DAG.GymnoT.COCO.fct.distinct.only)
str(DAG.COCO.GymnoT.fct.distinct.only)
```

Verify if DAGs duplicates
```{r}
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.COCO.Trees.fct.distinct$gene_cluster_id)
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id)
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.Gymno.Angio.fct.distinct$gene_cluster_id)
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id)
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.GymnoT.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.COCO.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.GymnoT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.Trees.COCO.fct.distinct.only$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct.only$gene_cluster_id)

which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id)
which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.Gymno.Angio.fct.distinct$gene_cluster_id)
which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id)
which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.GymnoT.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.COCO.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.GymnoT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.COCO.Trees.fct.distinct$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct.only$gene_cluster_id)

which(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id %in% DAG.Gymno.Angio.fct.distinct$gene_cluster_id)
which(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id)
which(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id %in% DAG.GymnoT.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id %in% DAG.COCO.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id %in% DAG.GymnoT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.Angio.Gymno.fct.distinct.only$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct.only$gene_cluster_id)

which(DAG.Gymno.Angio.fct.distinct$gene_cluster_id %in% DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id)
which(DAG.Gymno.Angio.fct.distinct$gene_cluster_id %in% DAG.GymnoT.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.Gymno.Angio.fct.distinct$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.Gymno.Angio.fct.distinct$gene_cluster_id %in% DAG.COCO.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.Gymno.Angio.fct.distinct$gene_cluster_id %in% DAG.GymnoT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.Gymno.Angio.fct.distinct$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct.only$gene_cluster_id)

which(DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id %in% DAG.GymnoT.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id %in% DAG.COCO.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id %in% DAG.GymnoT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.AngioT.GymnoT.fct.distinct.only$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct.only$gene_cluster_id)

which(DAG.GymnoT.AngioT.fct.distinct.only$gene_cluster_id %in% DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.GymnoT.AngioT.fct.distinct.only$gene_cluster_id %in% DAG.COCO.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.GymnoT.AngioT.fct.distinct.only$gene_cluster_id %in% DAG.GymnoT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.GymnoT.AngioT.fct.distinct.only$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct.only$gene_cluster_id)

which(DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id %in% DAG.COCO.AngioT.fct.distinct.only$gene_cluster_id)
which(DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id %in% DAG.GymnoT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.AngioT.COCO.fct.distinct.only$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct.only$gene_cluster_id)

which(DAG.COCO.AngioT.fct.distinct.only$gene_cluster_id %in% DAG.GymnoT.COCO.fct.distinct.only$gene_cluster_id)
which(DAG.COCO.AngioT.fct.distinct.only$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct.only$gene_cluster_id)

which(DAG.GymnoT.COCO.fct.distinct.only$gene_cluster_id %in% DAG.COCO.GymnoT.fct.distinct.only$gene_cluster_id)
```

### Prepare DAGs dataframe -->'DAGs' is for plotting (565 unique GC)
```{r}
DAGs.bind <- rbind(
  DAG.Trees.COCO.fct.distinct.only,
  DAG.COCO.Trees.fct.distinct,
  DAG.Gymno.Angio.fct.distinct,
  DAG.Angio.Gymno.fct.distinct.only,
  DAG.GymnoT.AngioT.fct.distinct.only,
  DAG.AngioT.GymnoT.fct.distinct.only,
  DAG.GymnoT.COCO.fct.distinct.only,
  DAG.COCO.GymnoT.fct.distinct.only,
  DAG.AngioT.COCO.fct.distinct.only,
  DAG.COCO.AngioT.fct.distinct.only
  )
#saveRDS(DAGs.bind, "../rds/DAGs.bind.rds")
DAGs.bind <- readRDS("../rds/DAGs.bind.rds")
DAGs <- DAGs.bind %>% distinct(gene_cluster_id, .keep_all = TRUE)
```

VERIFICATION CHUNK ONLY - Unique DAGs verification
```{r}
length(unique(DAGs$gene_cluster_id))
DAGs.fct <- DAGs[,c(1,5,11,19,21,28)] 
rownames(DAGs.fct) <- DAGs.fct$gene_cluster_id
DAGs.fct <- DAGs.fct[,-1]
```





# SCRIPT END




################################################################################
################################################################################
################################################################################

# OLD CODE


Note 251125: This section below code (11 chunks) leads to the first version of the DAGCs table (M.Sc.) (which does not contain COG category and pathways, nor KEGG brite, class and modules), as well as to creation of the long DAGCs figures, and volcano plots. To create latest version of the DAGCs table, see script 'DAGCs_analyses_followup.Rmd'.


### DAGs function curation (GC with functional annotations only)
```{r}
#DAGs.bind$Description <- ifelse(DAGs.bind$Description == "-", "", DAGs.bind$Description)
#DAGs.bind$Description <- ifelse(is.na(DAGs.bind$Description), "", DAGs.bind$Description)

# Functionally annotated GC
DAGs.bind.curated <- DAGs.bind[!(DAGs.bind$COG20_FUNCTION == "" & DAGs.bind$KOfam == "" & DAGs.bind$Pfam == ""), ]
DAGs.bind.curated <- droplevels(DAGs.bind.curated)
# number of unique DAGs that are functionaly annotated
length(unique(DAGs.bind.curated$gene_cluster_id)) # 757 DAGS
# percentage of functionally annotated DAGs
(length(unique(DAGs.bind.curated$gene_cluster_id))/length(unique(DAGs.bind$gene_cluster_id)))*100 # 65.31%

# Unannotated GC
DAGs.bind.wo.fct <- DAGs.bind[DAGs.bind$COG20_FUNCTION == "" & DAGs.bind$KOfam == "" & DAGs.bind$Pfam == "", ]
DAGs.bind.wo.fct <- DAGs.bind.wo.fct %>%
  distinct(gene_cluster_id, .keep_all = TRUE)
DAGs.bind.wo.fct <- droplevels(DAGs.bind.wo.fct)
DAGs.bind.wo.fct <- DAGs.bind.wo.fct[!(DAGs.bind.wo.fct$gene_cluster_id %in% DAGs.bind.curated$gene_cluster_id),]
DAGs.bind.wo.fct <- droplevels(DAGs.bind.wo.fct)
DAGs.bind.wo.fct$GC.function <- DAGs.bind.wo.fct$gene_cluster_id # put GC id name as function
DAGs.bind.wo.fct <- DAGs.bind.wo.fct[,c(1:5,74,6:73)]
# number of unique DAGs that are NOT functionaly annotated
length(unique(DAGs.bind.wo.fct$gene_cluster_id)) # 402 DAGS
# percentage of NON functionally annotated DAGs
(length(unique(DAGs.bind.wo.fct$gene_cluster_id))/length(unique(DAGs.bind$gene_cluster_id)))*100 # 34.69%

# do not keep unique description nor unique CAZyme
DAGs.bind.curated <- DAGs.bind.curated %>%
  distinct(gene_cluster_id, COG20_FUNCTION, KOfam, Pfam, .keep_all = TRUE)
# at this step 'DAGs.bind.curated' contains all COG, KOfam and Pfam annotations.

# create new vector 'GC.function' where the "main" function name will be in priority COG, KOfam, Pfam, then Description
# i.e. if no COG, then KOfam, if not KOfam, then Pfam
DAGs.bind.curated$GC.function <- coalesce(na_if(DAGs.bind.curated$COG20_FUNCTION, ""), na_if(DAGs.bind.curated$KOfam, ""), na_if(DAGs.bind.curated$Pfam, ""))
DAGs.bind.curated <- DAGs.bind.curated %>%
  distinct(gene_cluster_id, GC.function, .keep_all = TRUE)
DAGs.bind.curated <- DAGs.bind.curated[,c(1:5,74,6:73)]
# at this step, if there was multiple KOfam or Pfam for a unique given COG function, than we lose those multiple KOfam and Pfam annotations. However, if for same GC there is one row with one COG function and also one row without COG function, then the KOfam annotation in the row where no COG is found will still be there, even if first row has a COG function for the same GC.

# verify if coalesce worked ok
GC.multi.fct <- DAGs.bind.curated[,c(1,6,12,20,22)]
# get GC with multiple annotations
temp <- as.data.frame(table(DAGs.bind.curated$gene_cluster_id))
temp <- temp[temp$Freq > 1, ]
GC.multi.fct <- GC.multi.fct[GC.multi.fct$gene_cluster_id %in% temp$Var1,]
rm(temp)

# manually curate function list of GC with multiple functional annotations, in excel, than re-import and "merge" into DAGs.bind.curated
# write_csv(GC.multi.fct, file = "../GC.multi.fct.csv")

# to curate, inspect coverage values (non rarefied) of different gene calls associated with different functions
# keep most abundant function (or equally abundant)
temp <- data.fct.cov.c[,c(1:3,11,19,21,42:66)]
rm(temp)
# I favor COG function, then KOfam, then at last Pfam.

# Import curated functions
GC.multi.fct_curated <- read.table("../data_imported/GC.multi.fct_curated.txt", sep = "\t", dec = ".", header = T)

# Create final table with DAGs functions
DAGs.functions <- rbind(GC.multi.fct_curated, DAGs.bind.curated[!DAGs.bind.curated$gene_cluster_id %in% GC.multi.fct_curated$gene_cluster_id, c(1,6,12,20,22)], DAGs.bind.wo.fct[, c(1,6,12,20,22)])
```

### Prepare dataframes for figures and table
```{r}
# Prepare data for DAGs heatmap
DAGs.comm <- DAGs
rownames(DAGs.comm) <- DAGs.comm$gene_cluster_id
DAGs.comm <- DAGs.comm[,c(42:64)]
DAGs.comm <- data.frame(t(DAGs.comm))
abundances.GC <- picante::matrix2sample(DAGs.comm)
colnames(abundances.GC) <- c("sample","abund","gene_cluster_id")
abundances.GC$sample <- factor(x = abundances.GC$sample, levels = c("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))

# Prepare data for DAGs function table
DAGs.functions$gene_cluster_id <- factor(DAGs.functions$gene_cluster_id, levels = levels(abundances.GC$gene_cluster_id))
all(levels(DAGs.functions$gene_cluster_id)==levels(abundances.GC$gene_cluster_id))
DAGs.functions <- DAGs.functions %>% arrange(gene_cluster_id)
rownames(DAGs.functions) <- NULL
all((DAGs.functions$gene_cluster_id)==(DAGs$gene_cluster_id) )
DAGs.functions.all.info <- cbind(DAGs[,c(1:5,65:68)], DAGs.functions[,c(2:5)])

# NOTE I exported 'DAGs.functions.all.info' with the code line below (write_csv) ...
#write_csv(DAGs.functions.all.info, file = "../DAGs.functions.all.info.csv")
# Then in excel, I manually added comparison type in a column of 'DAGs.functions.all.info.csv'.
# SO, I need to import :
DAGs.functions.all.info <- read_delim("../DAGs.functions.all.info.csv")

# Edit 'DAGs.functions.all.info' to not have lfcSE, stat, pval, adj, so all following code should work
# DAGs.functions.all.info <- DAGs.functions.all.info[,c(5,1:4,10:13)]
# OR Keep all columns
DAGs.functions.all.info <- DAGs.functions.all.info[,c(5,1:4,6:13)]

# Prepare data for comparison map
DAGs.compare <- DAGs
rownames(DAGs.compare) <- DAGs.compare$gene_cluster_id
DAGs.compare <- DAGs.compare[,c(69:73)]
DAGs.compare <- data.frame(t(DAGs.compare))
DAGs.compare <- DAGs.compare + 2
compare.GC <- picante::matrix2sample(DAGs.compare)
colnames(compare.GC) <- c("host","compare","gene_cluster_id")
compare.GC$host <- factor(x = compare.GC$host, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
compare.GC$compare <- factor(x = compare.GC$compare, levels = c("3","2","1"))

# Prepare data for effect size (l2FC) 
DAGs.L2FC <- DAGs
DAGs.L2FC$gene_cluster_id <- factor(x = DAGs.L2FC$gene_cluster_id, levels = levels(abundances.GC$gene_cluster_id))

# Prepare data for Accessory vs Core GC
DAGs.Acc.Core <- DAGs
DAGs.Acc.Core$type <- factor(x = DAGs.Acc.Core$type, levels = c("core","accessory"))
DAGs.Acc.Core$gene_cluster_id <- factor(x = DAGs.Acc.Core$gene_cluster_id, levels = levels(abundances.GC$gene_cluster_id))
# number of accessory DAGs
nrow(DAGs.Acc.Core[DAGs.Acc.Core$type=="accessory",]) # 1050
# percentage of accessory DAGs
(nrow(DAGs.Acc.Core[DAGs.Acc.Core$type=="accessory",])/nrow(DAGs.Acc.Core))*100

```

###Plot counts
```{r}
plotCounts(GC.dds.6, gene="GC_00000165", intgroup="plant_clade")
```

### Plot
Long plot with all DAGs
```{r}
gg_heat <- ggplot(abundances.GC, aes(x = sample, y = gene_cluster_id)) + 
  geom_tile(aes(fill=abund)) + 
# geom_tile(aes(fill=sqrt(abund))) + # heat transformed to square root coverage value
  scale_fill_viridis(name = "Normalized\nabundance", option="magma", direction = -1
                    ,trans = "log"
                    ,breaks = c(2,5,12,25,50,100,200,400,800)
                    ) +
#  scale_fill_gradient(high= "darkblue", low = "yellow", name = "Coverage (rarefied)"
#                      ,trans = "sqrt"
#                      ) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Sample") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.25, 0),
      legend.title = element_text(size = 6, margin = margin(b = 7.5)),
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.45, "cm"),
      legend.spacing.y = unit(0.25, "cm"),
      legend.position = "right", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 6, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 6),
      axis.title.x = element_text(color = "black", size = 6.5, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 6.5, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      ) + coord_fixed(ratio = 1)
#gg_heat

gg_comp <- ggplot(compare.GC, aes(x = host, y = gene_cluster_id)) + 
  geom_tile(aes(fill=factor(compare))) + 
# geom_tile(aes(fill=sqrt(abund))) + # heat transformed to square root coverage value
  scale_fill_manual(values = c("3" = "red", "2" = "white", "1" = "blue"), name = "Comparisons", labels = c("Comparison", "Unconsidered", "Reference")) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Host") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.25, 0),
      legend.title = element_text(size = 6, margin = margin(b = 7.5)),
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.45, "cm"),
      legend.spacing.y = unit(0.25, "cm"),
      legend.position = "right", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 6, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 6),
      axis.title.x = element_text(color = "black", size = 6.5, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 6.5, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      ) + coord_fixed(ratio = 1)
#gg_comp

gg_point <- ggplot(DAGs.L2FC, aes(x = log2FoldChange, y = gene_cluster_id)) + 
  geom_point(size = 1.25) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "log2 Fold Change") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.25, 0),
      legend.title = element_text(size = 6, margin = margin(b = 7.5)),
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.45, "cm"),
      legend.spacing.y = unit(0.25, "cm"),
      legend.position = "right", 
      panel.grid.major.x = element_line(color = "gray", size = 0.125), 
      panel.grid.major.y = element_blank(),
      panel.background = element_blank(),
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 6, angle = 0, vjust = 0.35, hjust = 0.5),
      axis.text.y = element_text(color = "black", size = 6),
      axis.title.x = element_text(color = "black", size = 6.5, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 6.5, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
#gg_point

gg_acc.core <- ggplot(DAGs.Acc.Core, aes(x = 1, y = gene_cluster_id, fill = type)) +
  geom_tile() + 
  scale_fill_manual(values = c("core" = "forestgreen", "accessory" = "yellow"), name = "Type", labels = c("Core", "Accessory")) + 
  scale_y_discrete(limits=rev) +
  labs(x = "Type") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.25, 0),
      legend.title = element_text(size = 6, margin = margin(b = 7.5)),
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.45, "cm"),
      legend.spacing.y = unit(0.25, "cm"),
      legend.position = "right", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_blank(),
      axis.text.y = element_text(color = "black", size = 6),
      axis.title.x = element_text(color = "black", size = 6, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 7.5, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      ) + coord_fixed(ratio = 1)
#gg_acc.core


# GC order is the same for both heatmap.

# Remove GC names on gg_comp figure.
#gg_comp <- gg_comp +
#    theme(
#      axis.text.y = element_blank(),
#      axis.title.y = element_blank(),
#      axis.ticks.y = element_blank()
#      )

# Remove GC names on gg_heat and gg_point figures.
gg_heat <- gg_heat +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank()
      )

gg_point <- gg_point +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank()
      )

gg_acc.core <- gg_acc.core +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.ticks.x = element_blank()
      )

((gg_comp + gg_heat) + (gg_acc.core + gg_point))

ggsave("../figures/DAGs_heatmap_v5-2-241115.tiff", width = 8, height = 80, limitsize = FALSE)

gg_fct <- ggplot(DAGs.functions.all.info, aes(x = 1, y = GC.function, label = GC.function)) + 
  geom_text(hjust = 0) + 
  scale_x_continuous(limits = c(0, 2), breaks = NULL) + 
  scale_y_discrete(limits = rev(unique(DAGs.functions.all.info$GC.function))) +  
  theme_void() +  
  theme(
    text = element_text(family = "Arial"),
    axis.text.y = element_text(color = "black", size = 6, hjust = 0),
    axis.text.x = element_blank(),  
    axis.ticks.x = element_blank(),  
    axis.title.x = element_blank(),  
    axis.title.y = element_blank() 
  )

gg_fct

ggsave("../figures/DAGs_functions.tiff", width = 8, height = 80, limitsize = FALSE)


```

## Plot volcanos

### Import data
```{r}
volc.TS.data.fct <- readRDS("../rds/volc.TS.data.fct.rds")
volc.AG.data.fct <- readRDS("../rds/volc.AG.data.fct.rds")
GC.rar.fct.cov <- readRDS("../rds/GC.rar.fct.cov.rds")
```

### Prepare TS and AG data
```{r}
# TS

GC.rar.fct.cov.SUB <- GC.rar.fct.cov[,c(1,42:64)]
GC.rar.fct.cov.SUB <- unique(GC.rar.fct.cov.SUB)

volc.TS.data.fct <- merge(volc.TS.data.fct, GC.rar.fct.cov.SUB, by = "gene_cluster_id")
volc.TS.data.fct <- volc.TS.data.fct[order(-volc.TS.data.fct$log2FoldChange), ]
volc.TS.data.fct$COG20_FUNCTION <- as.factor(volc.TS.data.fct$COG20_FUNCTION)

# AG

volc.AG.data.fct <- merge(volc.AG.data.fct, GC.rar.fct.cov.SUB, by = "gene_cluster_id")
volc.AG.data.fct <- volc.AG.data.fct[order(-volc.AG.data.fct$log2FoldChange), ]
volc.AG.data.fct$COG20_FUNCTION <- as.factor(volc.AG.data.fct$COG20_FUNCTION)
```

Add COG Categories
```{r}
# Prep
cog.cat.fct <- readRDS("../rds/data.fct.cov.rds")
cog.cat.fct <- cog.cat.fct[,c(1,6,11)]
cog.cat.fct <- cog.cat.fct %>%
  distinct(COG20_FUNCTION, .keep_all = TRUE)

# TS
volc.TS.data.fct <- merge(volc.TS.data.fct, cog.cat.fct[, c("COG20_FUNCTION", "COG20_CATEGORY_ACC")], by = "COG20_FUNCTION", all.x = TRUE)
volc.TS.data.fct <- volc.TS.data.fct[,c(2:15,41,1,16:40)]

# Manually add double categories
volc.TS.data.fct$COG20_CATEGORY_ACC <- factor(volc.TS.data.fct$COG20_CATEGORY_ACC, levels = c("J","A","K","L","B","D","Y","V","T","M","N","Z","W","U","O","X","X|X","X|V","C","C|P","G","E","E|H","F","H","I","P","Q","Q|Q","R","S"))
volc.TS.data.fct$COG20_CATEGORY_ACC <- droplevels(volc.TS.data.fct$COG20_CATEGORY_ACC)
volc.TS.data.fct <- volc.TS.data.fct[order(volc.TS.data.fct$COG20_CATEGORY_ACC, -volc.TS.data.fct$log2FoldChange), ]

# AG 
volc.AG.data.fct <- merge(volc.AG.data.fct, cog.cat.fct[, c("COG20_FUNCTION", "COG20_CATEGORY_ACC")], by = "COG20_FUNCTION", all.x = TRUE)
volc.AG.data.fct <- volc.AG.data.fct[,c(2:15,41,1,16:40)]

# Manually add double categories
volc.AG.data.fct$COG20_CATEGORY_ACC <- factor(volc.AG.data.fct$COG20_CATEGORY_ACC, levels = c("J","A","K","K|V","K|G","L","B","D","Y","V","T","T|T","P|T","M","D|M","N","N|W","Z","W","U","U|W","O","X","X|X","X|V","C","C|P","G","E","E|H","F","H","H|R","I","I|S","P","Q","Q|Q","R","S|R","S"))
volc.AG.data.fct$COG20_CATEGORY_ACC <- droplevels(volc.AG.data.fct$COG20_CATEGORY_ACC)
volc.AG.data.fct <- volc.AG.data.fct[order(volc.AG.data.fct$COG20_CATEGORY_ACC, -volc.AG.data.fct$log2FoldChange), ]
```



### Prepare data for heatmap TS TREES-SHRUB
```{r}
volc.TS.heat.data <- volc.TS.data.fct

# Pre-Edit GC.function names
volc.TS.heat.data$GC.function <- gsub("\\(PDB.*?\\)", "", as.character(volc.TS.heat.data$GC.function))
volc.TS.heat.data$GC.function <- gsub("\\(PUBMED.*?\\)", "", as.character(volc.TS.heat.data$GC.function))
volc.TS.heat.data$GC.function <- gsub("\\[EC.*?\\]", "", as.character(volc.TS.heat.data$GC.function))
volc.TS.heat.data$GC.function <- gsub("\\s*\\|", "|", as.character(volc.TS.heat.data$GC.function))
volc.TS.heat.data$GC.function <- paste(volc.TS.heat.data$gene_cluster_id, volc.TS.heat.data$GC.function, sep = ": ")
volc.TS.heat.data$GC.function <- as.factor(volc.TS.heat.data$GC.function)

# Manually edit name on excel.
#write_csv(volc.TS.heat.data, file = "../volc.TS.heat.data_v5.3.csv")
# import
volc.TS.heat.data <- as.data.frame(read_delim("../volc.TS.heat.data_edit_v5.3.csv"))

volc.TS.heat.data$GC.function <- trimws(volc.TS.heat.data$GC.function)
volc.TS.heat.data$GC.function <- as.factor(volc.TS.heat.data$GC.function)
rownames(volc.TS.heat.data) <- volc.TS.heat.data$GC.function
volc.TS.heat.data <- volc.TS.heat.data[,c(19:41)]
volc.TS.heat.data.t <- data.frame(t(volc.TS.heat.data))
colnames(volc.TS.heat.data.t) <- rownames(volc.TS.heat.data)
volc.TS.heat.data <- volc.TS.heat.data.t
volc.TS.heat.data <- picante::matrix2sample(volc.TS.heat.data)
colnames(volc.TS.heat.data) <- c("sample","abund","GC.function")
volc.TS.heat.data$sample <- factor(x = volc.TS.heat.data$sample, levels = c("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
```

### Plot TS TREES-SHRUB
```{r}
# Abundance heatmap

gg_heat_TS <- ggplot(volc.TS.heat.data, aes(x = GC.function, y = sample)) + 
  geom_tile(aes(fill=abund)) + 
# geom_tile(aes(fill=sqrt(abund))) + # heat transformed to square root coverage value
  scale_fill_viridis(name = "Normalized\nabundance", option="magma", direction = -1
                    ,trans = "log"
                    ,breaks = c(2,5,12,25,50,100,200,400,800)
                    ) +
#  scale_fill_gradient(high= "darkblue", low = "yellow", name = "Coverage (rarefied)"
#                      ,trans = "sqrt"
#                      ) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Gene clusters & functions", y = "Samples") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 7.5)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "left", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x.bottom = element_text(vjust = 1),
      axis.text.x = element_text(color = "black", size = 10, angle = 70, hjust = 1),
      axis.text.y = element_text(color = "black", size = 10),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      ) + coord_fixed(ratio = 1)
gg_heat_TS
ggsave("../figures/volcano_DAGCs_heat_TS_v5.3.tiff", width = 8, height = 11, limitsize = FALSE)


# Log2FC 

# Prep
volc.TS.data.fct$gene_cluster_id <- factor(x = volc.TS.data.fct$gene_cluster_id, levels = unique(volc.TS.data.fct$gene_cluster_id))
levels(volc.TS.data.fct$gene_cluster_id)
volc.TS.data.fct$type <- factor(x = volc.TS.data.fct$type, levels = c("core","accessory"))


gg_point_TS <- ggplot(volc.TS.data.fct, aes(x = gene_cluster_id, y = log2FoldChange, fill = type)) + 
  geom_col(width = 0.9) + 
  scale_fill_manual(values = c("core" = "skyblue", "accessory" ="black")) +
  theme(legend.position="none") +
  labs(y = "log(2)FC") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.position = "none", 
      panel.grid.major.y = element_line(color = "gray90", size = 0.25), 
      panel.grid.major.x = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_text(color = "black", size = 10),
      axis.title.x = element_blank(),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
gg_point_TS
ggsave("../figures/volcano_DAGCs_l2fc_TS.tiff", width = 5.85, height = 1, limitsize = FALSE)
```


### Prepare data for heatmap AG ANGIO-GYMNO
```{r}
volc.AG.heat.data <- volc.AG.data.fct

# Pre-Edit GC.function names
volc.AG.heat.data$GC.function <- gsub("\\(PDB.*?\\)", "", as.character(volc.AG.heat.data$GC.function))
volc.AG.heat.data$GC.function <- gsub("\\(PUBMED.*?\\)", "", as.character(volc.AG.heat.data$GC.function))
volc.AG.heat.data$GC.function <- gsub("\\[EC.*?\\]", "", as.character(volc.AG.heat.data$GC.function))
volc.AG.heat.data$GC.function <- gsub("\\s*\\|", "|", as.character(volc.AG.heat.data$GC.function))
volc.AG.heat.data$GC.function <- paste(volc.AG.heat.data$gene_cluster_id, volc.AG.heat.data$GC.function, sep = ": ")
volc.AG.heat.data$GC.function <- as.factor(volc.AG.heat.data$GC.function)

# Manually edit name on excel.
#write_csv(volc.AG.heat.data, file = "../volc.AG.heat.data_v5.3.csv")
# import
volc.AG.heat.data <- as.data.frame(read_delim("../volc.AG.heat.data_edit_v5.3.csv"))

volc.AG.heat.data$GC.function <- trimws(volc.AG.heat.data$GC.function)
volc.AG.heat.data$GC.function <- as.factor(volc.AG.heat.data$GC.function)
rownames(volc.AG.heat.data) <- volc.AG.heat.data$GC.function
volc.AG.heat.data <- volc.AG.heat.data[,c(19:41)]
volc.AG.heat.data.t <- data.frame(t(volc.AG.heat.data))
colnames(volc.AG.heat.data.t) <- rownames(volc.AG.heat.data)
volc.AG.heat.data <- volc.AG.heat.data.t
volc.AG.heat.data <- picante::matrix2sample(volc.AG.heat.data)
colnames(volc.AG.heat.data) <- c("sample","abund","GC.function")
volc.AG.heat.data$sample <- factor(x = volc.AG.heat.data$sample, levels = c("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
```

### Plot AG ANGIO-GYMNO
```{r}
# Abundance heatmap

gg_heat_AG <- ggplot(volc.AG.heat.data, aes(x = GC.function, y = sample)) + 
  geom_tile(aes(fill=abund)) + 
# geom_tile(aes(fill=sqrt(abund))) + # heat transformed to square root coverage value
  scale_fill_viridis(name = "Normalized\nabundance", option="magma", direction = -1
                    ,trans = "log"
                    ,breaks = c(2,5,12,25,50,100,200,400,800)
                    ) +
#  scale_fill_gradient(high= "darkblue", low = "yellow", name = "Coverage (rarefied)"
#                      ,trans = "sqrt"
#                      ) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Gene clusters & functions", y = "Samples") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 7.5)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "left", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x.bottom = element_text(vjust = 1),
      axis.text.x = element_text(color = "black", size = 10, angle = 71, hjust = 1),
      axis.text.y = element_text(color = "black", size = 10),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      ) + coord_fixed(ratio = 1)
gg_heat_AG
ggsave("../figures/volcano_DAGCs_heat_AG_v5.3.tiff", width = 12, height = 12, limitsize = FALSE)



# Log2FC 

# Prep
volc.AG.data.fct$gene_cluster_id <- factor(x = volc.AG.data.fct$gene_cluster_id, levels = unique(volc.AG.data.fct$gene_cluster_id))
levels(volc.AG.data.fct$gene_cluster_id)
volc.AG.data.fct$type <- factor(x = volc.AG.data.fct$type, levels = c("core","accessory"))


gg_point_AG <- ggplot(volc.AG.data.fct, aes(x = gene_cluster_id, y = log2FoldChange, fill = type)) + 
  geom_col(width = 0.9) + 
  scale_fill_manual(values = c("core" = "skyblue", "accessory" ="black")) +
  #coord_cartesian(ylim = c(-4,5))+
  theme(legend.position="none") +
  labs(y = "log(2)FC") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.position = "none", 
      panel.grid.major.y = element_line(color = "gray90", size = 0.25), 
      panel.grid.major.x = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_text(color = "black", size = 10),
      axis.title.x = element_blank(),
      axis.title.y = element_text(color = "black", size = 12, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
gg_point_AG
ggsave("../figures/volcano_DAGCs_l2fc_AG.tiff", width = 12.6, height = 1)
```



Note 251125: This section above code (11 chunks) leads to the first version of the DAGCs table (which does not contain COG category and pathways, nor KEGG brite, class and modules), as well as to creation of the long DAGCs figures, and volcano plots. To create latest version of the DAGCs table, see script 'DAGCs_analyses_followup.Rmd'.



################################################################################

# OLD CODE

Note 251127: This section below leads to the first version of the COG categories figure (M.Sc.)


## Plot COGs

Relative abundances of COG categories within every comparison.
Abondance relative des catégories COG au sein de chaque comparaison.

### Prepare data
```{r}
# Preparation comparison types ('comp' vector)
DAGs.bind <- readRDS("../rds/DAGs.bind.rds")
DAGs.COG.CAT <- DAGs.bind[,c(1,7,69:73)]
DAGs.COG.CAT[,c(3:7)] <- DAGs.COG.CAT[,c(3:7)] + 2  # Transform '-1' to '1'; '0' to '2'; and '1' to '3'
DAGs.COG.CAT$comp <- apply(DAGs.COG.CAT[,c(3:7)], 1, function(x) {
  paste(x, collapse = "")  # Combine values into a single string
})
DAGs.COG.CAT$comp <- as.factor(DAGs.COG.CAT$comp)

# Change empty values (COG20 Categories) for "No annotation"
DAGs.COG.CAT$COG20_CATEGORY <- factor(ifelse(DAGs.COG.CAT$COG20_CATEGORY == "", "No annotation", as.character(DAGs.COG.CAT$COG20_CATEGORY)))

# For every gene call that have multiple (n) COG categories, make n rows, one row per category.
DAGs.COG.CAT.temp <- DAGs.COG.CAT
DAGs.COG.CAT.temp$COG20_CATEGORY <- as.character(DAGs.COG.CAT.temp$COG20_CATEGORY)
# Add row number column
df_split1 <- DAGs.COG.CAT.temp %>%
  rowid_to_column()  # Add a column representing the original row number
# Split rows based on "|" and unnest
df_split2 <- df_split1 %>%
  mutate(COG20_CATEGORY = strsplit(COG20_CATEGORY, "\\|")) %>%
  unnest(cols = COG20_CATEGORY)
# Remove duplicates within rows
df_split3 <- df_split2 %>%
  distinct(gene_cluster_id, COG20_CATEGORY, .keep_all = TRUE)  # Keep the first occurrence of each gene_cluster_id-COG20_CATEGORY combination
# Remove rowid column
df_split4 <- df_split3 %>%
  select(-rowid)
DAGs.COG.CAT.multi <- as.data.frame(df_split4)
rm(DAGs.COG.CAT.temp, df_split1, df_split2, df_split3, df_split4)
# arrange vectors in df
DAGs.COG.CAT.multi$COG20_CATEGORY <- as.factor(DAGs.COG.CAT.multi$COG20_CATEGORY)
DAGs.COG.CAT.multi$comp <- factor(DAGs.COG.CAT.multi$comp, levels = unique(DAGs.COG.CAT.multi$comp))

# Count each COG categories (COG.CAT) counts for each comparison type
DAGs.COG.CAT.counts = DAGs.COG.CAT.multi %>% 
  group_by(comp, COG20_CATEGORY) %>% 
  summarise(COG.CAT.counts.per.comp=n())

# Calculate total COG categories counts (COG.CAT.counts) per comparison type (total.COG.counts.per.comp)
comp.counts.ORIG <- DAGs.COG.CAT.counts %>%
  group_by(comp) %>%
  summarise(total.COG.counts.per.comp = sum(COG.CAT.counts.per.comp))
#### RESULTS ####
# I CAN USE 'comp.counts.ORIG' TO KNOW HOW MANY COUNTS FOR EACH COMPARISONS BEFORE REMOVING "No annotation" and "Function unknown" (NOTE: it is not how many DAGs for each comparison, because some DAGs have multiple categories assigned to).

# Calculate total COG categories counts (COG.CAT.counts) across comparisons type
COG.c.ORIG <- DAGs.COG.CAT.counts %>%
  group_by(COG20_CATEGORY) %>%
  summarise(COG.c = sum(COG.CAT.counts.per.comp)) %>% 
  ungroup()
#### RESULT ####
# I CAN USE 'COG.c.ORIG' TO KNOW HOW MANY COUNTS FOR EACH COG CATEGORIES INCLUDING "No annotation" and "Function unknown" across comparison types
# THERE IS 605 COUNTS OF "NO ANNOTATION" AND 49 COUNTS OF "FUNCTION UNKNOWN"

# Remove "No annotation" and "Function unknown"
DAGs.COG.CAT.counts <- subset(DAGs.COG.CAT.counts, !(COG20_CATEGORY %in% c("No annotation", "Function unknown")))

# Re-calculate total counts (COG.CAT.counts) per comparison type (total.COG.counts.per.comp) without "function unknown" nor "no annotation" 
comp.counts <- DAGs.COG.CAT.counts %>%
  group_by(comp) %>%
  summarise(total.COG.counts.per.comp = sum(COG.CAT.counts.per.comp))

DAGs.COG.CAT.counts <- merge(DAGs.COG.CAT.counts, comp.counts, by = "comp")
DAGs.COG.CAT.counts <- DAGs.COG.CAT.counts %>%
  arrange(comp)

# calculate relative abundance of each COG categories within comparison type
DAGs.COG.CAT.counts$rel.abund.COG.CAT.in.comp <- DAGs.COG.CAT.counts$COG.CAT.counts.per.comp/DAGs.COG.CAT.counts$total.COG.counts.per.comp

# Verify data 
sum(DAGs.COG.CAT.counts[DAGs.COG.CAT.counts$comp == "33331",]$rel.abund.COG.CAT.in.comp) # should sum to 1
sum(DAGs.COG.CAT.counts$rel.abund.COG.CAT.in.comp) # should sum to the number of comparison types (11)


# DO NOT RUN BELOW HERE. NOT NEEDED. KEPT AS BACKUP.

# Calculate relative abundance of categories across comparisons type (based on normalized COG.CAT counts)
#total.norm.COG.CAT.counts.across.comp <- DAGs.COG.CAT.counts %>%
#  group_by(COG20_CATEGORY) %>%
#  summarise(total.norm.COG.CAT.counts.across.comp = sum(norm.COG.CAT.counts.per.comp)) %>% 
#  ungroup()

#DAGs.COG.CAT.prop <- merge(DAGs.COG.CAT.counts, total.norm.COG.CAT.counts.across.comp, by = "COG20_CATEGORY")
#DAGs.COG.CAT.prop <- DAGs.COG.CAT.prop %>%
#  arrange(comp)

#DAGs.COG.CAT.prop$norm.rel.abund <- DAGs.COG.CAT.prop$norm.COG.CAT.counts.per.comp/DAGs.COG.CAT.prop$total.norm.COG.CAT.counts.across.comp

# test if it worked
#sum(DAGs.COG.CAT.prop[DAGs.COG.CAT.prop$COG20_CATEGORY == "Carbohydrate transport and metabolism",]$norm.rel.abund) # should sum to 1
#sum(DAGs.COG.CAT.prop$norm.COG.CAT.counts.per.comp) # should sum to the number of comparison types (7 in v5)
```


Prepare data for COG (calculate number of counts for each COG categories across comparison types)
```{r}
COG.c <- DAGs.COG.CAT.counts %>%
  group_by(COG20_CATEGORY) %>%
  summarise(COG.c = sum(COG.CAT.counts.per.comp)) %>% 
  ungroup()
```

Prepare data for comp (calculate number of COG counts for each comparison types across COG categories)
```{r}
# THIS IS NOT DAGs COUNTS!!!
# with "Function unknown" and "No annotation"
comp.counts.ORIG
# without "Function unknown" and "No annotation"
comp.c <- DAGs.COG.CAT.counts %>%
  distinct(comp, .keep_all = TRUE) 
comp.c <- comp.c[,c(1,4)]
comp.c$comp <- droplevels(comp.c$comp)
```

Prepare data for DAGs counts per comp
```{r}
DAGs.c <- DAGs.COG.CAT %>%
  distinct(gene_cluster_id, comp, .keep_all = F)

DAGs.c <- DAGs.c %>%
  group_by(comp) %>%
  summarise(DAGs.c = n_distinct(gene_cluster_id))
DAGs.c$comp <- factor(DAGs.c$comp, levels = levels(comp.c$comp))
DAGs.c <- DAGs.c[order(DAGs.c$comp), ]
```


Sorting dataframes in decreasing amount of total COG counts (so that in plot, COG category with the highest amount of counts across comparison types will be at the top of y axis)
```{r}
COG.c <- COG.c[order(-COG.c$COG.c), ]
COG.c$COG20_CATEGORY <- factor(COG.c$COG20_CATEGORY, levels = unique(COG.c$COG20_CATEGORY))
levels(COG.c$COG20_CATEGORY)
DAGs.COG.CAT.counts$COG20_CATEGORY <- factor(DAGs.COG.CAT.counts$COG20_CATEGORY, levels = levels(factor(COG.c$COG20_CATEGORY)))
all(levels(DAGs.COG.CAT.counts$COG20_CATEGORY)==levels(COG.c$COG20_CATEGORY))
```


### Plot

```{r}
# Plot relative abundance of COG categories within each comparison type
gg_heat.comp <- ggplot(DAGs.COG.CAT.counts, aes(x = comp, y = COG20_CATEGORY)) + 
  geom_tile(aes(fill=rel.abund.COG.CAT.in.comp)) + 
# geom_tile(aes(fill=sqrt(abund))) + # heat transformed to square root coverage value
  scale_fill_viridis(name = "Relative\nabundance", option="magma", direction = -1, limits = c(NA, 1)
                    ,trans = "sqrt"
                    ,breaks = c(0.10,0.25,0.5,1)
                    ) +
#  scale_fill_gradient(high= "darkblue", low = "yellow", name = "Coverage (rarefied)"
#                      ,trans = "sqrt"
#                      ) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Comparisons") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 12),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_blank(),
      axis.ticks.x = element_blank()      
      ) + coord_fixed(ratio = 1)
gg_heat.comp

# Plot raw counts of each COG category in each comparison type
gg_heat.comp <- ggplot(DAGs.COG.CAT.counts, aes(x = comp, y = COG20_CATEGORY)) + 
  geom_tile(aes(fill=COG.CAT.counts.per.comp)) + 
# geom_tile(aes(fill=sqrt(abund))) + # heat transformed to square root coverage value
  scale_fill_viridis(name = "Number of genes", option="magma", direction = -1, limits = c(NA, 27)
#                    ,trans = "sqrt"
#                    ,breaks = c(0.10,0.25,0.5,1)
                    ) +
#  scale_fill_gradient(high= "darkblue", low = "yellow", name = "Coverage (rarefied)"
#                      ,trans = "sqrt"
#                      ) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Comparisons") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 12),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_blank(),
      axis.ticks.x = element_blank()      
      ) + coord_fixed(ratio = 1)
gg_heat.comp




gg_COG.c <-  ggplot(COG.c, aes(x = COG.c, y = COG20_CATEGORY)) + 
  geom_col(fill = "black") + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Number of genes\nacross comparisons") +
  coord_cartesian(xlim = c(0,70))+
  scale_x_continuous(breaks = seq(0, 70, by = 10))+
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major.x = element_line(color = "gray", size = 0.25), 
      panel.grid.major.y = element_blank(),
      panel.background = element_blank(),
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 0, vjust = 0.35, hjust = 0.5),
#      axis.text.y = element_text(color = "black", size = 12),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
#      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold"),
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank()
      ) + coord_fixed(ratio = 6)
gg_COG.c

(gg_heat.comp + gg_COG.c)

#ggsave("../figures/DAGs_COGs_comp_heatmap_v5-3.tiff", width = 11, height = 5)
```

### Complementary Plots for COGs

### Prepare data
```{r}
comparisons <- unique(DAGs.COG.CAT[,c(3:8)])
comparisons <- comparisons[comparisons$comp %in% levels(comp.c$comp),]
rownames(comparisons) <- comparisons$comp
comparisons <- comparisons[,-6]
comparisons <- as.data.frame(t(comparisons))
comparisons <- picante::matrix2sample(comparisons)
colnames(comparisons) <- c("host","compare","comp_id")
comparisons$host <- factor(x = comparisons$host, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
comparisons$compare <- factor(x = comparisons$compare, levels = c("3","2","1"))
```

### Plot
```{r}
# Number of COG annotations per comparison (without "Funciton unknown" nor "No annotation")
gg_comp.c <- ggplot(comp.c, aes(x = comp, y = total.COG.counts.per.comp)) + 
  geom_col(fill = "black") +
  theme(legend.position="none") +
  labs(y = "Number of annotations\n per comparison") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major.x = element_blank(), 
      panel.grid.major.y = element_line(color = "gray", size = 0.25),
      panel.background = element_blank(),
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
#      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.x = element_blank(),
      axis.text.y = element_text(color = "black", size = 12),
#      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold"),
#      axis.text.y = element_blank(),
      axis.title.x = element_blank(),
      axis.ticks.x = element_blank()
      )
gg_comp.c

#ggsave("../figures/DAGs_COG_counts_per_comp_v5-2.tiff", width = 2.875, height = 2)

# Number of DAGs per comparison types (notwithstanding COG categories)
gg_dags.c <- ggplot(DAGs.c, aes(x = comp, y = DAGs.c)) + 
  geom_col(fill = "black") +
  theme(legend.position="none") +
  labs(y = "Number of DAGCs\n per comparison") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major.x = element_blank(), 
      panel.grid.major.y = element_line(color = "gray", size = 0.25),
      panel.background = element_blank(),
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
#      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.x = element_blank(),
      axis.text.y = element_text(color = "black", size = 12),
#      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold"),
#      axis.text.y = element_blank(),
      axis.title.x = element_blank(),
      axis.ticks.x = element_blank()
      )
gg_dags.c

#ggsave("../figures/DAGs_counts_per_comp_v5-2.tiff", width = 2.9, height = 2)


# Comparison types red-blue heatmap
gg_comp.small <- ggplot(comparisons, aes(x = comp_id, y = host)) + 
  geom_tile(aes(fill=factor(compare))) + 
  scale_fill_manual(values = c("1" = "blue", "2" = "white", "3" = "red")
#                   , name = "Comparisons", labels = c("Comparison", "Unconsidered", "Reference")
                    ) + 
  theme(legend.position="none") +
  guides(fill = "none") +
  scale_y_discrete(limits=rev, labels = c("ABBA" = "A. balsamea", "THOC" = "T. occidentalis","ACSA" = "A. Saccharum", "FAGR" = "F. grandifolia", "COCO" = "C. cornuta"))+
  labs(y = "Host\nspecies") +
  labs(x = "Comparisons") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
#      legend.justification = c(0.5, 0),
#      legend.title = element_text(size = 12, margin = margin(b = 1)),
#      legend.text = element_text(size = 12),
#      legend.key.size = unit(0.6, "cm"),
#      legend.spacing.y = unit(0.5, "cm"),
#      legend.position = "right", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_blank(),
      axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0.5, face = "italic"),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = 2, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold"),
      axis.ticks.x = element_blank()
      ) + coord_fixed(ratio = 1)  
gg_comp.small  

#ggsave("../figures/DAGs_comp_heatmap_v5-2.tiff", width = 3.756, height = 1.31)
```





################################################################################

# OLD CODE

Plot FUL-SIZE. DO NOT. See HALF-SIZE NEXT CHUNK
Long plot with all DAGs
```{r}
gg_heat <- ggplot(abundances.GC, aes(x = sample, y = gene_cluster_id)) + 
  geom_tile(aes(fill=abund)) + 
# geom_tile(aes(fill=sqrt(abund))) + # heat transformed to square root coverage value
  scale_fill_viridis(name = "Abundance", option="magma", direction = -1
#                    ,trans = "sqrt"
                    ) +
#  scale_fill_gradient(high= "darkblue", low = "yellow", name = "Coverage (rarefied)"
#                      ,trans = "sqrt"
#                      ) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Sample") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 12),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )+ coord_fixed(ratio = 1)
#gg_heat

gg_comp <- ggplot(compare.GC, aes(x = host, y = gene_cluster_id)) + 
  geom_tile(aes(fill=factor(compare))) + 
# geom_tile(aes(fill=sqrt(abund))) + # heat transformed to square root coverage value
  scale_fill_manual(values = c("3" = "red", "2" = "white", "1" = "blue"), name = "Comparisons", labels = c("Comparison", "Unconsidered", "Reference")) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Host") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 12),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      ) + coord_fixed(ratio = 1)
#gg_comp

gg_point <- ggplot(DAGs.L2FC, aes(x = log2FoldChange, y = gene_cluster_id)) + 
  geom_point(size = 3) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "log2 Fold Change") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major.x = element_line(color = "gray", size = 0.25), 
      panel.grid.major.y = element_blank(),
      panel.background = element_blank(),
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 12),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
#gg_point


# GC order is the same for both heatmap.

# Remove GC names on gg_comp figure.
#gg_comp <- gg_comp +
#    theme(
#      axis.text.y = element_blank(),
#      axis.title.y = element_blank(),
#      axis.ticks.y = element_blank()
#      )

# Remove GC names on gg_heat and gg_point figures.
gg_heat <- gg_heat +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank()
      )

gg_point <- gg_point +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank()
      )

(gg_comp + gg_heat + gg_point)

ggsave("../figures/DAGs_heatmap_v5-2.tiff", width = 15, height = 160, limitsize = FALSE)
```



  ###### DO NOT RUN BELOW THIS IS JUST A TEST ######

Prepare data (take 1)
```{r}
# Preparation
DAGs.COG.CAT <- DAGs.bind[,c(1,6,61:65)]
DAGs.COG.CAT[,c(3:7)] <- DAGs.COG.CAT[,c(3:7)] + 2

DAGs.COG.CAT$comp <- apply(DAGs.COG.CAT[,c(3:7)], 1, function(x) {
  paste(x, collapse = "")  # Combine values into a single string
})

DAGs.COG.CAT$comp <- as.factor(DAGs.COG.CAT$comp)

# Change empty values for "No annotation"
DAGs.COG.CAT$COG20_CATEGORY <- factor(ifelse(DAGs.COG.CAT$COG20_CATEGORY == "", "No annotation", as.character(DAGs.COG.CAT$COG20_CATEGORY)))
  
# For every gene call that have multiple (n) COG categories, make n rows, one row per category.
DAGs.COG.CAT.temp <- DAGs.COG.CAT
DAGs.COG.CAT.temp$COG20_CATEGORY <- as.character(DAGs.COG.CAT.temp$COG20_CATEGORY)

# Add row number column
df_split1 <- DAGs.COG.CAT.temp %>%
  rowid_to_column()  # Add a column representing the original row number

# Split rows based on "|" and unnest
df_split2 <- df_split1 %>%
  mutate(COG20_CATEGORY = strsplit(COG20_CATEGORY, "\\|")) %>%
  unnest(cols = COG20_CATEGORY)

# Remove duplicates within rows
df_split3 <- df_split2 %>%
  distinct(gene_cluster_id, COG20_CATEGORY, .keep_all = TRUE)  # Keep the first occurrence of each gene_cluster_id-COG20_CATEGORY combination

# Remove rowid column
df_split4 <- df_split3 %>%
  select(-rowid)

DAGs.COG.CAT.multi <- as.data.frame(df_split4)
DAGs.COG.CAT.multi$COG20_CATEGORY <- as.factor(DAGs.COG.CAT.multi$COG20_CATEGORY)
DAGs.COG.CAT.multi$comp <- factor(DAGs.COG.CAT.multi$comp, levels = unique(DAGs.COG.CAT.multi$comp))

rm(DAGs.COG.CAT.temp, df_split1, df_split2, df_split3, df_split4)

# Count each COG categories counts at each comparison type
DAGs.COG.CAT.counts = DAGs.COG.CAT.multi %>% 
  group_by(comp, COG20_CATEGORY) %>% 
  summarise(COG.counts=n())

# Calculate relative abundance of categories across comparisons type
DAGs.COG.CAT.prop <- DAGs.COG.CAT.counts %>%
  group_by(COG20_CATEGORY) %>%
  summarise(total_CAT.count = sum(COG.counts)) %>% 
  ungroup()

DAGs.COG.CAT.prop <- merge(DAGs.COG.CAT.counts, DAGs.COG.CAT.prop, by = "COG20_CATEGORY")
DAGs.COG.CAT.prop <- DAGs.COG.CAT.prop %>%
  arrange(comp)

DAGs.COG.CAT.prop$rel.abund <- DAGs.COG.CAT.prop$COG.counts/DAGs.COG.CAT.prop$total_CAT.count

# Calculate normalizing factor to apply to COG counts relatives abundances
comp.counts <- DAGs.COG.CAT.prop %>%
  group_by(comp) %>%
  summarise(comp.count = sum(COG.counts))
DAGs.COG.CAT.prop <- merge(DAGs.COG.CAT.prop, comp.counts, by = "comp")
DAGs.COG.CAT.prop <- DAGs.COG.CAT.prop %>%
  arrange(comp)
DAGs.COG.CAT.prop$norm.fact <- DAGs.COG.CAT.prop$comp.count/sum(comp.counts$comp.count)

# apply normalization
DAGs.COG.CAT.prop$norm.abund <- DAGs.COG.CAT.prop$rel.abund/DAGs.COG.CAT.prop$norm.fact



# archives - DO NOT RUN
# Calculate normalizing factor to apply to COG counts relatives abundances
comp.counts <- as.data.frame(apply(DAGs[,c(61:65)] + 2, 1, function(x) {
  paste(x, collapse = "")
}))
levels.temp <- levels(factor(comp.counts$comp, levels = unique(comp.counts$comp)))
comp.counts <- as.data.frame(table(comp.counts$comp))
colnames(comp.counts) <- c("comp","comp.count")
comp.counts$comp <- factor(x = comp.counts$comp, levels = levels.temp)
comp.counts <- comp.counts %>%
  arrange(comp)
DAGs.COG.CAT.prop$test <- DAGs.COG.CAT.prop$rel.abund/DAGs.COG.CAT.prop$norm.fact
rm(levels.temp)

sum(DAGs.COG.CAT.prop[DAGs.COG.CAT.prop$COG20_CATEGORY == "Amino acid transport and metabolism",]$norm.abund)
```

Plot 2.1
Normalized relative abundances of COG categories across all comparisons.

```{r}
gg_heat.COG <- ggplot(DAGs.COG.CAT.prop, aes(x = comp, y = COG20_CATEGORY)) + 
  geom_tile(aes(fill=norm.abund)) + 
# geom_tile(aes(fill=sqrt(abund))) + # heat transformed to square root coverage value
  scale_fill_viridis(name = "Relative Abundance", option="magma", direction = -1
                    ,trans = "sqrt"
                    ) +
#  scale_fill_gradient(high= "darkblue", low = "yellow", name = "Coverage (rarefied)"
#                      ,trans = "sqrt"
#                      ) + 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "COG Categories") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major = element_blank(), 
      panel.background = element_blank(),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 12),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
gg_heat.COG
```


  ###### DO NOT RUN ABOVE THIS IS JUST A TEST ######


################################################################################


################################################################################

  ###### DO NOT RUN BELOW THIS IS JUST A TEST ######

DAGs NON rarefied and NON filtered data

```{r}
countData <- as.data.frame(t(data.cov.c.agg.comm))
sum(rowSums(countData)==0)
colSums(countData)
```

Make sure count dataframe's columns and metadata's rows are in the same order
```{r}
all(rownames(metadata) == colnames(countData))
```

Analysis
```{r}
GC.dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = metadata, 
                              design = ~ host_species)
# Make COCO reference level
GC.dds$host_species <- relevel(GC.dds$host_species, ref = "COCO")

GC.dds <- DESeq(GC.dds, fitType = 'local')

resultsNames(GC.dds) # lists the coefficients
```

```{r}
GC.res.ABBA.ACSA <- results(GC.dds, contrast=c("host_species","ABBA","ACSA"), alpha = 0.05)
GC.res.ABBA.COCO <- results(GC.dds, contrast=c("host_species","ABBA","COCO"), alpha = 0.05)
GC.res.ABBA.FAGR <- results(GC.dds, contrast=c("host_species","ABBA","FAGR"), alpha = 0.05)
GC.res.ABBA.THOC <- results(GC.dds, contrast=c("host_species","ABBA","THOC"), alpha = 0.05)
GC.res.ACSA.COCO <- results(GC.dds, contrast=c("host_species","ACSA","COCO"), alpha = 0.05)
GC.res.ACSA.FAGR <- results(GC.dds, contrast=c("host_species","ACSA","FAGR"), alpha = 0.05)
GC.res.ACSA.THOC <- results(GC.dds, contrast=c("host_species","ACSA","THOC"), alpha = 0.05)
GC.res.COCO.FAGR <- results(GC.dds, contrast=c("host_species","COCO","FAGR"), alpha = 0.05)
GC.res.COCO.THOC <- results(GC.dds, contrast=c("host_species","COCO","THOC"), alpha = 0.05)
GC.res.FAGR.THOC <- results(GC.dds, contrast=c("host_species","FAGR","THOC"), alpha = 0.05)
```

Analyzing results
```{r}
# GC with signicant differential abundances between host tree species (adjusted_p_value<0.05)
GC.sig.ABBA.ACSA <- GC.res.ABBA.ACSA[is.na(GC.res.ABBA.ACSA$padj)==F&GC.res.ABBA.ACSA$padj<0.05,]
GC.sig.ABBA.COCO <- GC.res.ABBA.COCO[is.na(GC.res.ABBA.COCO$padj)==F&GC.res.ABBA.COCO$padj<0.05,]
GC.sig.ABBA.FAGR <- GC.res.ABBA.FAGR[is.na(GC.res.ABBA.FAGR$padj)==F&GC.res.ABBA.FAGR$padj<0.05,]
GC.sig.ABBA.THOC <- GC.res.ABBA.THOC[is.na(GC.res.ABBA.THOC$padj)==F&GC.res.ABBA.THOC$padj<0.05,] 
GC.sig.ACSA.COCO <- GC.res.ACSA.COCO[is.na(GC.res.ACSA.COCO$padj)==F&GC.res.ACSA.COCO$padj<0.05,]
GC.sig.ACSA.FAGR <- GC.res.ACSA.FAGR[is.na(GC.res.ACSA.FAGR$padj)==F&GC.res.ACSA.FAGR$padj<0.05,]
GC.sig.ACSA.THOC <- GC.res.ACSA.THOC[is.na(GC.res.ACSA.THOC$padj)==F&GC.res.ACSA.THOC$padj<0.05,] 
GC.sig.COCO.FAGR <- GC.res.COCO.FAGR[is.na(GC.res.COCO.FAGR$padj)==F&GC.res.COCO.FAGR$padj<0.05,] 
GC.sig.COCO.THOC <- GC.res.COCO.THOC[is.na(GC.res.COCO.THOC$padj)==F&GC.res.COCO.THOC$padj<0.05,] 
GC.sig.FAGR.THOC <- GC.res.FAGR.THOC[is.na(GC.res.FAGR.THOC$padj)==F&GC.res.FAGR.THOC$padj<0.05,] 
```

  ###### DO NOT RUN ABOVE THIS IS JUST A TEST ######


################################################################################



