---
title: "Methylobacterium species beta diversity analyses"
output: html_document
date: "2024-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(tidyverse)
library(vegan)
library(ggplot2)
library(ggtree)
library(patchwork)
library(rlang)
library(ggrepel)
library(ggdark)
library(pairwiseAdonis)
library(extrafont)
library(ggordiplots)
library(betapart)
library(picante)
library(gridExtra)
```

################################################################################

# Data import

## Detection

Import anvio genome detection data
```{r}
genome.detect <- read.delim("/mnt/storage/jocelyn/msc/anvio_analysis/ABBA/ABBA_PROFILE-SUMMARY/bins_across_samples/detection.txt", header = T, sep = "\t", dec = ".")
genome.detect <- merge(genome.detect, read.delim("/mnt/storage/jocelyn/msc/anvio_analysis/ACSA/ACSA_PROFILE-SUMMARY/bins_across_samples/detection.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.detect <- merge(genome.detect, read.delim("/mnt/storage/jocelyn/msc/anvio_analysis/COCO/COCO_PROFILE-SUMMARY/bins_across_samples/detection.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.detect <- merge(genome.detect, read.delim("/mnt/storage/jocelyn/msc/anvio_analysis/FAGR/FAGR_PROFILE-SUMMARY/bins_across_samples/detection.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.detect <- merge(genome.detect, read.delim("/mnt/storage/jocelyn/msc/anvio_analysis/THOC/THOC_PROFILE-SUMMARY/bins_across_samples/detection.txt", header = T, sep = "\t", dec = "."), by = "bins")
colnames(genome.detect)[1] <- "species"
rownames(genome.detect) <- genome.detect$species
genome.detect <- genome.detect[,-1]
#saveRDS(genome.detect, "../rds/genome.detect.rds")
genome.detect <- readRDS("../rds/genome.detect.rds")
```

## Mean coverage

Import anvio genome mean coverage data
```{r}
genome.mean.cov <- read.delim("/mnt/storage/jocelyn/msc/anvio_analysis/ABBA/ABBA_PROFILE-SUMMARY/bins_across_samples/mean_coverage.txt", header = T, sep = "\t", dec = ".")
genome.mean.cov <- merge(genome.mean.cov, read.delim("/mnt/storage/jocelyn/msc/anvio_analysis/ACSA/ACSA_PROFILE-SUMMARY/bins_across_samples/mean_coverage.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.mean.cov <- merge(genome.mean.cov, read.delim("/mnt/storage/jocelyn/msc/anvio_analysis/COCO/COCO_PROFILE-SUMMARY/bins_across_samples/mean_coverage.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.mean.cov <- merge(genome.mean.cov, read.delim("/mnt/storage/jocelyn/msc/anvio_analysis/FAGR/FAGR_PROFILE-SUMMARY/bins_across_samples/mean_coverage.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.mean.cov <- merge(genome.mean.cov, read.delim("/mnt/storage/jocelyn/msc/anvio_analysis/THOC/THOC_PROFILE-SUMMARY/bins_across_samples/mean_coverage.txt", header = T, sep = "\t", dec = "."), by = "bins")
colnames(genome.mean.cov)[1] <- "species"
rownames(genome.mean.cov) <- genome.mean.cov$species
genome.mean.cov <- genome.mean.cov[,-1]
#saveRDS(genome.mean.cov, "../rds/genome.mean.cov.rds")
genome.mean.cov <- readRDS("../rds/genome.mean.cov.rds")
```

## Metadata
```{r}
metadata <- read.delim("../data_imported/samples_metadata.txt", header = TRUE, sep = "\t", dec = ".", row.names = 1)
metadata <- metadata %>% arrange(row.names(.))
metadata <- metadata %>% mutate(across(c(plant_clade,plant_order,plant_family,host_species,plot,tree,leaf_wash_batch,extraction_batch), as.factor))
metadata$host_species <- factor(x = metadata$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
metadata$shrub_tree <- c("tree","tree","tree","tree","tree","tree","tree","tree","tree","tree","shrub","shrub","shrub","shrub","shrub","tree","tree","tree","tree","tree","tree","tree","tree","tree","tree")
metadata$shrub_tree <- factor(x = metadata$shrub_tree, levels = c("shrub","tree"))
metadata$gymno_angio_tree <- c("gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree")
metadata$gymno_angio_tree <- factor(x = metadata$gymno_angio_tree, levels = c("gymnosperm_tree","angiosperm_tree","angiosperm_shrub"))
```

## Reference genomes table
```{r}
ref.genomes <- read.delim("../data_imported/meth_all_genomes.txt", header = TRUE, sep = "\t", dec = ",")
ref.genomes <- ref.genomes[,c(2:4)]
ref.genomes <- ref.genomes %>% mutate(across(c(species,strain_original_name,group), as.factor))
colnames(ref.genomes) <- c("Meth_species","strain","clade")
```

## Reads data
```{r}
reads.data <- read.delim("../data_imported/nb_reads.txt", header = TRUE, sep = "\t", dec = ",", row.names = 1)
```

## Phylogenetic tree
```{r}
meth.tree <- read.tree("/mnt/storage/jocelyn/msc/msc_detection_coverage/data_imported/RAxML_bipartitionsBranchLabels.Partition-Concatenated-genes-nov21-b=500-rooted")
# <- root(meth.tree, outgroup = c("YourOutgroupSpecies",), resolve.root = TRUE)
strains.list <- read.csv2("/mnt/storage/jocelyn/msc/msc_detection_coverage/data_imported/List-genomes.csv", sep = ";", dec = ".")
# rename tips of the tree - prepare data
st.sp <- strains.list[,c(1,2,4)]
colnames(st.sp) <- c("st","strain","Meth_species")
tip.labels <- data.frame(meth.tree[["tip.label"]])
colnames(tip.labels) <- "st"
st.sp <- st.sp[st.sp$st %in% tip.labels$st,]
tip.labels$st <- factor(tip.labels$st, levels = unique(tip.labels$st))
st.sp$st <- factor(st.sp$st, levels = unique(tip.labels$st))
st.sp <- st.sp[order(st.sp$st), ]
rownames(st.sp) <- seq_len(nrow(st.sp))
tip.labels <- cbind(tip.labels, st.sp$strain, st.sp$Meth_species)
colnames(tip.labels) <- c("st","strain","Meth_species")

# rename with strain's names
meth.tree[["tip.label"]] <- tip.labels$strain
meth.tree[["tip.label"]]

# rename with Methylo species
meth.tree.sp <- meth.tree
meth.tree.sp[["tip.label"]] <- tip.labels$Meth_species
meth.tree.sp[["tip.label"]] <- gsub("Meth", "M", meth.tree.sp[["tip.label"]])
meth.tree.sp[["tip.label"]]
```

Plot tree (verification test)
```{r}
plot(meth.tree, show.tip.label = T)
ggtree(meth.tree) + 
  theme_tree() +
  geom_tiplab(size=2)
ggsave("../figures/meth_tree_test.tiff", width = 8, height = 16)
```


################################################################################

# Dectection threshold

Statistics if I would keep genomes with detection threshold >= 0.5
```{r}
genome.detect_0.5 <- data.frame(ifelse(genome.detect < 0.5, 0, 1))
genome.detect_0.5 <- genome.detect_0.5[rowSums(genome.detect_0.5)>0,]
genome.detect_0.5 <- genome.detect_0.5[,colSums(genome.detect_0.5)>0]
```

Statistics
```{r}
length(rownames(genome.detect_0.5))
rowSums(genome.detect_0.5)
length(colnames(genome.detect_0.5))
colSums(genome.detect_0.5)
sum(genome.detect_0.5)
```
With detection threshold >= 0.5 : 20 species, in 23 samples, for a total of 133 detections.

Keep genomes with detection threshold >= 0.34
```{r}
genome.detect_0.34 <- data.frame(ifelse(genome.detect < 0.34, 0, 1))
genome.detect_0.34 <- genome.detect_0.34[rowSums(genome.detect_0.34)>0,]
genome.detect_0.34 <- genome.detect_0.34[,colSums(genome.detect_0.34)>0]
```

Statistics
```{r}
length(rownames(genome.detect_0.34))
rowSums(genome.detect_0.34)
length(colnames(genome.detect_0.34))
colSums(genome.detect_0.34)
sum(genome.detect_0.34)
```
With detection threshold >= 0.34 : 23 species, in 24 samples, for a total of 198 detections.

## PA community matrix
```{r}
genomes.PA.comm <- data.frame(t(genome.detect_0.34))
```


################################################################################

# Coverage, metadata and reads data associated with present genomes

## Mean coverage

```{r}
genomes.mean.cov_0.34 <- genome.mean.cov[rownames(genome.detect_0.34), colnames(genome.detect_0.34)]
genomes.mean.cov_0.34 <- as.data.frame(mapply(function(x, y) ifelse(x == 1, y, 0), genome.detect_0.34, genomes.mean.cov_0.34))
rownames(genomes.mean.cov_0.34) <- row.names(genome.detect_0.34)
genomes.cov.comm <- data.frame(t(genomes.mean.cov_0.34))
#saveRDS(genomes.cov.comm, "../rds/genomes.cov.comm.rds")
```

## Count data : transform mean coverage to count data
```{r}
genomes.cov.comm.c <- round(genomes.cov.comm, digits = 0)
#saveRDS(genomes.cov.comm.c, "../rds/genomes.cov.comm.c.rds")
```

## Metadata & add reads data

```{r}
metadata.comm <- metadata[rownames(genomes.cov.comm), ]
reads.data.comm <- reads.data[rownames(genomes.cov.comm), ]
metadata.comm$filtered_pairs <- reads.data.comm$filtered_pairs
```

## Richness and Shannon Index
```{r}
metadata.comm$richness <- rowSums(genomes.PA.comm)
metadata.comm$shannon <- vegan::diversity(genomes.cov.comm, index = "shannon")
```

################################################################################

# Rarefaction ?

REFLECTION ON RAREFACTION (paragraph copied from 'genomes_cov_detect.Rmd')
Try to answer the question: "should I rarefy for Methylobacterium species composition analyses?"
Whatever the answer, I think I can't rarefy on genome coverage values (even if transformed to count values - i.e. integer) because of my data structure : I would lose too much information - Meth species as well as samples, because many genomes are only detected once per sample, and a few are detected a lot in samples. Furthermore, there is a huge difference among my samples in terms of absolute abundances of detections (based on mean coverage).
Furthermore, normally rarefaction is applied on number of DNA sequences. Normally, in amplicon analyses, 1 sequence (1 PE read) is tied to 1 ASV. Here, thousands of sequences are tied to 1 genome, and even for genes, it takes a few (dozens?) sequences of 150bp to cover 1 whole gene (~1000bp). However, my gene data structure still allows me to treat genes as "sequences".

## Data structure
```{r}
rowSums(genomes.cov.comm.c)
rowSums(genomes.PA.comm)
```

## Curve
```{r}
rarecurve(genomes.cov.comm.c, step = 1, label = T, xlim=c(0,150))
rarecurve(genomes.cov.comm.c, step = 1, label = T, xlim=c(0,50))
```

## Correlation between richness and number of filtered pairs.

### Distributions
```{r}
hist(metadata.comm$richness, breaks = 25)
shapiro.test(metadata.comm$richness)
hist(metadata.comm$filtered_pairs, breaks = 25)
shapiro.test(metadata.comm$filtered_pairs)
hist(reads.data$filtered_pairs, breaks = 25)
shapiro.test(reads.data$filtered_pairs)
```
The two variables distributions are normal.

### Models

Linear
```{r}
lm <- lm(richness ~ filtered_pairs, data = metadata.comm)
plot(richness ~ filtered_pairs, data = metadata.comm)
summary(lm)
anova(lm)
plot(lm)
hist(residuals(lm), breaks = 25)
shapiro.test(residuals(lm))
```
The model's assumptions are met, there is no correlation between the richness of samples and their number of filtered read pairs.


Fist, due to my data structure, I cannot rarefy my Methylobacterium species data without losing a huge amount of information.

However, not rarefying might not be a problem considering
A) There is no correlation between richness and number of filtered read pairs.
AB) There is no correlation between mapped reads to Methylo genomes and number of filtered read pairs. See script 'mapped_reads.Rmd'.
C) For some Methylobacterium species, there is a correlation between genome detection and number of filtered reads (sequencing depth). However, the significative correlations are all for species whose genomes were not detected in any samples (count = 0) (detection threshold >= 0.34). See script 'genomes_cov_and_detect.Rmd'.
D) For some Methylobacterium species, there is a correlation between mean genome coverage and number of filtered reads (sequencing depth). However, the significative correlations are all for species whose genomes were in average poorly covered across all samples. See script 'genomes_cov_and_detect.Rmd'.

## Conclusion : I will not rarefy my Methylobacterium species data.


################################################################################

# Prevalence of species across samples

Which species is present in the most samples (most prevalent species in the forest)?

## Coverage
```{r}
species.cov.across <- as.data.frame(rowSums(t(genomes.cov.comm)))
species.cov.across$species <- as.factor(rownames(species.cov.across))
colnames(species.cov.across) <- c("cov.across", "species")
species.cov.across <- merge(species.cov.across, ref.genomes, by = "species")
species.cov.across$isolated_from <- c("Cerastium holosteoides","Haplocladium microphyllum","Arabidopsis
thaliana","Lolium prenne","Pseudosasa japonica","Fagus grandifolia","Fagus grandifolia","Arabidopsis
thaliana","Arabidopsis thaliana","Arabidopsis thaliana","Acer saccharum","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Acer saccharum","Fagus grandifolia","Acer saccharum","Fagus grandifolia","Cerastium holosteoides","Trifolium repens")
```
Plot
```{r}
ggplot(species.cov.across, aes(x = species, y = cov.across))+
  geom_bar(stat = "identity", fill = "black") +
  labs(x = "Methylobacterium species", y = "Coverage across samples") +
  ylim(0,400) +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 11),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/Methylo_species_cov_across_samples.tiff", width = 6, height = 4)
```

## PA
```{r}
species.PA.across <- as.data.frame(rowSums(t(genomes.PA.comm)))
species.PA.across$species <- as.factor(rownames(species.PA.across))
colnames(species.PA.across) <- c("presence.across", "species")
species.PA.across <- merge(species.PA.across, ref.genomes, by = "species")
species.PA.across$isolated_from <- c("Cerastium holosteoides","Haplocladium microphyllum","Arabidopsis
thaliana","Lolium prenne","Pseudosasa japonica","Fagus grandifolia","Fagus grandifolia","Arabidopsis
thaliana","Arabidopsis thaliana","Arabidopsis thaliana","Acer saccharum","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Acer saccharum","Fagus grandifolia","Acer saccharum","Fagus grandifolia","Cerastium holosteoides","Trifolium repens")
```
Plot
```{r}
ggplot(species.PA.across, aes(x = species, y = presence.across))+
  geom_bar(stat = "identity", fill = "black") +
  labs(x = "Methylobacterium species", y = "Number of samples") +
  ylim(0,25) +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 11),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/Methylo_species_PA_across_samples.tiff", width = 6, height = 4)
```

```{r}
hist(species.PA.across$presence.across, breaks = 50)
```


################################################################################

# Alpha diversity

## Richness

```{r}
lm <- lm(richness ~ host_species, data = metadata.comm)
plot(richness ~ host_species, data = metadata.comm)
summary(lm)
anova(lm)
plot(lm)
hist(residuals(lm))
shapiro.test(residuals(lm))
```

Post-hoc
```{r}
# Shrub vs. trees

## Student t-test assumptions : normal distribution of data and equal variance
hist(metadata.comm[metadata.comm$shrub_tree == "tree",]$richness, breaks = 25)
shapiro.test(metadata.comm[metadata.comm$shrub_tree == "tree",]$richness)
hist(metadata.comm[metadata.comm$shrub_tree == "shrub",]$richness, breaks = 25)
shapiro.test(metadata.comm[metadata.comm$shrub_tree == "shrub",]$richness)
var(metadata.comm[metadata.comm$shrub_tree == "tree",]$richness)
var(metadata.comm[metadata.comm$shrub_tree == "shrub",]$richness)
## unequal variance so welch t-test
t1 <- t.test(richness ~ shrub_tree, data = metadata.comm, var.equal = F)

# Gymno vs. angio trees

## Student t-test assumptions : normal distribution of data and equal variance
hist(metadata.comm[metadata.comm$gymno_angio_tree == "gymnosperm_tree",]$richness, breaks = 25)
shapiro.test(metadata.comm[metadata.comm$gymno_angio_tree == "gymnosperm_tree",]$richness)
hist(metadata.comm[metadata.comm$gymno_angio_tree == "angiosperm_tree",]$richness, breaks = 25)
shapiro.test(metadata.comm[metadata.comm$gymno_angio_tree == "angiosperm_tree",]$richness)
var(metadata.comm[metadata.comm$gymno_angio_tree == "gymnosperm_tree",]$richness)
var(metadata.comm[metadata.comm$gymno_angio_tree == "angiosperm_tree",]$richness)
## unequal variance so welch t-test
t2 <- t.test(richness ~ gymno_angio_tree, data = metadata.comm[c(1:9,15:24),], var.equal = F)

# p-values adjustement
p.adjust(c(t1[["p.value"]],t2[["p.value"]]), method = "bonferroni")
```

Plot
```{r}
ggplot(metadata.comm, aes(x = host_species, y = richness)) +
geom_boxplot(alpha = 1, aes(fill = host_species), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = host_species), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
scale_fill_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"))+
scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Host Species", 
     y = "Methylobacterium species richness")+
scale_x_discrete(labels = c("ABBA" = "A. balsamea", "THOC" = "T. occidentalis","ACSA" = "A. Saccharum", "FAGR" = "F. grandifolia", "COCO" = "C. cornuta"))+
ylim(0,25) +
theme_bw()+
theme(text = element_text(family = "Arial"),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0, face = "italic"),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/Methylo_species_richness.tiff", width = 6, height = 4)
```

## Shannon index

```{r}
hist(metadata.comm$shannon, breaks = 25)
shapiro.test(metadata.comm$shannon)
# Non normal distribution, so will use Kruskall-Wallis
kruskal.test(shannon ~ host_species, data = metadata.comm)
rstatix::kruskal_effsize(shannon ~ host_species, data = metadata.comm)
# Can't use parametric ANOVA below
summary(lm(shannon ~ host_species, data = metadata.comm))
anova(lm(shannon ~ host_species, data = metadata.comm))
```

Plot
```{r}
ggplot(metadata.comm, aes(x = host_species, y = shannon)) +
geom_boxplot(alpha = 1, aes(fill = host_species), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = host_species), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
scale_fill_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"))+
scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Host Species", 
     y = "Shannon Index")+
scale_x_discrete(labels = c("ABBA" = "A. balsamea", "THOC" = "T. occidentalis","ACSA" = "A. Saccharum", "FAGR" = "F. grandifolia", "COCO" = "C. cornuta"))+
ylim(-0.1,2.5) +
theme_bw()+
theme(text = element_text(family = "Arial"),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0, face = "italic"),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/Methylo_species_shannon.tiff", width = 6, height = 4)
```


## Pielou Index

Calculation
```{r}
metadata.comm$pielou <- metadata.comm$shannon/log(metadata.comm$richness)
hist(metadata.comm$pielou, breaks = 25)
shapiro.test(metadata.comm$pielou)
```

Model
```{r}
lm <- lm(pielou ~ host_species, metadata.comm)
plot(pielou ~ host_species, metadata.comm)
summary(lm)
anova(lm)
hist(residuals(lm))
shapiro.test(residuals(lm))
```

Plot
```{r}
ggplot(metadata.comm, aes(x = host_species, y = pielou)) +
geom_boxplot(alpha = 1, aes(fill = host_species), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = host_species), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
scale_fill_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"))+
scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Host Species", 
     y = "Pielou Index")+
scale_x_discrete(labels = c("ABBA" = "A. balsamea", "THOC" = "T. occidentalis","ACSA" = "A. Saccharum", "FAGR" = "F. grandifolia", "COCO" = "C. cornuta"))+
ylim(0.5,1) +
theme_bw()+
theme(text = element_text(family = "Arial"),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0, face = "italic"),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/Methylo_species_pielou.tiff", width = 6, height = 4)
```

Correlation between richness and evenness
```{r}
lm <- lm(pielou ~ richness, metadata.comm)
plot(pielou ~ richness, metadata.comm)
summary(lm)
anova(lm)
hist(residuals(lm))
shapiro.test(residuals(lm))
```


################################################################################

# Beta diversity

## Composition

### Barplot
```{r}
abundances <- picante::matrix2sample(genomes.cov.comm)
colnames(abundances) <- c("sample","abund","Meth_species")
abundances$sample <- factor(x = abundances$sample, levels =c ("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M18","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
abundances$Meth_species <- factor(x = abundances$Meth_species, 
                                  levels = c(
"M_sp_026",
"M_sp_024",
"M_sp_029",
"M_pseudosasicola",
"M_sp_030",
"M_mesophilicum",
"M_sp_021",
"M_sp_027",
"M_sp_023",
"M_sp_022",
"M_longum",
"M_sp_025",
"M_sp_034",
"M_sp_018",
"M_cerastii",
"M_trifolii",
"M_sp_010",
"M_sp_008",
"M_sp_003",
"M_sp_011",
"M_sp_004",
"M_thuringiense",
"M_haplocladii"
))


palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
"#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888",
"#EE7744",  # Complement of #88CCEE
"#77CC99",  # Complement of #CC6677
"#7733CC",  # Complement of #DDCC77
"#CC88EE",  # Complement of #117733
"#88DD77",  # Complement of #332288
"#44BB66",  # Complement of #AA4499
"#BB4455",  # Complement of #44AA99
"#336699",  # Complement of #999933
"#55AA00",  # Complement of #882255
"#99EEFF",  # Complement of #661100
"#CC3366",  # Complement of #6699CC
"#777777")   # Complement of #888888


# Composition with abundances
ggplot(abundances, aes(x = sample, y = abund, fill = Meth_species)) +
  geom_bar(stat = "identity",
#          position = "fill" # add to get relative abundances
  ) +
  labs(x = "Sample", y = "Genome coverage") +
  scale_fill_manual(values=c(
  "#C2B280",
  "#1A8A00",
  "#8E6A48",
    "#B5D192",  

  "#3B5B29",
  "#E1B05D",  
    "#0095C8",
  "#B66D3B",  

  "#4E8",
  "#A77B3D",
  "#1F8C8B",
  "#E5B5B2", 
  "#B7C9D9", 
  "#E15F4C", 
  "#FFCC00",
  "#008B5B",
  "#7C9DB8", 
    "#D9C3B2",
  "#5A9C6A", 
    "#F4A300",

  "#C4A65D",
  "#A6C8C4", 
      "#4B8C2A" 

), name = "Methylobacterium species") +
  guides(fill = guide_legend(ncol = 1)) +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      legend.justification = c(0.5, 1),
      legend.text = element_text(size = 10.5),
      legend.key.size = unit(0.43, "cm"),
      legend.position = "right", 
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 11),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/Methylo_species_barplot-241220.tiff", width = 8, height = 5.5)
```


*** DO NOT RUN 2 CHUNKS BELOW ***
Baselga partition method - with 'betapart' package
Sorensen distance matrix
```{r}
# Compute 3 distance matrices
genomes.sor <- beta.pair(genomes.PA.comm, index.family = "sorensen")

write.table(t(genomes.PA.comm), file = "../genomes_PA_comm.csv", sep = ",", quote = TRUE)

# Proportion of turnover on total betadiv
matrix.sor.sim <- data.frame(as.matrix(genomes.sor[["beta.sim"]]))
#matrix.sor.sim <- data.frame(mapply(function(x) ifelse(x == 0, 0.000000000000001, x), matrix.sor.sim))
matrix.sor <- data.frame(as.matrix(genomes.sor[["beta.sor"]]))
rownames(matrix.sor.sim) <- rownames(matrix.sor)

sim.sor <- matrix.sor.sim/matrix.sor
sim.sor.2 <- picante::matrix2sample(sim.sor)
#sim.sor.2 <- sim.sor.2[sim.sor.2$abund!=Inf,]
#sim.sor.2$abund <- ifelse(sim.sor.2$abund < 0.001, 0, sim.sor.2$abund)
colnames(sim.sor.2)[2] <- "turnover_betadiv"

sim.sor.2$pairs <- paste0(gsub("_M[0-9]{2}", "", sim.sor.2$plot), "-", gsub("_M[0-9]{2}", "", sim.sor.2$id))
sim.sor.2$pairs <- as.factor(sim.sor.2$pairs)
#sim.sor.2 <- sim.sor.2[!gsub("-.*", "", sim.sor.2$pairs) == gsub(".*-", "", sim.sor.2$pairs), ]
sim.sor.2 <- sim.sor.2 %>% filter(pairs %in% c("ABBA-ACSA","ABBA-COCO","ABBA-FAGR","ABBA-THOC","ACSA-COCO","ACSA-FAGR","ACSA-THOC","COCO-FAGR","COCO-THOC","FAGR-THOC"))
sim.sor.2$pairs <- droplevels(sim.sor.2$pairs)

# Contribution of nestedness between species
mean(sim.sor.2$turnover_betadiv)
#mean(sim.sor.2$turnover_betadiv[sim.sor.2$turnover_betadiv!=0])

####

# Proportion of nestedness on total betadiv
matrix.sor.sne <- data.frame(as.matrix(genomes.sor[["beta.sne"]]))
matrix.sor.sne <- data.frame(mapply(function(x) ifelse(x == 0, 0.000000000000001, x), matrix.sor.sne))
rownames(matrix.sor.sne) <- rownames(matrix.sor)

sne.sor <- matrix.sor.sne/matrix.sor
sne.sor.2 <- picante::matrix2sample(sne.sor)
sne.sor.2 <- sne.sor.2[sne.sor.2$abund!=Inf,]
sne.sor.2$abund <- ifelse(sne.sor.2$abund < 0.001, 0, sne.sor.2$abund)
colnames(sne.sor.2)[2] <- "nestedness_betadiv"

sne.sor.2$pairs <- paste0(gsub("_M[0-9]{2}", "", sne.sor.2$plot), "-", gsub("_M[0-9]{2}", "", sne.sor.2$id))
sne.sor.2$pairs <- as.factor(sne.sor.2$pairs)
#sne.sor.2 <- sne.sor.2[!gsub("-.*", "", sne.sor.2$pairs) == gsub(".*-", "", sne.sor.2$pairs), ]
sne.sor.2 <- sne.sor.2 %>% filter(pairs %in% c("ABBA-ACSA","ABBA-COCO","ABBA-FAGR","ABBA-THOC","ACSA-COCO","ACSA-FAGR","ACSA-THOC","COCO-FAGR","COCO-THOC","FAGR-THOC"))
sne.sor.2$pairs <- droplevels(sne.sor.2$pairs)

# Contribution of nestedness between species
mean(sne.sor.2$nestedness_betadiv)

# Overall partitioning of turnover and nestedness
genomes.sor.multi <- beta.multi(genomes.PA.comm, index.family = "sorensen")

# Partition the turnover and nestedness within samples of same host species
genomes.sor.ABBA <- data.frame(beta.multi(genomes.PA.comm[c(1:5),], index.family = "sorensen"))
genomes.sor.ACSA <- data.frame(beta.multi(genomes.PA.comm[c(6:9),], index.family = "sorensen"))
genomes.sor.COCO <- data.frame(beta.multi(genomes.PA.comm[c(10:14),], index.family = "sorensen"))
genomes.sor.FAGR <- data.frame(beta.multi(genomes.PA.comm[c(15:19),], index.family = "sorensen"))
genomes.sor.THOC <- data.frame(beta.multi(genomes.PA.comm[c(20:24),], index.family = "sorensen"))
genomes.sor.multi.hosts <- rbind(genomes.sor.ABBA,genomes.sor.ACSA,genomes.sor.COCO,genomes.sor.FAGR,genomes.sor.THOC)
genomes.sor.multi.hosts$host_species <- c("ABBA","ACSA","COCO","FAGR","THOC")
rm(genomes.sor.ABBA,genomes.sor.ACSA,genomes.sor.COCO,genomes.sor.FAGR,genomes.sor.THOC)

mean(genomes.sor.multi.hosts$beta.SIM)
mean(genomes.sor.multi.hosts$beta.SNE)
mean(genomes.sor.multi.hosts$beta.SOR)
```
Plot turnover/betadiv
```{r}
ggplot(sim.sor.2, aes(x = pairs, y = turnover_betadiv)) +
#geom_boxplot(alpha = 1, aes(fill = NA), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = NA), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
#scale_fill_manual(values=c("#63AFF9","#CA0E78","#6104BB","#F39635","#1EA670"))+
#scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Pairs", 
     y = "Turnover contribution to beta diversity (proportion)")+
ylim(0,1) +
theme_bw()+
theme(text = element_text(family = "Arial"),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
#ggsave("../figures/Methylo_species_richness.tiff", width = 6, height = 4)
```
*** DO NOT RUN 2 CHUNKS ABOVE ***


## Beta diversity partitioning

### Bray-Curtis distance partitioning

Use of Baselga Bray-Curtis on quantitative matrix
```{r}
genomes.cov.beta <- adespatial::beta.div.comp(genomes.cov.comm, coef = "BS", quant = TRUE)
genomes.cov.beta[["part"]]
genomes.bray.dist <- genomes.cov.beta[["D"]]
genomes.bray <- data.frame(as.matrix(genomes.cov.beta[["D"]]))

genomes.cov.beta.host <- data.frame(genomes.cov.beta[["part"]])
colnames(genomes.cov.beta.host) <- "All samples"
genomes.cov.beta.ABBA <- adespatial::beta.div.comp(genomes.cov.comm[c(1:5),], coef = "BS", quant = TRUE)
genomes.cov.beta.host$ABBA <- genomes.cov.beta.ABBA[["part"]]
genomes.cov.beta.ACSA <- adespatial::beta.div.comp(genomes.cov.comm[c(6:9),], coef = "BS", quant = TRUE)
genomes.cov.beta.host$ACSA <- genomes.cov.beta.ACSA[["part"]]
genomes.cov.beta.COCO <- adespatial::beta.div.comp(genomes.cov.comm[c(10:14),], coef = "BS", quant = TRUE)
genomes.cov.beta.host$COCO <- genomes.cov.beta.COCO[["part"]]
genomes.cov.beta.FAGR <- adespatial::beta.div.comp(genomes.cov.comm[c(15:19),], coef = "BS", quant = TRUE)
genomes.cov.beta.host$FAGR <- genomes.cov.beta.FAGR[["part"]]
genomes.cov.beta.THOC <- adespatial::beta.div.comp(genomes.cov.comm[c(20:24),], coef = "BS", quant = TRUE)
genomes.cov.beta.host$THOC <- genomes.cov.beta.THOC[["part"]]
#write.csv(t(round(genomes.cov.beta.host, digits = 4)), "../genomes.cov.beta.csv")

# Proportion of turnover on total betadiv for pairs of different hosts
matrix.repl <- data.frame(as.matrix(genomes.cov.beta[["repl"]]))

repl.D <- matrix.repl/genomes.bray
repl.D <- picante::matrix2sample(repl.D)
colnames(repl.D)[2] <- "turnover_betadiv"

repl.D$pairs <- paste0(gsub("_M[0-9]{2}", "", repl.D$plot), "-", gsub("_M[0-9]{2}", "", repl.D$id))
repl.D$pairs <- as.factor(repl.D$pairs)
repl.D <- repl.D %>% filter(pairs %in% c("ABBA-ACSA","ABBA-COCO","ABBA-FAGR","ABBA-THOC","ACSA-COCO","ACSA-FAGR","ACSA-THOC","COCO-FAGR","COCO-THOC","FAGR-THOC"))
repl.D$pairs <- droplevels(repl.D$pairs)

# Contribution of nestedness between species
mean(repl.D$turnover_betadiv)
```
Plot turnover/betadiv
```{r}
ggplot(repl.D, aes(x = pairs, y = turnover_betadiv)) +
#geom_boxplot(alpha = 1, aes(fill = NA), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = NA), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
#scale_fill_manual(values=c("#63AFF9","#CA0E78","#6104BB","#F39635","#1EA670"))+
#scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Pairs", 
     y = "Turnover contribution to beta diversity (proportion)")+
ylim(0,1) +
theme_bw()+
theme(text = element_text(family = "Arial"),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
```

## PERMANOVA
```{r}
genomes.bray.perm <- adonis2(genomes.bray.dist ~ host_species, data = metadata.comm)
genomes.bray.perm
```

## BETADISPER
```{r}
genomes.bray.disper <- betadisper(genomes.bray.dist, metadata.comm$host_species)
summary(genomes.bray.disper)
anova(genomes.bray.disper)
plot(genomes.bray.disper)
```

## Pairwise PERMANOVAs
```{r}
shrub_tree_PERM <- adonis2(genomes.bray.dist ~ shrub_tree, data = metadata.comm)
shrub_tree_PERM
shrub_tree_DISP <- betadisper(genomes.bray.dist, metadata.comm$shrub_tree)
anova(shrub_tree_DISP)
plot(shrub_tree_DISP)

genomes.bray.dist.noCOCO <- data.frame(as.matrix(genomes.bray.dist))
genomes.bray.dist.noCOCO <- genomes.bray.dist.noCOCO[c(1:9,15:24),c(1:9,15:24)]
genomes.bray.dist.noCOCO <- as.dist(as.matrix(genomes.bray.dist.noCOCO), diag = FALSE, upper = FALSE)
gymno_angio_tree_PERM <- adonis2(genomes.bray.dist.noCOCO ~ gymno_angio_tree, data = metadata.comm[c(1:9,15:24),], perm =999)
gymno_angio_tree_PERM
gymno_angio_tree_DISP <- betadisper(genomes.bray.dist.noCOCO, metadata.comm[c(1:9,15:24),]$gymno_angio_tree)
anova(gymno_angio_tree_DISP)
plot(gymno_angio_tree_DISP, label = F)
```


## PCoA
```{r}
genomes.pcoa <- cmdscale(genomes.bray.dist, k =(nrow(genomes.cov.comm) - 1), eig = TRUE)
pcoa.eig <- eigenvals(genomes.pcoa) # extract eigenvalues
pcoa.eig <- abs(pcoa.eig) # absolute values
sum.pcoa.eig <- sum(pcoa.eig) # total of eigenvalues
pcoa.prop.exp <- pcoa.eig/sum.pcoa.eig # proportion explained by each axis
(pcoa.axis1 <- pcoa.prop.exp[1])
(pcoa.axis2 <- pcoa.prop.exp[2])
site.scores.pcoa <- scores(genomes.pcoa) # site scores
sp.scores.pcoa <- wascores(genomes.pcoa$points, genomes.cov.comm) # species scores 
site.scores.pcoa <- data.frame(site.scores.pcoa)
sp.scores.pcoa <- data.frame(sp.scores.pcoa)
host_species <- metadata.comm$host_species
site.scores.pcoa <- cbind(site.scores.pcoa, host_species)
pcoa.ggordiplot <- gg_ordiplot(genomes.pcoa, choices = c(1,2), groups = metadata.comm$host_species, kind = "sd", conf = 0.95, pt.size = 3)
pcoa.ellipses <- pcoa.ggordiplot$df_ellipse # extract confidence ellipse
```
Plot
```{r}
ggplot(data = site.scores.pcoa, 
         aes(x = Dim1, y = Dim2), 
         color = host_species) + 
  expand_limits(x=c(-0.8,0.8), y=c(-0.8,0.8)) +
  geom_point(aes(shape = host_species, color = host_species), show.legend = T, 
             size=3) +
  xlab(paste("PCoA1 (",round(pcoa.axis1*100, digits = 2),"%)", sep = "")) +
  ylab(paste("PCoA2 (",round(pcoa.axis2*100, digits = 2),"%)", sep = "")) +
  #geom_text_repel(data = sp.scores.pcoa, 
  #                aes(x = X1, y = X2), 
  #                label=rownames(sp.scores.pcoa), 
  #                color = "black") +
  scale_shape_manual(values = c(16,16,16,16,16)) +
  guides(shape = "none") +
  geom_path(data = pcoa.ellipses, 
            aes(x = x, y = y, color = Group), 
            show.legend = FALSE) + 
  scale_color_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"), name = "Host species", labels = c("ABBA" = expression(italic("A. balsamea")), "THOC" = expression(italic("T. occidentalis")),"ACSA" = expression(italic("A. Saccharum")), "FAGR" = expression(italic("F. grandifolia")), "COCO" = expression(italic("C. cornuta")))) +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.text = element_text(size = 11),
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/genomes_pcoa.tiff", width = 6, height = 4)
```


################################################################################

# Phylogeny

NOTE : To start section, use 'genomes.cov.comm','genomes.bray.dist','metadata.comm','meth.tree' and 'meth.tree.sp'.

## Replace 'genomes.cov.comm' columns names with strain instead of species.
```{r}
genomes.cov.comm.st <- genomes.cov.comm
ref.genomes.sub <- ref.genomes[ref.genomes$Meth_species %in% colnames(genomes.cov.comm.st),]
colnames(genomes.cov.comm.st) <- ref.genomes.sub$strain
```

## Prune tree
```{r}
pruned.tree <- prune.sample(genomes.cov.comm.st, meth.tree)
#pruned.tree.sp <- prune.sample(genomes.cov.comm, meth.tree.sp)
plot(pruned.tree, show.tip.label = TRUE)
nodelabels()
#plot(pruned.tree.sp, show.tip.label = TRUE)
```

## Test if Methylobacterium communities were composed of phylogenetically related species

i.e. Does communities harbored species more closely related than expected by chance.

### Cophenetic distance
```{r}
# Calculate the interspecific phylogenetic distance matrix (cophenetic distance)
# i.e. how species are phylogeneticaly related (close or far)
phydist <- cophenetic(pruned.tree) # same distances on pruned vs not pruned tree
```

### MPD : Mean pairwise distance (these mpd results are included in the ses.mdp results)
```{r}
mpd.results <- data.frame(matrix(ncol = 0, nrow = 24))
rownames(mpd.results) <- rownames(genomes.cov.comm.st)
mpd.results$mpd <- mpd(genomes.cov.comm.st, phydist, abundance.weighted = F)
mpd.results$mpd.abund <- mpd(genomes.cov.comm.st, phydist, abundance.weighted = T)
mpd.results$mntd <- mntd(genomes.cov.comm.st, phydist, abundance.weighted = F)
mpd.results$mntd.abund <- mntd(genomes.cov.comm.st, phydist, abundance.weighted = T)
```

### SES.MPD : Standardized Effect Size MPD
```{r}
# PRESENCE-ABSENCE
ses.mpd.results <- ses.mpd(genomes.cov.comm.st, phydist, null.model = "taxa.labels", abundance.weighted = FALSE, runs = 999, iterations = 1000)
ses.mpd.results$host_species <- metadata.comm$host_species
ses.mpd.results$plant_clade <- metadata.comm$plant_clade

ses.mpd.results$mpd.obs <- round(ses.mpd.results$mpd.obs, digits = 4)
ses.mpd.results$mpd.rand.mean <- round(ses.mpd.results$mpd.rand.mean, digits = 4)
ses.mpd.results$mpd.rand.sd <- round(ses.mpd.results$mpd.rand.sd, digits = 4)
ses.mpd.results$mpd.obs.z <- round(ses.mpd.results$mpd.obs.z, digits = 4)
#saveRDS(ses.mpd.results, "../rds/ses.mpd.results.rds")
write_excel_csv(ses.mpd.results, "../ses.mpd.results.csv")
ses.mpd.results <- readRDS("../rds/ses.mpd.results.rds")

plot(mpd.obs.z ~ host_species, ses.mpd.results)
plot(mpd.obs.z ~ plant_clade, ses.mpd.results)
mean(na.omit(ses.mpd.results$mpd.obs.z))
sd(na.omit(ses.mpd.results$mpd.obs.z))


# ABUNDANCES 
ses.mpd.results.abund <- ses.mpd(genomes.cov.comm.st, phydist, null.model = "taxa.labels", abundance.weighted = TRUE, runs = 999, iterations = 1000)
ses.mpd.results.abund$host_species <- metadata.comm$host_species
plot(mpd.obs.z ~ host_species, ses.mpd.results.abund)
```
NOTE : I think it makes more sense to use non abundance-weighted values, since we don't want abundances differences to mask phylogenetic effects. (changing to relative abundances instead ff 'true' abundances in community matrix do not significantly change the results). Since E-025 is super abundant in some COCO samples, it might distort phylogenetic reality (communities appears as less phylogenetically clustered)

## Do some host species harbors communities composed of more closely related species than other hosts?

We compare the z-scores of observed MPD values among host species with an ANOVA to test if host species was having an effect on the phylogenetic relatedness of communities. 

### ANOVA
```{r}
lm <- lm(mpd.obs.z ~ host_species, ses.mpd.results)
summary(lm)
anova(lm)
plot(lm)
hist(residuals(lm))
shapiro.test(residuals(lm))
```

CONCLUSION : Overall, Methylobacterium species communities appear to be phylogenetically clustered (species are more related to one another in communities than they would be expected by pure chance). BUT, based on non abundance weithed data, there is no difference in ses.mpd between host species.

### Plot
```{r}
ggplot(ses.mpd.results, aes(x = host_species, y = mpd.obs.z)) +
geom_boxplot(alpha = 1, aes(fill = host_species), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = host_species), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
scale_fill_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"))+
scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Host Species", 
     y = "Standardized effect size of MPD")+
scale_x_discrete(labels = c("ABBA" = "A. balsamea", "THOC" = "T. occidentalis","ACSA" = "A. Saccharum", "FAGR" = "F. grandifolia", "COCO" = "C. cornuta"))+
ylim(-6,1) +
theme_bw()+
theme(text = element_text(family = "Arial"),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0, face = "italic"),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/Methylo_SESMPD.tiff", width = 6, height = 4)
```


## Do host species harbor distinct phylogenetic groups of Methylobacterium species?


### Phylogenetic beta diversity
```{r}
# calculate inter-community mean pairwise distance (MPD)
comm.dist <- comdist(genomes.cov.comm.st, phydist) # Object 'comm.dist' used to be named 'comdist.result'
```

### Cluster
```{r}
comm.dist.clusters <- hclust(comm.dist)
plot(comm.dist.clusters)
```
### Mantel test (correlation between : MPD between pairs of communities & Bray-Curtis distance (taxonomic composition))
```{r}
mantel(genomes.bray.dist, comm.dist)
```
### PERMANOVA
```{r}
phy.perm <- adonis2(comm.dist ~ host_species, data = metadata.comm)
phy.perm
```

### BETADISPER
```{r}
phy.disper <- betadisper(comm.dist, metadata.comm$host_species)
anova(phy.disper)
plot(phy.disper)
```


### Conclusion :
There seems to be no clear patterns of phylogenetic relatedness among communities.


### Visualize tree with heatmap of abundance

Prepare data
```{r}
# https://thackl.github.io/ggtree-composite-plots
abundances <- picante::matrix2sample(genomes.cov.comm.st)
colnames(abundances) <- c("sample","abund","Meth_species")
abundances$sample <- factor(x = abundances$sample, levels =c ("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M18","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
```



```{r}
# Plot tree with node labels to get info (exploration stage)
gg_tr <- ggtree(pruned.tree) + geom_tiplab(align=TRUE) +
  scale_x_continuous(expand=expansion(0.5)) + # make more room for the labels
  geom_text(aes(label = node), vjust = -0.5, hjust = 1.2, size = 4, color = "red", fontface = "italic")
gg_tr

# Create dataframe with clades and associated colors 
clades <- data.frame(
  node = c(1,2,25,19,20,21,22,23,42,43,44,45,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,24,26),  # Nodes for different clades
  color = c("darkorange","darkorange","black", "red","red","red","red","red","black","red","red","red", "blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","black","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","black","black"),
  clade = c("Clade B","Clade B","none", "Clade D","Clade D","Clade D","Clade D","Clade D","none","Clade D","Clade D","Clade D", "Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","none","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A", "none","none")
)

# Plot clade-colored tree with strains as tip labels
gg_tr <- ggtree(pruned.tree, aes(color = as.factor(node))) + geom_tiplab(align=TRUE, color="black", size = 4.25) +
  scale_x_continuous(expand=expansion(0.5)) +
  scale_color_manual(values = clades$color, 
                     breaks = clades$node,
                     labels = clades$clade)+
  theme(legend.position="none",
        text = element_text(family = "Arial"))
gg_tr

# Get tip order (for data arrangment)
tip_order <- paste("", 
        paste0(get_taxa_name(gg_tr), collapse=","))
tip_order

#abundances$Meth_species <- factor(x = abundances$Meth_species, levels = #c("M_sp_021","M_sp_027","M_sp_023","M_sp_022","M_sp_025","M_longum","M_pseudosasicola","M_sp_029","M_sp_030","M_sp_#026","M_sp_024","M_mesophilicum","M_sp_034","M_sp_018","M_cerastii","M_trifolii","M_haplocladii","M_thuringiense","#M_sp_010","M_sp_008","M_sp_003","M_sp_004","M_sp_011"))

# Rearrange factor levels
abundances$Meth_species <- factor(x = abundances$Meth_species, levels = c("J-077","J-048","E-066","BL36","J-001","NBRC15688T","E-065","J-088","J-030","J-026","DSM23933T","J-070","J-076","E-025","DSM23679T","DSM23632","Leaf125","Leaf89","J-078","Leaf99","J-090","DSM23674","NBRC107714"))

# Plot the heatmap with strains in y axis
gg_heat <- ggplot(abundances, aes(x = sample, y = Meth_species)) + 
  geom_tile(aes(fill=abund)) + 
 # geom_tile(aes(fill=sqrt(abund))) + # heat transformed to square root coverage value
  scale_fill_gradient2(high= "darkgreen", name = "Coverage"
                      ,transform = "sqrt"
                      )+ 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Sample") +
    theme_bw() +
      theme(
      text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      panel.background = element_rect(fill = "black"),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 12),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )

(gg_tr + gg_heat) # Colored tree with strains name on both figures

# Note : Order of strains is ok, let's change names of strain to species on heatmap.

#gg_tr.nolabs <- ggtree(pruned.tree) + geom_tiplab(align=TRUE, size = 0) + 
#  scale_x_continuous(expand=expansion(0.5)) # make more room for the labels
#gg_tr.nolabs

gg_heat.sp <- gg_heat +
              theme(axis.title.y = element_blank()) +
              scale_y_discrete(labels = c(
              "J-077" = "M_sp_026",
              "J-048" = "M_sp_024",
              "E-066" = "M_sp_029",
              "BL36" = "M_pseudosasicola",
              "J-001" = "M_sp_030",
              "NBRC15688T" = "M_mesophilicum",
              "E-065" = "M_sp_021",
              "J-088" = "M_sp_027",
              "J-030" = "M_sp_023",
              "J-026" = "M_sp_022",
              "DSM23933T" = "M_longum",
              "J-070" = "M_sp_025",
              "J-076" = "M_sp_034",
              "E-025" = "M_sp_018",
              "DSM23679T" = "M_cerastii",
              "DSM23632" = "M_trifolii",
              "Leaf125" = "M_sp_010",
              "Leaf89" = "M_sp_008",
              "J-078" = "M_sp_003",
              "Leaf99" = "M_sp_011",
              "J-090" = "M_sp_004",
              "DSM23674" = "M_thuringiense",
              "NBRC107714" = "M_haplocladii"
              ),
              limits = rev(levels(abundances$Meth_species)))
gg_heat.sp

#(gg_tr + gg_heat)
#(gg_tr.nolabs + gg_heat.sp)
plot.heat.tree <- gg_tr + gg_heat.sp
plot.heat.tree

ggsave("../figures/Methylobacterium_tree_heatmap.tiff", width = 12, height = 6)
```






################################################################################
DO NOT RUN ALL CODE BELOW, KEPT AS ARCHIVE.



dbRDA
```{r}
dbrda.detect <- dbrda(detect.bray ~ host_species, data = metadata.comm)
anova.cca(dbrda.detect, by = "terms", permutations = 999)
RsquareAdj(dbrda.detect)
```
Extract proportions explained by axes 1 and 2
```{r}
dbrda.detect.eig <- eigenvals(dbrda.detect, model = "constrained") # extract eigenvalues
dbrda.detect.eig <- abs(dbrda.detect.eig) # calculate eigenvalues absolute values
dbrda.detect.eig.sum <- sum(dbrda.detect.eig) # total of all eigenvalues
dbrda.detect.prop.exp <- dbrda.detect.eig/dbrda.detect.eig.sum # proportion explained by every axes
(dbrda.detect.axis1 <- round(dbrda.detect.prop.exp[1]*100, digits = 2))
(dbrda.detect.axis2 <- round(dbrda.detect.prop.exp[2]*100, digits = 2))
```

Extract dbRDA scores
```{r}
dbrda.detect.site.scores <- data.frame(scores(dbrda.detect, display = "sites", scaling = 3)) # site scores
sppscores(dbrda.detect) <- genome.detect.comm
dbrda.detect.sp.scores <- data.frame(scores(dbrda.detect, display = "species", scaling = 3)) # species scores 
```

Add host_species factor in site scores data frame for ggplot 
```{r}
host_species <- metadata.comm$host_species
dbrda.detect.site.scores <- cbind(dbrda.detect.site.scores,host_species)
rm(host_species)
```

dbRDA plot
```{r}
dbrda.detect.plot <- 
  ggplot(data = dbrda.detect.site.scores, 
         aes(x = dbRDA1, y = dbRDA2), 
         color = host_species) + 
  #geom_text_repel(data = dbrda.detect.sp.scores, 
  #                aes(x = dbRDA1, y = dbRDA2), 
  #                label=rownames(dbrda.detect.sp.scores), 
  #                color = "black", max.overlaps = 23) +
  expand_limits(x=c(-1,1), y=c(-1,1.3)) +
  geom_point(aes(shape = host_species, color = host_species), 
             size=4) +
  xlab(paste("dbRDA1 (",dbrda.detect.axis1,"%)")) +
  ylab(paste("dbRDA2 (",dbrda.detect.axis2,"%)")) +
  #geom_segment(data = vect.pcoa.env, 
  #             aes(x = 0, xend = Dim1, y = 0, yend = Dim3),
  #             arrow = arrow(length = unit(0.75, "cm")), 
  #             color = "black") +
  #geom_text_repel(data = vect.pcoa.env, 
  #                aes(x = Dim1, y = Dim3, label = variable.env, size = 3), 
  #                color = "firebrick1", 
  #                fontface = "bold", 
  #                show.legend = FALSE) +
  scale_shape_manual(values = c(17,16,16,16,17)) +
  #geom_path(data = pcoa1.ellipses, 
  #          aes(x = x, y = y, color = Group), 
  #          show.legend = FALSE) + 
  scale_color_manual(values=c("turquoise1","red1", "gold1","orchid1","limegreen")) +
  dark_theme_linedraw() +
  theme(axis.text.x = element_text(color = "grey80", size = 10, angle = 0, hjust = 0.5, vjust = 0),
        axis.text.y = element_text(color = "grey80", size = 10, angle = 0, hjust = 0, vjust = 0.5),  
        axis.title.x = element_text(color = "grey80", size = 14, angle = 0, hjust = 0.5, vjust = -1),
        axis.title.y = element_text(color = "grey80", size = 14, angle = 90, hjust = 0.5, vjust = 1),
        plot.background = element_rect(fill = "grey15"),
        panel.background = element_blank(),
        legend.background = element_blank(),
        panel.grid.major = element_line(size = 0, linetype = 'solid', colour = "grey90"), 
        panel.grid.minor = element_line(size = 0, linetype = 'solid', colour = "grey90"))
dbrda.detect.plot
ggsave("../figures/dbrda.detect.tiff", height = 5, width = 7)
```
CONCLUSION

Based on the pairwise adonis test, and as we can see on the plot above, THOC, ACSA, and COCO are mainly driving the compositional difference (based on genome detection) in Methylobacterium communities. ABBA is clearly not driving the difference.

Significative difference in composition : ACSA_vs_COCO, ACSA_vs_THOC, COCO_vs_THOC, FAGR_vs_THOC.


dbRDA
```{r}
dbrda.mean.cov <- dbrda(mean.cov.bray ~ host_species, data = metadata.comm)
anova.cca(dbrda.mean.cov, by = "terms", permutations = 999)
RsquareAdj(dbrda.mean.cov)
```
Extract proportions explained by axes 1 and 2
```{r}
dbrda.mean.cov.eig <- eigenvals(dbrda.mean.cov, model = "constrained") # extract eigenvalues
dbrda.mean.cov.eig <- abs(dbrda.mean.cov.eig) # calculate eigenvalues absolute values
dbrda.mean.cov.eig.sum <- sum(dbrda.mean.cov.eig) # total of all eigenvalues
dbrda.mean.cov.prop.exp <- dbrda.mean.cov.eig/dbrda.mean.cov.eig.sum # proportion explained by every axes
(dbrda.mean.cov.axis1 <- round(dbrda.mean.cov.prop.exp[1]*100, digits = 2))
(dbrda.mean.cov.axis2 <- round(dbrda.mean.cov.prop.exp[2]*100, digits = 2))
```

Extract dbRDA scores
```{r}
dbrda.mean.cov.site.scores <- data.frame(scores(dbrda.mean.cov, display = "sites", scaling = 3)) # site scores
sppscores(dbrda.mean.cov) <- genome.mean.cov.comm
dbrda.mean.cov.sp.scores <- data.frame(scores(dbrda.mean.cov, display = "species", scaling = 3)) # species scores 
```

Add host_species factor in site scores data frame for ggplot 
```{r}
host_species <- metadata.comm$host_species
dbrda.mean.cov.site.scores <- cbind(dbrda.mean.cov.site.scores,host_species)
rm(host_species)
```

dbRDA plot
```{r}
dbrda.mean.cov.plot <- 
  ggplot(data = dbrda.mean.cov.site.scores, 
         aes(x = dbRDA1, y = dbRDA2), 
         color = host_species) + 
  #geom_text_repel(data = dbrda.mean.cov.sp.scores, 
  #                aes(x = dbRDA1, y = dbRDA2), 
  #                label=rownames(dbrda.mean.cov.sp.scores), 
  #                color = "black", max.overlaps = 23) +
  expand_limits(x=c(-1,1), y=c(-1,1.3)) +
  geom_point(aes(shape = host_species, color = host_species), 
             size=4) +
  xlab(paste("dbRDA1 (",dbrda.mean.cov.axis1,"%)")) +
  ylab(paste("dbRDA2 (",dbrda.mean.cov.axis2,"%)")) +
  #geom_segment(data = vect.pcoa.env, 
  #             aes(x = 0, xend = Dim1, y = 0, yend = Dim3),
  #             arrow = arrow(length = unit(0.75, "cm")), 
  #             color = "black") +
  #geom_text_repel(data = vect.pcoa.env, 
  #                aes(x = Dim1, y = Dim3, label = variable.env, size = 3), 
  #                color = "firebrick1", 
  #                fontface = "bold", 
  #                show.legend = FALSE) +
  scale_shape_manual(values = c(17,16,16,16,17)) +
  #geom_path(data = pcoa1.ellipses, 
  #          aes(x = x, y = y, color = Group), 
  #          show.legend = FALSE) + 
  scale_color_manual(values=c("turquoise1","red1", "gold1","orchid1","limegreen")) +
  dark_theme_linedraw() +
  theme(axis.text.x = element_text(color = "grey80", size = 10, angle = 0, hjust = 0.5, vjust = 0),
        axis.text.y = element_text(color = "grey80", size = 10, angle = 0, hjust = 0, vjust = 0.5),  
        axis.title.x = element_text(color = "grey80", size = 14, angle = 0, hjust = 0.5, vjust = -1),
        axis.title.y = element_text(color = "grey80", size = 14, angle = 90, hjust = 0.5, vjust = 1),
        plot.background = element_rect(fill = "grey15"),
        panel.background = element_blank(),
        legend.background = element_blank(),
        panel.grid.major = element_line(size = 0, linetype = 'solid', colour = "grey90"), 
        panel.grid.minor = element_line(size = 0, linetype = 'solid', colour = "grey90"))
dbrda.mean.cov.plot
ggsave("../figures/dbrda.mean.cov.tiff", height = 5, width = 7)
```

CONCLUSION

Based on the pairwise adonis test, and as we can see on the plot above, COCO strongly drives the compositional difference (based on genome dmean coverage) in Methylobacterium communities. ACSA and THOC are also driving the difference.

Significative difference in composition : ABBA_vs_COCO, ACSA_vs_COCO (also in detection), ACSA_vs_THOC (also in detection), COCO_vs_FAGR, COCO_vs_THOC (also in detection).

I suspect that these results are partly explained by the fact that more genomes have been detected in COCO samples.


