---
title: "Methylobacterium species analyses"
output: html_document
date: "2024-04-24"
---

-################################################################################

# Libraries
```{r}
library(tidyverse)
library(vegan)
library(picante)
library(ggtree)
library(adespatial)
library(ggordiplots)
```

-################################################################################

# Import data

## Mean coverage

Import anvio genome mean coverage data
```{r}
genome.mean.cov <- read.delim("../data/anvio_outputs/ABBA/ABBA_PROFILE-SUMMARY/bins_across_samples/mean_coverage.txt", header = T, sep = "\t", dec = ".")
genome.mean.cov <- merge(genome.mean.cov, read.delim("../data/anvio_outputs/ACSA/ACSA_PROFILE-SUMMARY/bins_across_samples/mean_coverage.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.mean.cov <- merge(genome.mean.cov, read.delim("../data/anvio_outputs/COCO/COCO_PROFILE-SUMMARY/bins_across_samples/mean_coverage.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.mean.cov <- merge(genome.mean.cov, read.delim("../data/anvio_outputs/FAGR/FAGR_PROFILE-SUMMARY/bins_across_samples/mean_coverage.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.mean.cov <- merge(genome.mean.cov, read.delim("../data/anvio_outputs/THOC/THOC_PROFILE-SUMMARY/bins_across_samples/mean_coverage.txt", header = T, sep = "\t", dec = "."), by = "bins")
colnames(genome.mean.cov)[1] <- "species"
rownames(genome.mean.cov) <- genome.mean.cov$species
genome.mean.cov <- genome.mean.cov[,-1]
```

## Horizontal coverage

Import anvio genome 'detection' data
```{r}
genome.detect <- read.delim("../data/anvio_outputs/ABBA/ABBA_PROFILE-SUMMARY/bins_across_samples/detection.txt", header = T, sep = "\t", dec = ".")
genome.detect <- merge(genome.detect, read.delim("../data/anvio_outputs/ACSA/ACSA_PROFILE-SUMMARY/bins_across_samples/detection.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.detect <- merge(genome.detect, read.delim("../data/anvio_outputs/COCO/COCO_PROFILE-SUMMARY/bins_across_samples/detection.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.detect <- merge(genome.detect, read.delim("../data/anvio_outputs/FAGR/FAGR_PROFILE-SUMMARY/bins_across_samples/detection.txt", header = T, sep = "\t", dec = "."), by = "bins")
genome.detect <- merge(genome.detect, read.delim("../data/anvio_outputs/THOC/THOC_PROFILE-SUMMARY/bins_across_samples/detection.txt", header = T, sep = "\t", dec = "."), by = "bins")
colnames(genome.detect)[1] <- "species"
rownames(genome.detect) <- genome.detect$species
genome.detect <- genome.detect[,-1]
```

## Reads data
```{r}
reads.data <- read.delim("../data/nb_reads.txt", header = TRUE, sep = "\t", dec = ",", row.names = 1)
```

## Metadata
```{r}
metadata <- read.delim("../data/samples_metadata.txt", header = TRUE, sep = "\t", dec = ".", row.names = 1)
metadata <- metadata %>% arrange(row.names(.))
metadata <- metadata %>% mutate(across(c(plant_clade,plant_order,plant_family,host_species,tree), as.factor))
reads.data$host_species <- metadata$host_species
metadata$host_species <- factor(x = metadata$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
reads.data$host_species <- factor(x = reads.data$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
metadata$shrub_tree <- c("tree","tree","tree","tree","tree","tree","tree","tree","tree","tree","shrub","shrub","shrub","shrub","shrub","tree","tree","tree","tree","tree","tree","tree","tree","tree","tree")
metadata$shrub_tree <- factor(x = metadata$shrub_tree, levels = c("shrub","tree"))
metadata$gymno_angio_tree <- c("gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_shrub","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","angiosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree","gymnosperm_tree")
metadata$gymno_angio_tree <- factor(x = metadata$gymno_angio_tree, levels = c("gymnosperm_tree","angiosperm_tree","angiosperm_shrub"))
metadata.bac.comm <- metadata
```

## Reference genomes table
```{r}
ref.genomes <- read.delim("../data/meth_all_genomes.txt", header = TRUE, sep = "\t", dec = ",")
ref.genomes <- ref.genomes[,c(2:4)]
ref.genomes <- ref.genomes %>% mutate(across(c(species,strain_original_name,group), as.factor))
colnames(ref.genomes) <- c("species","strain","clade")
```

## Phylogenetic tree
```{r}
meth.tree <- read.tree("../data/RAxML_bipartitionsBranchLabels.Partition-Concatenated-genes-nov21-b=500-rooted")
strains.list <- read.csv2("../data/List-genomes.csv", sep = ";", dec = ".")
# rename tips of the tree - prepare data
st.sp <- strains.list[,c(1,2,4)]
colnames(st.sp) <- c("st","strain","Meth_species")
tip.labels <- data.frame(meth.tree[["tip.label"]])
colnames(tip.labels) <- "st"
st.sp <- st.sp[st.sp$st %in% tip.labels$st,]
tip.labels$st <- factor(tip.labels$st, levels = unique(tip.labels$st))
st.sp$st <- factor(st.sp$st, levels = unique(tip.labels$st))
st.sp <- st.sp[order(st.sp$st), ]
rownames(st.sp) <- seq_len(nrow(st.sp))
tip.labels <- cbind(tip.labels, st.sp$strain, st.sp$Meth_species)
colnames(tip.labels) <- c("st","strain","Meth_species")

# rename with strain's names
meth.tree[["tip.label"]] <- tip.labels$strain
meth.tree[["tip.label"]]

# rename with Methylo species
meth.tree.sp <- meth.tree
meth.tree.sp[["tip.label"]] <- tip.labels$Meth_species
meth.tree.sp[["tip.label"]] <- gsub("Meth", "M", meth.tree.sp[["tip.label"]])
meth.tree.sp[["tip.label"]]
```

Plot tree
```{r}
plot(meth.tree, show.tip.label = T)
ggtree(meth.tree) + 
  theme_tree() +
  geom_tiplab(size=2)
```


-################################################################################

# Horizontal coverage threshold

Keep genomes with detection threshold >= 0.34 (outliers based on z values, see script 'genomes_horizontal_cov_hist.Rmd')
```{r}
genome.detect_0.34 <- data.frame(ifelse(genome.detect < 0.34, 0, 1))
genome.detect_0.34 <- genome.detect_0.34[rowSums(genome.detect_0.34)>0,]
genome.detect_0.34 <- genome.detect_0.34[,colSums(genome.detect_0.34)>0]
```

Statistics
```{r}
length(rownames(genome.detect_0.34))
rowSums(genome.detect_0.34)
length(colnames(genome.detect_0.34))
colSums(genome.detect_0.34)
sum(genome.detect_0.34)
```

## Presence-absence community matrix
```{r}
genomes.PA.comm <- data.frame(t(genome.detect_0.34))
```

-################################################################################

# Coverage, metadata and reads data associated with 'detected' genomes

## Mean coverage

```{r}
genomes.mean.cov_0.34 <- genome.mean.cov[rownames(genome.detect_0.34), colnames(genome.detect_0.34)]
genomes.mean.cov_0.34 <- as.data.frame(mapply(function(x, y) ifelse(x == 1, y, 0), genome.detect_0.34, genomes.mean.cov_0.34))
rownames(genomes.mean.cov_0.34) <- row.names(genome.detect_0.34)
genomes.cov.comm <- data.frame(t(genomes.mean.cov_0.34))
```

## Count data : transform mean coverage to count data
```{r}
genomes.cov.comm.c <- round(genomes.cov.comm, digits = 0)
```

## Metadata & add reads data

```{r}
metadata.comm <- metadata[rownames(genomes.cov.comm), ]
reads.data.comm <- reads.data[rownames(genomes.cov.comm), ]
metadata.comm$filtered_pairs <- reads.data.comm$filtered_pairs
```

## Richness and Shannon Index
```{r}
metadata.comm$richness <- rowSums(genomes.PA.comm)
metadata.comm$shannon <- vegan::diversity(genomes.cov.comm, index = "shannon")
```

-################################################################################

# Correlation between richness and number of filtered pairs.

```{r}
lm <- lm(richness ~ filtered_pairs, data = metadata.comm)
plot(richness ~ filtered_pairs, data = metadata.comm)
summary(lm)
plot(lm)
hist(residuals(lm), breaks = 25)
shapiro.test(residuals(lm))
```
-################################################################################

# Prevalence of species across samples

Which species is present in the most samples (most prevalent species in the forest)?

## Coverage
```{r}
species.cov.across <- as.data.frame(rowSums(t(genomes.cov.comm)))
species.cov.across$species <- as.factor(rownames(species.cov.across))
colnames(species.cov.across) <- c("cov.across", "species")
species.cov.across <- merge(species.cov.across, ref.genomes, by = "species")
species.cov.across$isolated_from <- c("Cerastium holosteoides","Haplocladium microphyllum","Arabidopsis
thaliana","Lolium prenne","Pseudosasa japonica","Fagus grandifolia","Fagus grandifolia","Arabidopsis
thaliana","Arabidopsis thaliana","Arabidopsis thaliana","Acer saccharum","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Acer saccharum","Fagus grandifolia","Acer saccharum","Fagus grandifolia","Cerastium holosteoides","Trifolium repens")
```

Plot
```{r}
ggplot(species.cov.across, aes(x = species, y = cov.across))+
  geom_bar(stat = "identity", fill = "black") +
  labs(x = "Methylobacterium species", y = "Coverage across samples") +
  ylim(0,400) +
  theme_bw() +
  theme(text = element_text(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 11),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
```

## Presence-absence
```{r}
species.PA.across <- as.data.frame(rowSums(t(genomes.PA.comm)))
species.PA.across$species <- as.factor(rownames(species.PA.across))
colnames(species.PA.across) <- c("presence.across", "species")
species.PA.across <- merge(species.PA.across, ref.genomes, by = "species")
species.PA.across$isolated_from <- c("Cerastium holosteoides","Haplocladium microphyllum","Arabidopsis
thaliana","Lolium prenne","Pseudosasa japonica","Fagus grandifolia","Fagus grandifolia","Arabidopsis
thaliana","Arabidopsis thaliana","Arabidopsis thaliana","Acer saccharum","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Fagus grandifolia","Acer saccharum","Fagus grandifolia","Acer saccharum","Fagus grandifolia","Cerastium holosteoides","Trifolium repens")
```

Plot
```{r}
ggplot(species.PA.across, aes(x = species, y = presence.across))+
  geom_bar(stat = "identity", fill = "black") +
  labs(x = "Methylobacterium species", y = "Number of samples") +
  ylim(0,25) +
  theme_bw() +
  theme(text = element_text(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 11),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
```

```{r}
hist(species.PA.across$presence.across, breaks = 50)
```


-################################################################################

# Alpha diversity

## Richness

```{r}
lm <- lm(richness ~ host_species, data = metadata.comm)
plot(richness ~ host_species, data = metadata.comm)
summary(lm)
plot(lm)
hist(residuals(lm))
shapiro.test(residuals(lm))
```

Post-hoc
```{r}
# Shrub vs. trees

## Student t-test assumptions : normal distribution of data and equal variance
hist(metadata.comm[metadata.comm$shrub_tree == "tree",]$richness, breaks = 25)
shapiro.test(metadata.comm[metadata.comm$shrub_tree == "tree",]$richness)
hist(metadata.comm[metadata.comm$shrub_tree == "shrub",]$richness, breaks = 25)
shapiro.test(metadata.comm[metadata.comm$shrub_tree == "shrub",]$richness)
var(metadata.comm[metadata.comm$shrub_tree == "tree",]$richness)
var(metadata.comm[metadata.comm$shrub_tree == "shrub",]$richness)
## unequal variance so welch t-test
t1 <- t.test(richness ~ shrub_tree, data = metadata.comm, var.equal = F)
t1

# Gymno vs. angio trees

## Student t-test assumptions : normal distribution of data and equal variance
hist(metadata.comm[metadata.comm$gymno_angio_tree == "gymnosperm_tree",]$richness, breaks = 25)
shapiro.test(metadata.comm[metadata.comm$gymno_angio_tree == "gymnosperm_tree",]$richness)
hist(metadata.comm[metadata.comm$gymno_angio_tree == "angiosperm_tree",]$richness, breaks = 25)
shapiro.test(metadata.comm[metadata.comm$gymno_angio_tree == "angiosperm_tree",]$richness)
var(metadata.comm[metadata.comm$gymno_angio_tree == "gymnosperm_tree",]$richness)
var(metadata.comm[metadata.comm$gymno_angio_tree == "angiosperm_tree",]$richness)
## unequal variance so welch t-test
t2 <- t.test(richness ~ gymno_angio_tree, data = metadata.comm[c(1:9,15:24),], var.equal = F)
t2

# p-values adjustement
p.adjust(c(t1[["p.value"]],t2[["p.value"]]), method = "bonferroni")
```

Plot
```{r}
ggplot(metadata.comm, aes(x = host_species, y = richness)) +
geom_boxplot(alpha = 1, aes(fill = host_species), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = host_species), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
scale_fill_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"))+
scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Host Species", 
     y = expression(bolditalic("Methylobacterium") ~ bold("species richness")))+
scale_x_discrete(labels = c("ABBA" = "A. balsamea", "THOC" = "T. occidentalis","ACSA" = "A. saccharum", "FAGR" = "F. grandifolia", "COCO" = "C. cornuta"))+
ylim(0,25) +
theme_bw()+
theme(text = element_text(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0, face = "italic"),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/figure_S8_Methylo_species_richness.pdf", width = 6, height = 4)
```

## Shannon index

```{r}
hist(metadata.comm$shannon, breaks = 25)
shapiro.test(metadata.comm$shannon)
kruskal.test(shannon ~ host_species, data = metadata.comm)
rstatix::kruskal_effsize(shannon ~ host_species, data = metadata.comm)
```

## Pielou Index

Calculation
```{r}
metadata.comm$pielou <- metadata.comm$shannon/log(metadata.comm$richness)
hist(metadata.comm$pielou, breaks = 25)
shapiro.test(metadata.comm$pielou)
```

Model
```{r}
lm <- lm(pielou ~ host_species, metadata.comm)
plot(pielou ~ host_species, metadata.comm)
summary(lm)
hist(residuals(lm))
shapiro.test(residuals(lm))
```

Plot
```{r}
ggplot(metadata.comm, aes(x = host_species, y = pielou)) +
geom_boxplot(alpha = 1, aes(fill = host_species), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = host_species), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
scale_fill_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"))+
scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Host Species", 
     y = "Pielou Index")+
scale_x_discrete(labels = c("ABBA" = "A. balsamea", "THOC" = "T. occidentalis","ACSA" = "A. saccharum", "FAGR" = "F. grandifolia", "COCO" = "C. cornuta"))+
ylim(0.5,1) +
theme_bw()+
theme(text = element_text(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0, face = "italic"),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/figure_S9_Methylo_species_pielou.pdf", width = 6, height = 4)
```


Correlation between richness and evenness
```{r}
lm <- lm(pielou ~ richness, metadata.comm)
plot(pielou ~ richness, metadata.comm)
summary(lm)
hist(residuals(lm))
```


-################################################################################

# Beta diversity

## Composition

### Barplot
```{r}
abundances <- picante::matrix2sample(genomes.cov.comm)
colnames(abundances) <- c("sample","abund","Meth_species")
abundances$sample <- factor(x = abundances$sample, levels =c ("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M18","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
abundances$Meth_species <- factor(x = abundances$Meth_species, 
levels = c(
"M_sp_026",
"M_sp_024",
"M_sp_029",
"M_pseudosasicola",
"M_sp_030",
"M_mesophilicum",
"M_sp_021",
"M_sp_027",
"M_sp_023",
"M_sp_022",
"M_longum",
"M_sp_025",
"M_sp_034",
"M_sp_018",
"M_cerastii",
"M_trifolii",
"M_sp_010",
"M_sp_008",
"M_sp_003",
"M_sp_011",
"M_sp_004",
"M_thuringiense",
"M_haplocladii"
))

ggplot(abundances, aes(x = sample, y = abund, fill = Meth_species)) +
  geom_bar(stat = "identity") +
  labs(x = "Sample", y = "Genome coverage") +
  scale_fill_manual(values=c(
  "#C2B280",
  "#1A8A00",
  "#8E6A48",
  "#B5D192",  
  "#3B5B29",
  "#E1B05D",  
  "#0095C8",
  "#B66D3B",  
  "#4E8",
  "#A77B3D",
  "#1F8C8B",
  "#E5B5B2", 
  "#B7C9D9", 
  "#E15F4C", 
  "#FFCC00",
  "#008B5B",
  "#7C9DB8", 
  "#D9C3B2",
  "#5A9C6A", 
  "#F4A300",
  "#C4A65D",
  "#A6C8C4", 
  "#4B8C2A" 
), name = expression(italic("Methylobacterium") ~ "species")) +
  guides(fill = guide_legend(ncol = 1)) +
  theme_bw() +
  theme(text = element_text(),
      legend.justification = c(0.5, 1),
      legend.text = element_text(size = 10.5),
      legend.key.size = unit(0.43, "cm"),
      legend.position = "right", 
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 11),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/figure_2_Methylo_species_barplot.pdf", width = 8, height = 5.5)
```

## Bray-Curtis distance
```{r}
genomes.cov.beta <- adespatial::beta.div.comp(genomes.cov.comm, coef = "BS", quant = TRUE)
genomes.bray.dist <- genomes.cov.beta[["D"]]
```

## PERMANOVA
```{r}
genomes.bray.perm <- adonis2(genomes.bray.dist ~ host_species, data = metadata.comm)
genomes.bray.perm
```

## BETADISPER
```{r}
genomes.bray.disper <- betadisper(genomes.bray.dist, metadata.comm$host_species)
anova(genomes.bray.disper)
plot(genomes.bray.disper)
```

## Pairwise PERMANOVAs
```{r}
shrub_tree_PERM <- adonis2(genomes.bray.dist ~ shrub_tree, data = metadata.comm)
shrub_tree_PERM
shrub_tree_DISP <- betadisper(genomes.bray.dist, metadata.comm$shrub_tree)
anova(shrub_tree_DISP)
plot(shrub_tree_DISP)

genomes.bray.dist.noCOCO <- data.frame(as.matrix(genomes.bray.dist))
genomes.bray.dist.noCOCO <- genomes.bray.dist.noCOCO[c(1:9,15:24),c(1:9,15:24)]
genomes.bray.dist.noCOCO <- as.dist(as.matrix(genomes.bray.dist.noCOCO), diag = FALSE, upper = FALSE)
gymno_angio_tree_PERM <- adonis2(genomes.bray.dist.noCOCO ~ gymno_angio_tree, data = metadata.comm[c(1:9,15:24),], perm =999)
gymno_angio_tree_PERM
gymno_angio_tree_DISP <- betadisper(genomes.bray.dist.noCOCO, metadata.comm[c(1:9,15:24),]$gymno_angio_tree)
anova(gymno_angio_tree_DISP)
plot(gymno_angio_tree_DISP, label = F)
```

## PCoA
```{r}
genomes.pcoa <- cmdscale(genomes.bray.dist, k =(nrow(genomes.cov.comm) - 1), eig = TRUE)
pcoa.eig <- eigenvals(genomes.pcoa)
pcoa.eig <- abs(pcoa.eig) 
sum.pcoa.eig <- sum(pcoa.eig) 
pcoa.prop.exp <- pcoa.eig/sum.pcoa.eig 
(pcoa.axis1 <- pcoa.prop.exp[1])
(pcoa.axis2 <- pcoa.prop.exp[2])
site.scores.pcoa <- scores(genomes.pcoa) 
sp.scores.pcoa <- wascores(genomes.pcoa$points, genomes.cov.comm) 
site.scores.pcoa <- data.frame(site.scores.pcoa)
sp.scores.pcoa <- data.frame(sp.scores.pcoa)
host_species <- metadata.comm$host_species
site.scores.pcoa <- cbind(site.scores.pcoa, host_species)
pcoa.ggordiplot <- gg_ordiplot(genomes.pcoa, choices = c(1,2), groups = metadata.comm$host_species, kind = "sd", conf = 0.95, pt.size = 3)
pcoa.ellipses <- pcoa.ggordiplot$df_ellipse # extract confidence ellipse
```

Plot
```{r}
ggplot(data = site.scores.pcoa, 
         aes(x = Dim1, y = Dim2), 
         color = host_species) + 
  expand_limits(x=c(-0.8,0.8), y=c(-0.8,0.8)) +
  geom_point(aes(shape = host_species, color = host_species), show.legend = T, 
             size=3) +
  xlab(paste("PCoA1 (",round(pcoa.axis1*100, digits = 2),"%)", sep = "")) +
  ylab(paste("PCoA2 (",round(pcoa.axis2*100, digits = 2),"%)", sep = "")) +
  scale_shape_manual(values = c(16,16,16,16,16)) +
  guides(shape = "none") +
  geom_path(data = pcoa.ellipses, 
            aes(x = x, y = y, color = Group), 
            show.legend = FALSE) + 
  scale_color_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"), name = "Host species", labels = c("ABBA" = expression(italic("A. balsamea")), "THOC" = expression(italic("T. occidentalis")),"ACSA" = expression(italic("A. saccharum")), "FAGR" = expression(italic("F. grandifolia")), "COCO" = expression(italic("C. cornuta")))) +
  theme_bw() +
  theme(text = element_text(),
      legend.justification = c(0.5, 0),
      legend.text = element_text(size = 11),
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/figure_3A_genomes_pcoa.pdf", width = 6, height = 4)
```

-################################################################################

# Phylogeny

## Replace 'genomes.cov.comm' columns names with strain instead of species.
```{r}
genomes.cov.comm.st <- genomes.cov.comm
ref.genomes.sub <- ref.genomes[ref.genomes$species %in% colnames(genomes.cov.comm.st),]
colnames(genomes.cov.comm.st) <- ref.genomes.sub$strain
```

## Prune tree
```{r}
pruned.tree <- prune.sample(genomes.cov.comm.st, meth.tree)
plot(pruned.tree, show.tip.label = TRUE)
nodelabels()
```

## Test if Methylobacterium communities were composed of phylogenetically related species

### Cophenetic distance
```{r}
phydist <- cophenetic(pruned.tree)
```

### SES.MPD : Standardized Effect Size MPD
```{r}
# PRESENCE-ABSENCE
ses.mpd.results <- ses.mpd(genomes.cov.comm.st, phydist, null.model = "taxa.labels", abundance.weighted = FALSE, runs = 999, iterations = 1000)
ses.mpd.results$host_species <- metadata.comm$host_species
ses.mpd.results$plant_clade <- metadata.comm$plant_clade
ses.mpd.results$mpd.obs <- round(ses.mpd.results$mpd.obs, digits = 4)
ses.mpd.results$mpd.rand.mean <- round(ses.mpd.results$mpd.rand.mean, digits = 4)
ses.mpd.results$mpd.rand.sd <- round(ses.mpd.results$mpd.rand.sd, digits = 4)
ses.mpd.results$mpd.obs.z <- round(ses.mpd.results$mpd.obs.z, digits = 4)
#saveRDS(ses.mpd.results, "../rds/ses.mpd.results.rds")
ses.mpd.results <- readRDS("../rds/ses.mpd.results.rds")

plot(mpd.obs.z ~ host_species, ses.mpd.results)
plot(mpd.obs.z ~ plant_clade, ses.mpd.results)
mean(na.omit(ses.mpd.results$mpd.obs.z))
sd(na.omit(ses.mpd.results$mpd.obs.z))
```

## Do some host species harbors communities composed of more closely related species than other hosts?

### ANOVA
```{r}
lm <- lm(mpd.obs.z ~ host_species, ses.mpd.results)
summary(lm)
plot(lm)
hist(residuals(lm))
shapiro.test(residuals(lm))
```

### Plot
```{r}
ggplot(ses.mpd.results, aes(x = host_species, y = mpd.obs.z)) +
geom_boxplot(alpha = 1, aes(fill = host_species), show.legend = FALSE, outlier.shape = NA)+
geom_point(aes(color = host_species), 
           position = position_jitter(0.075), 
           shape = 16, 
           size = 3, show.legend = FALSE)+
scale_fill_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"))+
scale_color_manual(values=c("black","black","black","black","black"))+
labs(x = "Host Species", 
     y = "Standardized effect size of MPD")+
scale_x_discrete(labels = c("ABBA" = "A. balsamea", "THOC" = "T. occidentalis","ACSA" = "A. saccharum", "FAGR" = "F. grandifolia", "COCO" = "C. cornuta"))+
ylim(-6,1) +
theme_bw()+
theme(text = element_text(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0, face = "italic"),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )
ggsave("../figures/figure_S10_Methylo_SESMPD.pdf", width = 6, height = 4)
```


## Do host species harbor distinct phylogenetic groups of Methylobacterium species?

### Phylogenetic beta diversity
```{r}
# calculate inter-community mean pairwise distance (MPD)
comm.dist <- comdist(genomes.cov.comm.st, phydist)
```

### Cluster
```{r}
comm.dist.clusters <- hclust(comm.dist)
plot(comm.dist.clusters)
```

### PERMANOVA
```{r}
phy.perm <- adonis2(comm.dist ~ host_species, data = metadata.comm)
phy.perm
```

### BETADISPER
```{r}
phy.disper <- betadisper(comm.dist, metadata.comm$host_species)
anova(phy.disper)
plot(phy.disper)
```

## Tree with heatmap of abundance

Prepare data
```{r}
abundances <- matrix2sample(genomes.cov.comm.st)
colnames(abundances) <- c("sample","abund","Meth_species")
abundances$sample <- factor(x = abundances$sample, levels =c ("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M18","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
```



```{r}
# Create dataframe with clades and associated colors 
clades <- data.frame(
  node = c(1,2,25,19,20,21,22,23,42,43,44,45,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,24,26),  # Nodes for different clades
  color = c("darkorange","darkorange","black", "red","red","red","red","red","black","red","red","red", "blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","black","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","blue","black","black"),
  clade = c("Clade B","Clade B","none", "Clade D","Clade D","Clade D","Clade D","Clade D","none","Clade D","Clade D","Clade D", "Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","none","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A","Clade A", "none","none")
)

# Plot clade-colored tree with strains as tip labels
gg_tr <- ggtree(pruned.tree, aes(color = as.factor(node))) + geom_tiplab(align=TRUE, color="black", size = 4.25) +
  scale_x_continuous(expand=expansion(0.5)) +
  scale_color_manual(values = clades$color, 
                     breaks = clades$node,
                     labels = clades$clade)+
  theme(legend.position="none",
        text = element_text())
gg_tr

# Get tip order
tip_order <- paste("", 
        paste0(get_taxa_name(gg_tr), collapse=","))
tip_order

# Rearrange factor levels
abundances$Meth_species <- factor(x = abundances$Meth_species, levels = c("J-077","J-048","E-066","BL36","J-001","NBRC15688T","E-065","J-088","J-030","J-026","DSM23933T","J-070","J-076","E-025","DSM23679T","DSM23632","Leaf125","Leaf89","J-078","Leaf99","J-090","DSM23674","NBRC107714"))

# Plot the heatmap with strains in y axis
gg_heat <- ggplot(abundances, aes(x = sample, y = Meth_species)) + 
  geom_tile(aes(fill=abund)) + 
  scale_fill_gradient2(high= "darkgreen", name = "Coverage", transform = "sqrt")+ 
  theme(legend.position="none") +
  scale_y_discrete(limits=rev) +
  labs(x = "Sample") +
    theme_bw() +
      theme(
      text = element_text(),
      legend.justification = c(0.5, 0),
      legend.title = element_text(size = 12, margin = margin(b = 15)),
      legend.text = element_text(size = 12),
      legend.key.size = unit(0.9, "cm"),
      legend.spacing.y = unit(0.5, "cm"),
      legend.position = "right", 
      panel.grid.major = element_line(color = "gray90", size = 0.25), 
      panel.background = element_rect(fill = "black"),
      #panel.grid.minor = element_line(color = "lightgray", size = 0.25), 
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, vjust = 0.35, hjust = 1),
      axis.text.y = element_text(color = "black", size = 12),
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2, face = "bold")
      )

(gg_tr + gg_heat) # Colored tree with strains name on both figures

# Order of strains is ok, let's change names of strain to species on heatmap.

gg_heat.sp <- gg_heat +
              theme(axis.title.y = element_blank()) +
              scale_y_discrete(labels = c(
              "J-077" = "M_sp_026",
              "J-048" = "M_sp_024",
              "E-066" = "M_sp_029",
              "BL36" = "M_pseudosasicola",
              "J-001" = "M_sp_030",
              "NBRC15688T" = "M_mesophilicum",
              "E-065" = "M_sp_021",
              "J-088" = "M_sp_027",
              "J-030" = "M_sp_023",
              "J-026" = "M_sp_022",
              "DSM23933T" = "M_longum",
              "J-070" = "M_sp_025",
              "J-076" = "M_sp_034",
              "E-025" = "M_sp_018",
              "DSM23679T" = "M_cerastii",
              "DSM23632" = "M_trifolii",
              "Leaf125" = "M_sp_010",
              "Leaf89" = "M_sp_008",
              "J-078" = "M_sp_003",
              "Leaf99" = "M_sp_011",
              "J-090" = "M_sp_004",
              "DSM23674" = "M_thuringiense",
              "NBRC107714" = "M_haplocladii"
              ),
              limits = rev(levels(abundances$Meth_species)))
gg_heat.sp

plot.heat.tree <- gg_tr + gg_heat.sp
plot.heat.tree

ggsave("../figures/figure_4_Methylobacterium_tree_heatmap.pdf", width = 12, height = 6)
```

-################################################################################