---
title: "pN/pS analyses"
output: html_document
date: "2024-11-14"
---

-################################################################################

# Libraries
```{r}
library(tidyverse)
library(picante)
library(car)
#library(lmerTest)
library(emmeans)
library(ggeffects)
library(MuMIn)
#library(tls)
```

-################################################################################

# Import function table
```{r}
data.fct.cov <- readRDS("../rds/data.fct.cov.rds")
```

-################################################################################

# M_sp_018

## Import data
```{r}
pNpS_sp_018 <- read.table("../data/anvio_outputs/pNpS_sp_018-03/pNpS.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pN_sp_018 <- read.table("../data/anvio_outputs/pNpS_sp_018-03/pN.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pS_sp_018 <- read.table("../data/anvio_outputs/pNpS_sp_018-03/pS.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pNpS_sp_018$pN <- pN_sp_018$pN_gene_consensus
pNpS_sp_018$pS <- pS_sp_018$pS_gene_consensus
length(which(pNpS_sp_018$pNpS_gene_consensus == 0)) #180
str(pNpS_sp_018)
```

## Edit dataframe
```{r}
# Remove rows where there was missing values to calculate pN/pS ratio (NA, Inf)
pNpS_sp_018 <- pNpS_sp_018[!is.na(pNpS_sp_018$pNpS_gene_consensus) & pNpS_sp_018$pNpS_gene_consensus != "Inf" & pNpS_sp_018$pNpS_gene_consensus != 0,]
# We must remove NAs (not enough data to compute pS/pN) as well as extreme values (0 and Inf), because those values might be due to higher stochasticity when number of variants is low.

length(unique(pNpS_sp_018$corresponding_gene_call)) # 4802 genes left after removal
str(pNpS_sp_018)

# Remove samples in which the species was not present
samples_presence <- c("ABBA_M12","ABBA_M13","ABBA_M14","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M18","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05")
pNpS_sp_018 <- pNpS_sp_018[pNpS_sp_018$sample_id %in% samples_presence,]
pNpS_sp_018 <- droplevels(pNpS_sp_018)

# Add host species factor
pNpS_sp_018$host_species <- as.factor(substr(pNpS_sp_018$sample_id, 1, 4))
pNpS_sp_018$host_species <- factor(pNpS_sp_018$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
str(pNpS_sp_018)

# Add clade factor
pNpS_sp_018$plant_clade <- factor(
  ifelse(grepl("^(ACSA|FAGR|COCO)", pNpS_sp_018$sample_id),
         "angiosperm",
         ifelse(grepl("^(ABBA|THOC)", pNpS_sp_018$sample_id),
                "gymnosperm",
                NA))
)
str(pNpS_sp_018)

# Add plant type factor
pNpS_sp_018$plant_type <- factor(
  ifelse(grepl("^(ACSA|FAGR|ABBA|THOC)", pNpS_sp_018$sample_id),
         "tree",
         ifelse(grepl("^(COCO)", pNpS_sp_018$sample_id),
                "shrub",
                NA))
)
str(pNpS_sp_018)
```

## Distribution of pN/pS measurements across all genes and all samples
```{r}
hist(pNpS_sp_018$pNpS_gene_consensus, breaks = 1000000)
hist(log(pNpS_sp_018$pNpS_gene_consensus), breaks = 10000)
```

## Add core/accessory and order table
```{r}
colnames(pNpS_sp_018)[1] <- "gene_callers_id"
data.pangenome <- readRDS("../rds/data.pangenome.rds")
pNpS_sp_018 <- merge(pNpS_sp_018, data.pangenome[,c(3:4)], by = "gene_callers_id")
pNpS_sp_018$type <- factor(x = pNpS_sp_018$type, levels = c("core","accessory"))
pNpS_sp_018 <- pNpS_sp_018[order(pNpS_sp_018$type, -pNpS_sp_018$pNpS_gene_consensus), ]
pNpS_sp_018$gene_callers_id <- factor(x = pNpS_sp_018$gene_callers_id, levels = unique(pNpS_sp_018$gene_callers_id))
pNpS_sp_018$sample_id <- factor(x = pNpS_sp_018$sample, levels =c ("ABBA_M12","ABBA_M13","ABBA_M14","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M18","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
str(pNpS_sp_018)
```

## Genes under positive selection

### Data prep
```{r}
# Number of unique gene calls for which at least in one sample the pN/pS ratio > 1 (across all samples and host species)
length(unique(pNpS_sp_018[pNpS_sp_018$pNpS_gene_consensus > 1,]$gene_callers_id))

# Make a vector of all genes that have at least one positive pN/pS value in a sample (any host)
pos.sel.genes <- unique(pNpS_sp_018[pNpS_sp_018$pNpS_gene_consensus > 1,]$gene_callers_id)
pos.sel.genes <- droplevels(pos.sel.genes)
```

```{r}
# Create pNpS dataframe with only genes that have at least one positive value
pNpS_sp_018.pos <- pNpS_sp_018[pNpS_sp_018$gene_callers_id %in% pos.sel.genes,]
pNpS_sp_018.pos <- droplevels(pNpS_sp_018.pos)

# Create a table to know in how many samples of each host_species there is variability info on a given gene
freq.table <- as.data.frame(table(pNpS_sp_018.pos$gene_callers_id, pNpS_sp_018.pos$host_species))
colnames(freq.table) <- c("gene_callers_id","host_species","num_samples")

# Keep only gene-host info for hosts that have at least 3 samples for the given gene
freq.table <- freq.table[freq.table$num_samples >= 3,]
freq.table <- droplevels(freq.table)

# Separate freq.table per host species
freq.table.ABBA <- freq.table[freq.table$host_species == "ABBA",]
freq.table.ABBA <- droplevels(freq.table.ABBA)
freq.table.ACSA <- freq.table[freq.table$host_species == "ACSA",]
freq.table.ACSA <- droplevels(freq.table.ACSA)
freq.table.COCO <- freq.table[freq.table$host_species == "COCO",]
freq.table.COCO <- droplevels(freq.table.COCO)
freq.table.FAGR <- freq.table[freq.table$host_species == "FAGR",]
freq.table.FAGR <- droplevels(freq.table.FAGR)
freq.table.THOC <- freq.table[freq.table$host_species == "THOC",]
freq.table.THOC <- droplevels(freq.table.THOC)
```

### Confidence interval tests 

We will keep only genes with a mean log(pN/pS(gene)) > 0 and a lower confidence interval > 0.
```{r}
sig.pos.sel.genes_sp_018 <- data.frame(
  gene_callers_id = character(0),
  mean = numeric(0),
  conf_int_lower = numeric(0),
  conf_int_upper = numeric(0),
  t_statistic = numeric(0),
  p_value = numeric(0),
  adj.p = numeric(0),
  host_species = character(0),
  stringsAsFactors = FALSE
)

host_species <- c("ABBA","THOC","ACSA","FAGR","COCO")

for (host in host_species) {

genes.to.t.test <- unique(freq.table[freq.table$host_species == host,]$gene_callers_id)
genes.to.t.test <- droplevels(genes.to.t.test)  

means.of.genes <- data.frame(
  gene_callers_id = character(0),
  mean_pNpS = numeric(0),
  sd_pNpS = numeric(0)
)

for (gene in genes.to.t.test) {
  mean <- mean(log(pNpS_sp_018[pNpS_sp_018$gene_callers_id == gene & pNpS_sp_018$host_species == host,]$pNpS_gene_consensus))
  sd <- sd(pNpS_sp_018[pNpS_sp_018$gene_callers_id == gene & pNpS_sp_018$host_species == host,]$pNpS_gene_consensus)
  means.of.genes <- rbind(means.of.genes, data.frame(
    gene_callers_id = gene,
    mean_pNpS = mean,
    sd_pNpS = sd))
}

means.of.genes$mean_minus_sd <- means.of.genes$mean - means.of.genes$sd

genes.to.t.test <- as.factor(unique(means.of.genes[means.of.genes$mean_pNpS > 0,]$gene_callers_id))
genes.to.t.test <- droplevels(genes.to.t.test)

t_test_results <- data.frame(
  gene_callers_id = character(0),
  mean_pNpS = numeric(0),
  log_mean_pNpS = numeric(0),
  conf_int_lower = numeric(0),
  conf_int_upper = numeric(0),
  t_statistic = numeric(0),
  p_value = numeric(0),
  stringsAsFactors = FALSE
)

for (gene in genes.to.t.test) {
  # Perform the t-test for the current level
  t_test_result <- t.test(log(pNpS_sp_018[pNpS_sp_018$gene_callers_id == gene 
                                      & pNpS_sp_018$host_species == host,]$pNpS_gene_consensus), mu = 0, alternative = "greater")
  # Store the result in the data frame
  t_test_results <- rbind(t_test_results, data.frame(
    gene_callers_id = gene,
    mean_pNpS = exp(t_test_result[["estimate"]][["mean of x"]]),
    log_mean_pNpS = t_test_result[["estimate"]][["mean of x"]],
    conf_int_lower = t_test_result$conf.int[1],
    conf_int_upper = t_test_result$conf.int[2],
    t_statistic = t_test_result$statistic,
    p_value = t_test_result$p.value
  ))
}

t_test_results$adj.p <- p.adjust(t_test_results$p_value, method = "BH") # Benjamini-Hochberg adjustment
t_test_results.sig <- t_test_results[t_test_results$conf_int_lower > 0,]

# Add host species column
t_test_results.sig$host_species <- rep(host, nrow(t_test_results.sig))

# Add to final dataframe
sig.pos.sel.genes_sp_018 <- rbind(sig.pos.sel.genes_sp_018,t_test_results.sig)

}

rownames(sig.pos.sel.genes_sp_018) <- NULL
```

### Import functions
```{r}
data.fct.cov <- readRDS("../rds/data.fct.cov.rds")
sig.pos.sel.genes_sp_018.fct <- merge(sig.pos.sel.genes_sp_018, data.fct.cov, by = "gene_callers_id")
length(unique(sig.pos.sel.genes_sp_018.fct$gene_callers_id))
# 41 unique positively selected genes. 42 "gene-host" combinations, including 2 FAGR, and 40 COCO
# All accessory genes
# 13 unique ANNOTATED genes
```

### DAGCs?
```{r}
DAGs.functions.all.info <- read_delim("../data/DAGCs_fully_annotated.csv", delim = ",")
pos.sel.DAGCs_sp_018 <- DAGs.functions.all.info[DAGs.functions.all.info$gene_cluster_id %in% sig.pos.sel.genes_sp_018.fct$gene_cluster_id,]
# includes 5 DAGCs

sig.pos.sel.genes_sp_018.fct$DAGCs <- ifelse(
  sig.pos.sel.genes_sp_018.fct$gene_cluster_id %in% pos.sel.DAGCs_sp_018$gene_cluster_id,
  paste(pos.sel.DAGCs_sp_018$focal, pos.sel.DAGCs_sp_018$reference, sep = "_vs_"),
  "non-DAGC"
)

sig.pos.sel.genes_sp_018.fct <- sig.pos.sel.genes_sp_018.fct[,c(1:13,58,14:57)]
```

### Save data
```{r}
#saveRDS(sig.pos.sel.genes_sp_018.fct, "../rds/sig.pos.sel.genes_sp_018.fct.rds")
sig.pos.sel.genes_sp_018.fct <- readRDS("../rds/sig.pos.sel.genes_sp_018.fct.rds")
```

-################################################################################

# M_sp_021

## Import data

```{r}
pNpS_sp_021 <- read.table("../data/anvio_outputs/pNpS_sp_021-03/pNpS.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pN_sp_021 <- read.table("../data/anvio_outputs/pNpS_sp_021-03/pN.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pS_sp_021 <- read.table("../data/anvio_outputs/pNpS_sp_021-03/pS.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pNpS_sp_021$pN <- pN_sp_021$pN_gene_consensus
pNpS_sp_021$pS <- pS_sp_021$pS_gene_consensus
length(which(pNpS_sp_021$pNpS_gene_consensus == 0)) #139
str(pNpS_sp_021)
```

## Edit dataframe
```{r}
# Remove rows where there was missing values to calculate pN/pS ratio (NA, Inf)
pNpS_sp_021 <- pNpS_sp_021[!is.na(pNpS_sp_021$pNpS_gene_consensus) & pNpS_sp_021$pNpS_gene_consensus != "Inf" & pNpS_sp_021$pNpS_gene_consensus != 0,]

length(unique(pNpS_sp_021$corresponding_gene_call)) # 3829 genes left after removal
str(pNpS_sp_021)

# Remove samples in which the species was not present
samples_presence <- c("ABBA_M12","ABBA_M13","ABBA_M14","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05")
pNpS_sp_021 <- pNpS_sp_021[pNpS_sp_021$sample_id %in% samples_presence,]
pNpS_sp_021 <- droplevels(pNpS_sp_021)

# Add host species factor
pNpS_sp_021$host_species <- as.factor(substr(pNpS_sp_021$sample_id, 1, 4))
pNpS_sp_021$host_species <- factor(pNpS_sp_021$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
str(pNpS_sp_021)

# Add clade factor
pNpS_sp_021$plant_clade <- factor(
  ifelse(grepl("^(ACSA|FAGR|COCO)", pNpS_sp_021$sample_id),
         "angiosperm",
         ifelse(grepl("^(ABBA|THOC)", pNpS_sp_021$sample_id),
                "gymnosperm",
                NA))
)
str(pNpS_sp_021)

# Add plant type factor
pNpS_sp_021$plant_type <- factor(
  ifelse(grepl("^(ACSA|FAGR|ABBA|THOC)", pNpS_sp_021$sample_id),
         "tree",
         ifelse(grepl("^(COCO)", pNpS_sp_021$sample_id),
                "shrub",
                NA))
)
str(pNpS_sp_021)
```

## Distribution of pN/pS measurements across all genes and all samples
```{r}
hist(pNpS_sp_021$pNpS_gene_consensus, breaks = 1000000)
hist(log(pNpS_sp_021$pNpS_gene_consensus), breaks = 10000)
```

## Add core/accessory
```{r}
colnames(pNpS_sp_021)[1] <- "gene_callers_id"
data.pangenome <- readRDS("../rds/data.pangenome.rds")
pNpS_sp_021 <- merge(pNpS_sp_021, data.pangenome[,c(3:4)], by = "gene_callers_id")
pNpS_sp_021$type <- factor(x = pNpS_sp_021$type, levels = c("core","accessory"))
pNpS_sp_021 <- pNpS_sp_021[order(pNpS_sp_021$type, -pNpS_sp_021$pNpS_gene_consensus), ]
pNpS_sp_021$gene_callers_id <- factor(x = pNpS_sp_021$gene_callers_id, levels = unique(pNpS_sp_021$gene_callers_id))
pNpS_sp_021$sample_id <- factor(x = pNpS_sp_021$sample, levels = c("ABBA_M12","ABBA_M13","ABBA_M14","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M21","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
str(pNpS_sp_021)
```

## Genes under positive selection

### Data prep
```{r}
# Number of unique gene calls for which at least in one sample the pN/pS ratio > 1 (across all samples and host species)
length(unique(pNpS_sp_021[pNpS_sp_021$pNpS_gene_consensus > 1,]$gene_callers_id))

# Make a vector of all genes that have at least one positive pN/pS value in a sample (any host)
pos.sel.genes <- unique(pNpS_sp_021[pNpS_sp_021$pNpS_gene_consensus > 1,]$gene_callers_id)
pos.sel.genes <- droplevels(pos.sel.genes)
```

```{r}
# Create pNpS dataframe with only genes that have at least one positive value
pNpS_sp_021.pos <- pNpS_sp_021[pNpS_sp_021$gene_callers_id %in% pos.sel.genes,]
pNpS_sp_021.pos <- droplevels(pNpS_sp_021.pos)

# Create a table to know in how many samples of each host_species there is variability info on a given gene
freq.table <- as.data.frame(table(pNpS_sp_021.pos$gene_callers_id, pNpS_sp_021.pos$host_species))
colnames(freq.table) <- c("gene_callers_id","host_species","num_samples")

# Keep only gene-host info for hosts that have at least 3 samples for the given gene
freq.table <- freq.table[freq.table$num_samples >= 3,]
freq.table <- droplevels(freq.table)

# Separate freq.table per host species
freq.table.ABBA <- freq.table[freq.table$host_species == "ABBA",]
freq.table.ABBA <- droplevels(freq.table.ABBA)
freq.table.ACSA <- freq.table[freq.table$host_species == "ACSA",]
freq.table.ACSA <- droplevels(freq.table.ACSA)
freq.table.COCO <- freq.table[freq.table$host_species == "COCO",]
freq.table.COCO <- droplevels(freq.table.COCO)
freq.table.FAGR <- freq.table[freq.table$host_species == "FAGR",]
freq.table.FAGR <- droplevels(freq.table.FAGR)
freq.table.THOC <- freq.table[freq.table$host_species == "THOC",]
freq.table.THOC <- droplevels(freq.table.THOC)
```

### Confidence interval tests

We will keep only genes with a mean log(pN/pS(gene)) > 0 and a lower confidence interval > 0.
```{r}
sig.pos.sel.genes_sp_021 <- data.frame(
  gene_callers_id = character(0),
  mean = numeric(0),
  conf_int_lower = numeric(0),
  conf_int_upper = numeric(0),
  t_statistic = numeric(0),
  p_value = numeric(0),
  adj.p = numeric(0),
  host_species = character(0),
  stringsAsFactors = FALSE
)

host_species <- c("ABBA","THOC","ACSA","FAGR","COCO")

for (host in host_species) {

genes.to.t.test <- unique(freq.table[freq.table$host_species == host,]$gene_callers_id)
genes.to.t.test <- droplevels(genes.to.t.test)  

means.of.genes <- data.frame(
  gene_callers_id = character(0),
  mean_pNpS = numeric(0),
  sd_pNpS = numeric(0)
)

for (gene in genes.to.t.test) {
  mean <- mean(log(pNpS_sp_021[pNpS_sp_021$gene_callers_id == gene & pNpS_sp_021$host_species == host,]$pNpS_gene_consensus))
  sd <- sd(pNpS_sp_021[pNpS_sp_021$gene_callers_id == gene & pNpS_sp_021$host_species == host,]$pNpS_gene_consensus)
  means.of.genes <- rbind(means.of.genes, data.frame(
    gene_callers_id = gene,
    mean_pNpS = mean,
    sd_pNpS = sd))
}

means.of.genes$mean_minus_sd <- means.of.genes$mean - means.of.genes$sd

genes.to.t.test <- as.factor(unique(means.of.genes[means.of.genes$mean_pNpS > 0,]$gene_callers_id))
genes.to.t.test <- droplevels(genes.to.t.test)

t_test_results <- data.frame(
  gene_callers_id = character(0),
  mean_pNpS = numeric(0),
  log_mean_pNpS = numeric(0),
  conf_int_lower = numeric(0),
  conf_int_upper = numeric(0),
  t_statistic = numeric(0),
  p_value = numeric(0),
  stringsAsFactors = FALSE
)

for (gene in genes.to.t.test) {
  # Perform the t-test for the current level
  t_test_result <- t.test(log(pNpS_sp_021[pNpS_sp_021$gene_callers_id == gene 
                                      & pNpS_sp_021$host_species == host,]$pNpS_gene_consensus), mu = 0, alternative = "greater")
  # Store the result in the data frame
  t_test_results <- rbind(t_test_results, data.frame(
    gene_callers_id = gene,
    mean_pNpS = exp(t_test_result[["estimate"]][["mean of x"]]),
    log_mean_pNpS = t_test_result[["estimate"]][["mean of x"]],
    conf_int_lower = t_test_result$conf.int[1],
    conf_int_upper = t_test_result$conf.int[2],
    t_statistic = t_test_result$statistic,
    p_value = t_test_result$p.value
  ))
}


t_test_results$adj.p <- p.adjust(t_test_results$p_value, method = "BH") # Benjamini-Hochberg adjustment
t_test_results.sig <- t_test_results[t_test_results$conf_int_lower > 0,]

# Add host species column
t_test_results.sig$host_species <- rep(host, nrow(t_test_results.sig))

# Add to final dataframe
sig.pos.sel.genes_sp_021 <- rbind(sig.pos.sel.genes_sp_021,t_test_results.sig)

}

rownames(sig.pos.sel.genes_sp_021) <- NULL
```

### Functions
```{r}
data.fct.cov <- readRDS("../rds/data.fct.cov.rds")
sig.pos.sel.genes_sp_021.fct <- merge(sig.pos.sel.genes_sp_021, data.fct.cov, by = "gene_callers_id")
length(unique(sig.pos.sel.genes_sp_021.fct$gene_callers_id))
# 6 unique positively selected genes. 6 "gene-host" combinations, including 6 THOC
# All accessory genes
# 1 unique ANNOTATED gene
```

### DAGCs?
```{r}
DAGs.functions.all.info <- read_delim("../data/DAGCs_fully_annotated.csv", delim = ",")
pos.sel.DAGCs_sp_021 <- DAGs.functions.all.info[DAGs.functions.all.info$gene_cluster_id %in% sig.pos.sel.genes_sp_021.fct$gene_cluster_id,]
# no DAGC

sig.pos.sel.genes_sp_021.fct$DAGCs <- ifelse(
  sig.pos.sel.genes_sp_021.fct$gene_cluster_id %in% pos.sel.DAGCs_sp_021$gene_cluster_id,
  paste(pos.sel.DAGCs_sp_021$focal, pos.sel.DAGCs_sp_021$reference, sep = "_vs_"),
  "non-DAGC"
)

sig.pos.sel.genes_sp_021.fct <- sig.pos.sel.genes_sp_021.fct[,c(1:13,58,14:57)]
```

### Save data
```{r}
#saveRDS(sig.pos.sel.genes_sp_021.fct, "../rds/sig.pos.sel.genes_sp_021.fct.rds")
sig.pos.sel.genes_sp_021.fct <- readRDS("../rds/sig.pos.sel.genes_sp_021.fct.rds")
```

-################################################################################

# M_sp_022

## Import data

```{r}
pNpS_sp_022 <- read.table("../data/anvio_outputs/pNpS_sp_022-03/pNpS.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pN_sp_022 <- read.table("../data/anvio_outputs/pNpS_sp_022-03/pN.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pS_sp_022 <- read.table("../data/anvio_outputs/pNpS_sp_022-03/pS.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pNpS_sp_022$pN <- pN_sp_022$pN_gene_consensus
pNpS_sp_022$pS <- pS_sp_022$pS_gene_consensus
length(which(pNpS_sp_022$pNpS_gene_consensus == 0)) #60
str(pNpS_sp_022)
```

## Edit dataframe
```{r}
# Remove rows where there was missing values to calculate pN/pS ratio (NA, Inf)
pNpS_sp_022 <- pNpS_sp_022[!is.na(pNpS_sp_022$pNpS_gene_consensus) & pNpS_sp_022$pNpS_gene_consensus != "Inf" & pNpS_sp_022$pNpS_gene_consensus != 0,]

length(unique(pNpS_sp_022$corresponding_gene_call)) # 1308 genes left after removal
str(pNpS_sp_022)

# Remove samples in which the species was not present
samples_presence <- c("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05")
pNpS_sp_022 <- pNpS_sp_022[pNpS_sp_022$sample_id %in% samples_presence,]
pNpS_sp_022 <- droplevels(pNpS_sp_022)

# Add host species factor
pNpS_sp_022$host_species <- as.factor(substr(pNpS_sp_022$sample_id, 1, 4))
pNpS_sp_022$host_species <- factor(pNpS_sp_022$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
str(pNpS_sp_022)

# Add clade factor
pNpS_sp_022$plant_clade <- factor(
  ifelse(grepl("^(ACSA|FAGR|COCO)", pNpS_sp_022$sample_id),
         "angiosperm",
         ifelse(grepl("^(ABBA|THOC)", pNpS_sp_022$sample_id),
                "gymnosperm",
                NA))
)
str(pNpS_sp_022)

# Add plant type factor
pNpS_sp_022$plant_type <- factor(
  ifelse(grepl("^(ACSA|FAGR|ABBA|THOC)", pNpS_sp_022$sample_id),
         "tree",
         ifelse(grepl("^(COCO)", pNpS_sp_022$sample_id),
                "shrub",
                NA))
)
str(pNpS_sp_022)
```

## Distribution of pN/pS measurements across all genes and all samples
```{r}
hist(pNpS_sp_022$pNpS_gene_consensus, breaks = 1000000)
hist(log(pNpS_sp_022$pNpS_gene_consensus), breaks = 10000)
```

## Add core/accessory
```{r}
colnames(pNpS_sp_022)[1] <- "gene_callers_id"
data.pangenome <- readRDS("../rds/data.pangenome.rds")
pNpS_sp_022 <- merge(pNpS_sp_022, data.pangenome[,c(3:4)], by = "gene_callers_id")
pNpS_sp_022$type <- factor(x = pNpS_sp_022$type, levels = c("core","accessory"))
pNpS_sp_022 <- pNpS_sp_022[order(pNpS_sp_022$type, -pNpS_sp_022$pNpS_gene_consensus), ]
pNpS_sp_022$gene_callers_id <- factor(x = pNpS_sp_022$gene_callers_id, levels = unique(pNpS_sp_022$gene_callers_id))
pNpS_sp_022$sample_id <- factor(x = pNpS_sp_022$sample, levels = c("ABBA_M11","ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M06","THOC_M07","THOC_M08","THOC_M09","THOC_M10","ACSA_M22","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M19","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
str(pNpS_sp_022)
```

## Genes under positive selection

### Data prep
```{r}
# Number of unique gene calls for which at least in one sample the pN/pS ratio > 1 (across all samples and host species)
length(unique(pNpS_sp_022[pNpS_sp_022$pNpS_gene_consensus > 1,]$gene_callers_id))

# Make a vector of all genes that have at least one positive pN/pS value in a sample (any host)
pos.sel.genes <- unique(pNpS_sp_022[pNpS_sp_022$pNpS_gene_consensus > 1,]$gene_callers_id)
pos.sel.genes <- droplevels(pos.sel.genes)
```

```{r}
# Create pNpS dataframe with only genes that have at least one positive value
pNpS_sp_022.pos <- pNpS_sp_022[pNpS_sp_022$gene_callers_id %in% pos.sel.genes,]
pNpS_sp_022.pos <- droplevels(pNpS_sp_022.pos)

# Create a table to know in how many samples of each host_species there is variability info on a given gene
freq.table <- as.data.frame(table(pNpS_sp_022.pos$gene_callers_id, pNpS_sp_022.pos$host_species))
colnames(freq.table) <- c("gene_callers_id","host_species","num_samples")

# Keep only gene-host info for hosts that have at least 3 samples for the given gene
freq.table <- freq.table[freq.table$num_samples >= 3,]
freq.table <- droplevels(freq.table)

# Separate freq.table per host species
freq.table.ABBA <- freq.table[freq.table$host_species == "ABBA",]
freq.table.ABBA <- droplevels(freq.table.ABBA)
freq.table.ACSA <- freq.table[freq.table$host_species == "ACSA",]
freq.table.ACSA <- droplevels(freq.table.ACSA)
freq.table.COCO <- freq.table[freq.table$host_species == "COCO",]
freq.table.COCO <- droplevels(freq.table.COCO)
freq.table.FAGR <- freq.table[freq.table$host_species == "FAGR",]
freq.table.FAGR <- droplevels(freq.table.FAGR)
freq.table.THOC <- freq.table[freq.table$host_species == "THOC",]
freq.table.THOC <- droplevels(freq.table.THOC)
```


### Confidence interval tests

We will keep only genes with a mean log(pN/pS(gene)) > 0 and a lower confidence interval > 0.
```{r}
sig.pos.sel.genes_sp_022 <- data.frame(
  gene_callers_id = character(0),
  mean = numeric(0),
  conf_int_lower = numeric(0),
  conf_int_upper = numeric(0),
  t_statistic = numeric(0),
  p_value = numeric(0),
  adj.p = numeric(0),
  host_species = character(0),
  stringsAsFactors = FALSE
)

host_species <- c("ABBA","THOC","ACSA","FAGR","COCO")

for (host in host_species) {

genes.to.t.test <- unique(freq.table[freq.table$host_species == host,]$gene_callers_id)
genes.to.t.test <- droplevels(genes.to.t.test)  

means.of.genes <- data.frame(
  gene_callers_id = character(0),
  mean_pNpS = numeric(0),
  sd_pNpS = numeric(0)
)

for (gene in genes.to.t.test) {
  mean <- mean(log(pNpS_sp_022[pNpS_sp_022$gene_callers_id == gene & pNpS_sp_022$host_species == host,]$pNpS_gene_consensus))
  sd <- sd(pNpS_sp_022[pNpS_sp_022$gene_callers_id == gene & pNpS_sp_022$host_species == host,]$pNpS_gene_consensus)
  means.of.genes <- rbind(means.of.genes, data.frame(
    gene_callers_id = gene,
    mean_pNpS = mean,
    sd_pNpS = sd))
}

means.of.genes$mean_minus_sd <- means.of.genes$mean - means.of.genes$sd

genes.to.t.test <- as.factor(unique(means.of.genes[means.of.genes$mean_pNpS > 0,]$gene_callers_id))
genes.to.t.test <- droplevels(genes.to.t.test)

t_test_results <- data.frame(
  gene_callers_id = character(0),
  mean_pNpS = numeric(0),
  log_mean_pNpS = numeric(0),
  conf_int_lower = numeric(0),
  conf_int_upper = numeric(0),
  t_statistic = numeric(0),
  p_value = numeric(0),
  stringsAsFactors = FALSE
)

for (gene in genes.to.t.test) {
  # Perform the t-test for the current level
  t_test_result <- t.test(log(pNpS_sp_022[pNpS_sp_022$gene_callers_id == gene 
                                      & pNpS_sp_022$host_species == host,]$pNpS_gene_consensus), mu = 0, alternative = "greater")
  # Store the result in the data frame
  t_test_results <- rbind(t_test_results, data.frame(
    gene_callers_id = gene,
    mean_pNpS = exp(t_test_result[["estimate"]][["mean of x"]]),
    log_mean_pNpS = t_test_result[["estimate"]][["mean of x"]],
    conf_int_lower = t_test_result$conf.int[1],
    conf_int_upper = t_test_result$conf.int[2],
    t_statistic = t_test_result$statistic,
    p_value = t_test_result$p.value
  ))
}


t_test_results$adj.p <- p.adjust(t_test_results$p_value, method = "BH") # Benjamini-Hochberg adjustment
t_test_results.sig <- t_test_results[t_test_results$conf_int_lower > 0,]

# Add host species column
t_test_results.sig$host_species <- rep(host, nrow(t_test_results.sig))

# Add to final dataframe
sig.pos.sel.genes_sp_022 <- rbind(sig.pos.sel.genes_sp_022,t_test_results.sig)

}

rownames(sig.pos.sel.genes_sp_022) <- NULL
```


NO positively selected genes detected within M. sp. 022 population.

-################################################################################

# M_sp_024

## Import data
```{r}
pNpS_sp_024 <- read.table("../data/anvio_outputs/pNpS_sp_024-03/pNpS.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pN_sp_024 <- read.table("../data/anvio_outputs/pNpS_sp_024-03/pN.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pS_sp_024 <- read.table("../data/anvio_outputs/pNpS_sp_024-03/pS.txt", header = T, sep = "\t", colClasses = c("NULL", "factor", "factor", "numeric"))
pNpS_sp_024$pN <- pN_sp_024$pN_gene_consensus
pNpS_sp_024$pS <- pS_sp_024$pS_gene_consensus
length(which(pNpS_sp_024$pNpS_gene_consensus == 0)) #263
str(pNpS_sp_024)
```

## Edit dataframe
```{r}
# Remove rows where there was missing values to calculate pN/pS ratio (NA, Inf)
pNpS_sp_024 <- pNpS_sp_024[!is.na(pNpS_sp_024$pNpS_gene_consensus) & pNpS_sp_024$pNpS_gene_consensus != "Inf" & pNpS_sp_024$pNpS_gene_consensus != 0,]

length(unique(pNpS_sp_024$corresponding_gene_call)) # 4802 genes left after removal
str(pNpS_sp_024)

# Remove samples in which the species was not present
samples_presence <- c("ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M08","THOC_M09","THOC_M10","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05")
pNpS_sp_024 <- pNpS_sp_024[pNpS_sp_024$sample_id %in% samples_presence,]
pNpS_sp_024 <- droplevels(pNpS_sp_024)

# Add host species factor
pNpS_sp_024$host_species <- as.factor(substr(pNpS_sp_024$sample_id, 1, 4))
pNpS_sp_024$host_species <- factor(pNpS_sp_024$host_species, levels = c("ABBA","THOC","ACSA","FAGR","COCO"))
str(pNpS_sp_024)

# Add clade factor
pNpS_sp_024$plant_clade <- factor(
  ifelse(grepl("^(ACSA|FAGR|COCO)", pNpS_sp_024$sample_id),
         "angiosperm",
         ifelse(grepl("^(ABBA|THOC)", pNpS_sp_024$sample_id),
                "gymnosperm",
                NA))
)
str(pNpS_sp_024)

# Add plant type factor
pNpS_sp_024$plant_type <- factor(
  ifelse(grepl("^(ACSA|FAGR|ABBA|THOC)", pNpS_sp_024$sample_id),
         "tree",
         ifelse(grepl("^(COCO)", pNpS_sp_024$sample_id),
                "shrub",
                NA))
)
str(pNpS_sp_024)
```

## Distribution of pN/pS measurements across all genes and all samples
```{r}
hist(pNpS_sp_024$pNpS_gene_consensus, breaks = 1000000)
hist(log(pNpS_sp_024$pNpS_gene_consensus), breaks = 10000)
```

## Add core/accessory
```{r}
colnames(pNpS_sp_024)[1] <- "gene_callers_id"
data.pangenome <- readRDS("../rds/data.pangenome.rds")
pNpS_sp_024 <- merge(pNpS_sp_024, data.pangenome[,c(3:4)], by = "gene_callers_id")
pNpS_sp_024$type <- factor(x = pNpS_sp_024$type, levels = c("core","accessory"))
pNpS_sp_024 <- pNpS_sp_024[order(pNpS_sp_024$type, -pNpS_sp_024$pNpS_gene_consensus), ]
pNpS_sp_024$gene_callers_id <- factor(x = pNpS_sp_024$gene_callers_id, levels = unique(pNpS_sp_024$gene_callers_id))
pNpS_sp_024$sample_id <- factor(x = pNpS_sp_024$sample, levels = c("ABBA_M12","ABBA_M13","ABBA_M14","ABBA_M15","THOC_M08","THOC_M09","THOC_M10","ACSA_M24","ACSA_M25","FAGR_M16","FAGR_M17","FAGR_M20","COCO_M01","COCO_M02","COCO_M03","COCO_M04","COCO_M05"))
str(pNpS_sp_024)
```

## Genes under positive selection

### Data prep
```{r}
# Number of unique gene calls for which at least in one sample the pN/pS ratio > 1 (across all samples and host species)
length(unique(pNpS_sp_024[pNpS_sp_024$pNpS_gene_consensus > 1,]$gene_callers_id))

# Make a vector of all genes that have at least one positive pN/pS value in a sample (any host)
pos.sel.genes <- unique(pNpS_sp_024[pNpS_sp_024$pNpS_gene_consensus > 1,]$gene_callers_id)
pos.sel.genes <- droplevels(pos.sel.genes)
```

```{r}
# Create pNpS dataframe with only genes that have at least one positive value
pNpS_sp_024.pos <- pNpS_sp_024[pNpS_sp_024$gene_callers_id %in% pos.sel.genes,]
pNpS_sp_024.pos <- droplevels(pNpS_sp_024.pos)

# Create a table to know in how many samples of each host_species there is variability info on a given gene
freq.table <- as.data.frame(table(pNpS_sp_024.pos$gene_callers_id, pNpS_sp_024.pos$host_species))
colnames(freq.table) <- c("gene_callers_id","host_species","num_samples")

# Keep only gene-host info for hosts that have at least 3 samples for the given gene
freq.table <- freq.table[freq.table$num_samples >= 3,]
freq.table <- droplevels(freq.table)

# Separate freq.table per host species
freq.table.ABBA <- freq.table[freq.table$host_species == "ABBA",]
freq.table.ABBA <- droplevels(freq.table.ABBA)
freq.table.ACSA <- freq.table[freq.table$host_species == "ACSA",]
freq.table.ACSA <- droplevels(freq.table.ACSA)
freq.table.COCO <- freq.table[freq.table$host_species == "COCO",]
freq.table.COCO <- droplevels(freq.table.COCO)
freq.table.FAGR <- freq.table[freq.table$host_species == "FAGR",]
freq.table.FAGR <- droplevels(freq.table.FAGR)
freq.table.THOC <- freq.table[freq.table$host_species == "THOC",]
freq.table.THOC <- droplevels(freq.table.THOC)
```

### Confidence interval tests

We will keep only genes with a mean log(pN/pS(gene)) > 0 and a lower confidence interval > 0.
```{r}
sig.pos.sel.genes_sp_024 <- data.frame(
  gene_callers_id = character(0),
  mean = numeric(0),
  conf_int_lower = numeric(0),
  conf_int_upper = numeric(0),
  t_statistic = numeric(0),
  p_value = numeric(0),
  adj.p = numeric(0),
  host_species = character(0),
  stringsAsFactors = FALSE
)

host_species <- c("ABBA","THOC","ACSA","FAGR","COCO")

for (host in host_species) {

genes.to.t.test <- unique(freq.table[freq.table$host_species == host,]$gene_callers_id)
genes.to.t.test <- droplevels(genes.to.t.test)  

means.of.genes <- data.frame(
  gene_callers_id = character(0),
  mean_pNpS = numeric(0),
  sd_pNpS = numeric(0)
)

for (gene in genes.to.t.test) {
  mean <- mean(log(pNpS_sp_024[pNpS_sp_024$gene_callers_id == gene & pNpS_sp_024$host_species == host,]$pNpS_gene_consensus))
  sd <- sd(pNpS_sp_024[pNpS_sp_024$gene_callers_id == gene & pNpS_sp_024$host_species == host,]$pNpS_gene_consensus)
  means.of.genes <- rbind(means.of.genes, data.frame(
    gene_callers_id = gene,
    mean_pNpS = mean,
    sd_pNpS = sd))
}

means.of.genes$mean_minus_sd <- means.of.genes$mean - means.of.genes$sd

genes.to.t.test <- as.factor(unique(means.of.genes[means.of.genes$mean_pNpS > 0,]$gene_callers_id))
genes.to.t.test <- droplevels(genes.to.t.test)

t_test_results <- data.frame(
  gene_callers_id = character(0),
  mean_pNpS = numeric(0),
  log_mean_pNpS = numeric(0),
  conf_int_lower = numeric(0),
  conf_int_upper = numeric(0),
  t_statistic = numeric(0),
  p_value = numeric(0),
  stringsAsFactors = FALSE
)

for (gene in genes.to.t.test) {
  # Perform the t-test for the current level
  t_test_result <- t.test(log(pNpS_sp_024[pNpS_sp_024$gene_callers_id == gene 
                                      & pNpS_sp_024$host_species == host,]$pNpS_gene_consensus), mu = 0, alternative = "greater")
  # Store the result in the data frame
  t_test_results <- rbind(t_test_results, data.frame(
    gene_callers_id = gene,
    mean_pNpS = exp(t_test_result[["estimate"]][["mean of x"]]),
    log_mean_pNpS = t_test_result[["estimate"]][["mean of x"]],
    conf_int_lower = t_test_result$conf.int[1],
    conf_int_upper = t_test_result$conf.int[2],
    t_statistic = t_test_result$statistic,
    p_value = t_test_result$p.value
  ))
}


t_test_results$adj.p <- p.adjust(t_test_results$p_value, method = "BH") # Benjamini-Hochberg adjustment
t_test_results.sig <- t_test_results[t_test_results$conf_int_lower > 0,]

# Add host species column
t_test_results.sig$host_species <- rep(host, nrow(t_test_results.sig))

# Add to final dataframe
sig.pos.sel.genes_sp_024 <- rbind(sig.pos.sel.genes_sp_024,t_test_results.sig)

}

rownames(sig.pos.sel.genes_sp_024) <- NULL
```

### Functions
```{r}
data.fct.cov <- readRDS("../rds/data.fct.cov.rds")
sig.pos.sel.genes_sp_024.fct <- merge(sig.pos.sel.genes_sp_024, data.fct.cov, by = "gene_callers_id")
length(unique(sig.pos.sel.genes_sp_024.fct$gene_callers_id))
# 1 unique positively selected genes. 1 "gene-host" combinations, including 1 COCO
# accessory gene
# 0 unique ANNOTATED genes
```

### DAGCs?
```{r}
DAGs.functions.all.info <- read_delim("../data/DAGCs_fully_annotated.csv", delim = ",")
pos.sel.DAGCs_sp_024 <- DAGs.functions.all.info[DAGs.functions.all.info$gene_cluster_id %in% sig.pos.sel.genes_sp_024.fct$gene_cluster_id,]
# includes 0 DAGC

sig.pos.sel.genes_sp_024.fct$DAGCs <- ifelse(
  sig.pos.sel.genes_sp_024.fct$gene_cluster_id %in% pos.sel.DAGCs_sp_024$gene_cluster_id,
  paste(pos.sel.DAGCs_sp_024$focal, pos.sel.DAGCs_sp_024$reference, sep = "_vs_"),
  "non-DAGC"
)

sig.pos.sel.genes_sp_024.fct <- sig.pos.sel.genes_sp_024.fct[,c(1:13,58,14:57)]
```

### Save data
```{r}
#saveRDS(sig.pos.sel.genes_sp_024.fct, "../rds/sig.pos.sel.genes_sp_024.fct.rds")
sig.pos.sel.genes_sp_024.fct <- readRDS("../rds/sig.pos.sel.genes_sp_024.fct.rds")
```

-################################################################################

# Supplementary File D

Create table for positively selected genes

## Import data
```{r}
sig.pos.sel.genes_sp_018.fct <- readRDS("../rds/sig.pos.sel.genes_sp_018.fct.rds")
sig.pos.sel.genes_sp_021.fct <- readRDS("../rds/sig.pos.sel.genes_sp_021.fct.rds")
sig.pos.sel.genes_sp_024.fct <- readRDS("../rds/sig.pos.sel.genes_sp_024.fct.rds")
```

## Edit table
```{r}
all(colnames(sig.pos.sel.genes_sp_018.fct)==colnames(sig.pos.sel.genes_sp_021.fct))
all(colnames(sig.pos.sel.genes_sp_018.fct)==colnames(sig.pos.sel.genes_sp_024.fct))

all.pos.sel.genes <- rbind(sig.pos.sel.genes_sp_018.fct, 
                           sig.pos.sel.genes_sp_021.fct, 
                           sig.pos.sel.genes_sp_024.fct)
colnames(all.pos.sel.genes)[4] <- "CI_lower.log"
all.pos.sel.genes <- all.pos.sel.genes[,c(1,10,12,11,9,14,13,2,3,4,16,18,20,22,24,26,28,30)]

all.pos.sel.genes$host_species <- factor(x = all.pos.sel.genes$host_species, levels = c("THOC","FAGR","COCO"))
all.pos.sel.genes <- all.pos.sel.genes[order(all.pos.sel.genes$genome_name, all.pos.sel.genes$host_species, -all.pos.sel.genes$mean_pNpS),]

saveRDS(all.pos.sel.genes,"../rds/all.pos.sel.genes.rds")
```

## Write table
```{r}
#write_csv(all.pos.sel.genes, file = "../supplementary_files/Suppl_file_D-pNpS.csv")
```

-################################################################################

# Combine pN/pS data for the 4 methylo species

## Edit tables
```{r}
pNpS_sp_018$methylo_sp <- "M. sp. 018"
pNpS_sp_021$methylo_sp <- "M. sp. 021"
pNpS_sp_022$methylo_sp <- "M. sp. 022"
pNpS_sp_024$methylo_sp <- "M. sp. 024"
pNpS_all_sp <- rbind(pNpS_sp_018,pNpS_sp_021,pNpS_sp_022,pNpS_sp_024)
pNpS_all_sp$methylo_sp <- as.factor(pNpS_all_sp$methylo_sp)
levels(pNpS_all_sp$type)[levels(pNpS_all_sp$type) == "core"] <- "Core genes"
levels(pNpS_all_sp$type)[levels(pNpS_all_sp$type) == "accessory"] <- "Accessory genes"
pNpS_all_sp$log_pNpS <- log(pNpS_all_sp$pNpS_gene_consensus)
contrasts(pNpS_all_sp$type) <- contr.sum
contrasts(pNpS_all_sp$methylo_sp) <- contr.sum
contrasts(pNpS_all_sp$host_species) <- contr.sum
str(pNpS_all_sp) # 14,566 genes
```

## Make second table with COG category
```{r}
data.fct.cov <- readRDS("../rds/data.fct.cov.rds")
pNpS_all_sp <- merge(pNpS_all_sp, data.fct.cov[c(2,7)], by = "gene_callers_id")

# For every gene call that have multiple (n) COG categories, make n rows, one row per category.
pNpS_all_sp.temp <- pNpS_all_sp
pNpS_all_sp.temp$COG20_CATEGORY <- as.character(pNpS_all_sp.temp$COG20_CATEGORY)
# Add row number column
df_split1 <- pNpS_all_sp.temp %>%
  rowid_to_column()  # Add a column representing the original row number
# Split rows based on "|" and unnest
df_split2 <- df_split1 %>%
  mutate(COG20_CATEGORY = strsplit(COG20_CATEGORY, "\\|")) %>%
  unnest(cols = COG20_CATEGORY)
# Remove duplicates within rows
df_split3 <- df_split2 %>%
  distinct(gene_callers_id, sample_id, COG20_CATEGORY, .keep_all = TRUE) 
# Remove rowid column
df_split4 <- df_split3 %>%
  select(-rowid)
pNpS_all_sp.multi <- as.data.frame(df_split4)
rm(pNpS_all_sp.temp, df_split1, df_split2, df_split3, df_split4)
# arrange vectors
pNpS_all_sp.multi$COG20_CATEGORY <- as.factor(pNpS_all_sp.multi$COG20_CATEGORY)
contrasts(pNpS_all_sp.multi$COG20_CATEGORY) <- contr.sum
pNpS_all_sp.multi <- droplevels(pNpS_all_sp.multi)
str(pNpS_all_sp.multi) ## 11,711 genes
```

## Linear Model without COG category
```{r}
lm <- lm(log(pNpS_gene_consensus) ~ type * methylo_sp * host_species, data = pNpS_all_sp) # report
summary(lm) # report p and adj.R2
Anova(lm, type = 3) # report supp mat
joint_tests(lm) # corresponds to type III ANOVA

plot(lm)
hist(residuals(lm))

# Estimated marginal means and contrasts
emm.type <-emmeans(lm, specs = pairwise ~ type, type = "response")
emm.type # report supp mat

emm.meth_sp <- emmeans(lm, specs = pairwise ~ methylo_sp, adjust = "bonferroni", type = "response")
emm.meth_sp # report supp mat

emm.host_sp <- emmeans(lm, specs = pairwise ~ host_species, adjust = "bonferroni", type = "response")
emm.host_sp # report supp mat

pred.type <- ggemmeans(lm, terms = c("type"))
plot(pred.type)
#pred.type

pred.meth_sp <- ggemmeans(lm, terms = c("methylo_sp"))
plot(pred.meth_sp)
#pred.meth_sp

pred.host_sp <- ggemmeans(lm, terms = c("host_species"))
plot(pred.host_sp)
#pred.host_sp 

pred <- ggemmeans(lm, terms = c("methylo_sp", "host_species", "type"), back.transform = TRUE, interval = "confidence")
#pred
plot(pred)
```

### Plot
```{r}
pred <- ggemmeans(lm, terms = c("methylo_sp", "host_species", "type"), back.transform = TRUE, interval = "confidence")
plot(pred)

mod.plot <- plot(pred, 
                 #show_data = T, limit_range = F, dot_alpha = 0.01, jitter = 0.2,
                 dodge = 0.5)

mod.plot +
  xlab(expression(bolditalic("Methylobacterium") ~ bold("species"))) +
  ylab(expression(bold("Model-predicted") ~ bolditalic("pN/pS")^bold("(gene)"))) +
#  ylim(0, 0.25) +
  scale_color_manual(values=c("#63AFF9","#1EA670","#CA0E78","#F39635","#6104BB"), name = expression(bold("Host species")), labels = c("ABBA" = expression(italic("A. balsamea")), "THOC" = expression(italic("T. occidentalis")),"ACSA" = expression(italic("A. saccharum")), "FAGR" = expression(italic("F. grandifolia")), "COCO" = expression(italic("C. cornuta")))) +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      strip.background = element_rect(fill = "gray25", color = "black"),  # black fill with white border
      strip.text = element_text(face = "bold", color = "white", size = 11), # bold white text
      legend.justification = c(0.5, 0),  
      legend.text = element_text(size = 9),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      panel.grid.major.y = element_line(),
      panel.grid.minor.y = element_line(),
      plot.title = element_blank(),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0, face = "italic"),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2)
      )
ggsave("../figures/predicted_pNpS.tiff", width = 9.5, height = 4)
```

## relation between pN/pS and gene/genome coverage

### Data prep
```{r}
genome.mean.cov <- readRDS("../rds/genome.mean.cov.rds")
#genome.std.cov <- readRDS("../rds/genome.std.cov.rds")
genome.mean.cov.sub <- genome.mean.cov[rownames(genome.mean.cov) %in% c("M_sp_018", "M_sp_021", "M_sp_022", "M_sp_024"),]
genome.cov <- matrix2sample(genome.mean.cov.sub)
colnames(genome.cov) <- c("methylo_sp","genome_cov","sample_id")
levels(genome.cov$methylo_sp) <- gsub("_", ". ", levels(genome.cov$methylo_sp))

data.fct.cov <- readRDS("../rds/data.fct.cov.rds")
data.fct.cov.sub <- data.fct.cov[,c(2,42:66)]
rownames(data.fct.cov.sub) <- data.fct.cov.sub$gene_callers_id
data.fct.cov.sub <- data.fct.cov.sub[,-1]
gene.cov <- matrix2sample(data.fct.cov.sub)
colnames(gene.cov) <- c("gene_callers_id","gene_cov","sample_id")

pNpS_all_sp_cov <- merge(pNpS_all_sp, gene.cov, by = c("gene_callers_id","sample_id"))
pNpS_all_sp_cov <- merge(pNpS_all_sp_cov, genome.cov, by = c("methylo_sp","sample_id"))
```

### Model and plot - genome coverage
```{r}
lm.cov <- lm(log(pNpS_gene_consensus) ~ type * methylo_sp * genome_cov, data = pNpS_all_sp_cov) # report
summary(lm.cov) # report p and adj.R2
Anova(lm.cov, type = 3) # report supp mat

plot(lm.cov)
hist(residuals(lm.cov))

lm.cov.2 <- lm(log(pNpS_gene_consensus) ~ genome_cov, data = pNpS_all_sp_cov) # report
summary(lm.cov.2)
plot(log(pNpS_gene_consensus) ~ genome_cov, data = pNpS_all_sp_cov)

trends <- emtrends(lm.cov, specs = pairwise ~ type * methylo_sp, var = "genome_cov", adjust = "bonferroni", type = "response")
trends # report supp mat

pred.cov <- ggemmeans(lm.cov, terms = c("genome_cov [0.87:81.75]", "methylo_sp", "type"), back.transform = TRUE, interval = "confidence")
plot(pred.cov)

mod.cov.plot <- plot(pred.cov
                 #, show_data = T, limit_range = F, dot_alpha = 0.01, jitter = 0.2,
                 #, dodge = 0.5
                 , limit_range = TRUE 
                   )

mod.cov.plot <- mod.cov.plot +
  xlab(expression(bold("Genome coverage"))) +
  ylab(expression(bold("Model-predicted") ~ bolditalic("pN/pS")^bold("(gene)"))) +
  scale_color_manual(values=c("#E15F4C","#0095C8","#A77B3D","#1A8A00"), name = expression(bold("Methylo.\nspecies")), labels = c("M. sp. 018" = expression(italic("M. sp. 018")), "M. sp. 021" = expression(italic("M. sp. 021")),"M. sp. 022" = expression(italic("M. sp. 022")), "M. sp. 024" = expression(italic("M. sp. 024")))) +
  scale_fill_manual(values=c("#E15F4C","#0095C8","#A77B3D","#1A8A00"))+
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      strip.background = element_rect(fill = "gray25", color = "black"),  # black fill with white border
      strip.text = element_text(face = "bold", color = "white", size = 11), # bold white text
      legend.justification = c(0.5, 0),  
      legend.text = element_text(size = 9),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      panel.grid.major.y = element_line(),
      panel.grid.minor.y = element_line(),
      plot.title = element_blank(),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2)
      )

mod.cov.plot

ggsave("../figures/predicted_pNpS_genome_cov.tiff", width = 9.5, height = 4)
```

### Model and plot - gene coverage
```{r}
lm.cov.gene <- lm(log(pNpS_gene_consensus) ~ type * methylo_sp * gene_cov, data = pNpS_all_sp_cov) # report
summary(lm.cov.gene) # report p and adj.R2
Anova(lm.cov.gene, type = 3) # report supp mat

plot(lm.cov.gene)
hist(residuals(lm.cov.gene))

lm.cov.gene.2 <- lm(log(pNpS_gene_consensus) ~ log(gene_cov), data = pNpS_all_sp_cov) 
cor(log(pNpS_all_sp_cov$pNpS_gene_consensus), log(pNpS_all_sp_cov$gene_cov))
summary(lm.cov.gene.2) # report p and adj.R2
plot(log(pNpS_gene_consensus) ~ gene_cov, data = pNpS_all_sp_cov)

trends.gene <- emtrends(lm.cov.gene, specs = pairwise ~ type * methylo_sp, var = "gene_cov", adjust = "bonferroni", type = "response")
trends.gene # report supp mat

pred.cov.gene <- ggemmeans(lm.cov.gene, terms = c("gene_cov [1.25:297.5]", "methylo_sp", "type"), back.transform = TRUE, interval = "confidence")
plot(pred.cov.gene)

mod.cov.plot.gene <- plot(pred.cov.gene
                 #, show_data = T, limit_range = F, dot_alpha = 0.01, jitter = 0.2,
                 #, dodge = 0.5
                 , limit_range = TRUE 
                   )

mod.cov.plot.gene <- mod.cov.plot.gene +
  xlab(expression(bold("Gene coverage"))) +
  #ylim(0,0.3)+xlim(0,80)+
  ylab(expression(bold("Model-predicted") ~ bolditalic("pN/pS")^bold("(gene)"))) +
  scale_color_manual(values=c("#E15F4C","#0095C8","#A77B3D","#1A8A00"), name = expression(bold("Methylo.\nspecies")), labels = c("M. sp. 018" = expression(italic("M. sp. 018")), "M. sp. 021" = expression(italic("M. sp. 021")),"M. sp. 022" = expression(italic("M. sp. 022")), "M. sp. 024" = expression(italic("M. sp. 024")))) +
  scale_fill_manual(values=c("#E15F4C","#0095C8","#A77B3D","#1A8A00"))+
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      strip.background = element_rect(fill = "gray25", color = "black"),  # black fill with white border
      strip.text = element_text(face = "bold", color = "white", size = 11), # bold white text
      legend.justification = c(0.5, 0),  
      legend.text = element_text(size = 9),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      panel.grid.major.y = element_line(),
      panel.grid.minor.y = element_line(),
      plot.title = element_blank(),
      axis.text.x = element_text(color = "black", size = 11, angle = 0, hjust = 0.5, vjust = 0),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2)
      )

mod.cov.plot.gene

ggsave("../figures/predicted_pNpS_gene_cov.tiff", width = 9.5, height = 4)
```

## Linear Model with COG category - Exploration
```{r}
# Exporation only
lm.COG.1 <- lm(log(pNpS_gene_consensus) ~ COG20_CATEGORY, data = pNpS_all_sp.multi) 

lm.COG.2 <- lm(log(pNpS_gene_consensus) ~ type * COG20_CATEGORY, data = pNpS_all_sp.multi) 

lm.COG.3 <- lm(log(pNpS_gene_consensus) ~ type * host_species * COG20_CATEGORY, data = pNpS_all_sp.multi)

lm.COG.4 <- lm(log(pNpS_gene_consensus) ~ type * host_species * methylo_sp * COG20_CATEGORY, data = pNpS_all_sp.multi) 

lm.COG.5 <- lm(log(pNpS_gene_consensus) ~ host_species * COG20_CATEGORY, data = pNpS_all_sp.multi)

lm.COG.6 <- lm(log(pNpS_gene_consensus) ~ host_species * methylo_sp * COG20_CATEGORY, data = pNpS_all_sp.multi) 

lm.COG.7 <- lm(log(pNpS_gene_consensus) ~ methylo_sp * COG20_CATEGORY, data = pNpS_all_sp.multi) 

lm.COG.8 <- lm(log(pNpS_gene_consensus) ~ type * methylo_sp * COG20_CATEGORY, data = pNpS_all_sp.multi) 

lm.COG.9 <- lm(log(pNpS_gene_consensus) ~ type * host_species * methylo_sp + COG20_CATEGORY, data = pNpS_all_sp.multi) 

lm.COG.10 <- lm(log(pNpS_gene_consensus) ~ type * host_species * methylo_sp + COG20_CATEGORY + type:COG20_CATEGORY, data = pNpS_all_sp.multi) 

lm.COG.11 <- lmer(log(pNpS_gene_consensus) ~ type * COG20_CATEGORY + (1|host_species) + (1|methylo_sp), data = pNpS_all_sp.multi)

lm.COG.12 <- lmer(log(pNpS_gene_consensus) ~ type * COG20_CATEGORY + (1|host_species*methylo_sp), data = pNpS_all_sp.multi)

lm.COG.13 <- lm(log(pNpS_gene_consensus) ~ host_species + methylo_sp + host_species:methylo_sp + type + COG20_CATEGORY + type:COG20_CATEGORY, data = pNpS_all_sp.multi)

lm.COG.14 <- lm(log(pNpS_gene_consensus) ~ type + host_species + methylo_sp + host_species:methylo_sp + type:host_species + type:methylo_sp + type:host_species:methylo_sp + COG20_CATEGORY + type:COG20_CATEGORY, data = pNpS_all_sp.multi)

model.sel(lm,lm.COG.1,lm.COG.2,lm.COG.3,lm.COG.4,lm.COG.5,lm.COG.6,lm.COG.7,lm.COG.8,lm.COG.9,lm.COG.10,lm.COG.11,lm.COG.12,lm.COG.13,lm.COG.14)
```

### Plot - Exploration
```{r}
pred.COG <- ggemmeans(lm.COG.10, terms = c("COG20_CATEGORY", "host_species", "methylo_sp"), back.transform = TRUE, interval = "confidence")

pred.COG <- ggemmeans(lm.COG.10, terms = "COG20_CATEGORY")

pred.COG.2 <- ggemmeans(lm.COG.2, terms = c("COG20_CATEGORY", "type"))
pred.COG.9 <- ggemmeans(lm.COG.9, terms = c("COG20_CATEGORY"))
pred.COG.10.1 <- ggemmeans(lm.COG.10, terms = c("COG20_CATEGORY"))
pred.COG.10.2 <- ggemmeans(lm.COG.10, terms = c("COG20_CATEGORY", "type"))
pred.COG.11 <- ggemmeans(lm.COG.11, terms = c("COG20_CATEGORY", "type"))
pred.COG.12 <- ggemmeans(lm.COG.12, terms = c("COG20_CATEGORY", "type"))
pred.COG.13 <- ggemmeans(lm.COG.13, terms = c("COG20_CATEGORY", "type"))
pred.COG.14 <- ggemmeans(lm.COG.14, terms = c("COG20_CATEGORY", "type"))
plot(pred.COG.2)
plot(pred.COG.9)
plot(pred.COG.10.1)
plot(pred.COG.10.2)
plot(pred.COG.11)
plot(pred.COG.12)
plot(pred.COG.13)
plot(pred.COG.14)

pred.COG <- ggemmeans(lm.COG.10, terms = c("COG20_CATEGORY", "host_species"))

pred.COG <- ggemmeans(lm.COG.10, terms = c("COG20_CATEGORY","host_species", "type", "methylo_sp"))

pred.COG <- ggemmeans(lm.COG.10, terms = c("COG20_CATEGORY", "type","methylo_sp"))


pred.COG <- ggemmeans(lm.COG.10, terms = c("COG20_CATEGORY"))

```
 
## Linear Model with COG category (!! important to run all the block!!)
```{r}
lm.COG <- lm(log(pNpS_gene_consensus) ~ type * host_species * methylo_sp + COG20_CATEGORY + type:COG20_CATEGORY, data = pNpS_all_sp.multi) 

summary(lm.COG) 
anova(lm.COG) # type I ... type III does not work

# Get the levels of factor 'COG20_CATEGORY' based on decreasing EMM averaged across type, methylo_sp and host_species
emm.cat.COG.means <-emmeans(lm.COG, specs = pairwise ~ COG20_CATEGORY, adjust = "bonferroni", type = "response")
emm.cat.COG.means$emmeans
emm.cat.COG.means <- as.data.frame(emm.cat.COG.means$emmeans)
emm.cat.COG.means[6,2] <- 0

emm.cat.COG.means.decr <- emm.cat.COG.means %>%
  mutate(COG20_CATEGORY = fct_reorder(COG20_CATEGORY, response, .fun = mean, .desc = TRUE))
levels(emm.cat.COG.means.decr$COG20_CATEGORY)

# !!! IMPORTANT TO RERUN THE MODEL

## re-run model with good factor order
pNpS_all_sp.multi$COG20_CATEGORY <- factor(pNpS_all_sp.multi$COG20_CATEGORY, levels = levels(emm.cat.COG.means.decr$COG20_CATEGORY))

lm.COG <- lm(log(pNpS_gene_consensus) ~ type * host_species * methylo_sp + COG20_CATEGORY + type:COG20_CATEGORY, data = pNpS_all_sp.multi) 
summary(lm.COG)
anova(lm.COG)

emm.cat.COG <-emmeans(lm.COG, specs = pairwise ~ COG20_CATEGORY|type, adjust = "bonferroni", type = "response")
emm.cat.COG$emmeans

# Table to report in supp mat
emm.COG.df <- merge(emm.cat.COG$emmeans[1:23],emm.cat.COG$emmeans[24:46], by = "COG20_CATEGORY")
emm.COG.df <- emm.COG.df[c(1:3,6:9,12:13)]
emm.COG.df$response.x <- round(emm.COG.df$response.x, 3)
emm.COG.df$lower.CL.x <- round(emm.COG.df$lower.CL.x, 3)
emm.COG.df$upper.CL.x <- round(emm.COG.df$upper.CL.x, 3)
emm.COG.df$response.y <- round(emm.COG.df$response.y, 3)
emm.COG.df$lower.CL.y <- round(emm.COG.df$lower.CL.y, 3)
emm.COG.df$upper.CL.y <- round(emm.COG.df$upper.CL.y, 3)
emm.COG.df <- emm.COG.df[order(emm.COG.df$COG20_CATEGORY), ]
rownames(emm.COG.df) <- NULL
```

### Plot
```{r}
pred.COG <- ggemmeans(lm.COG, terms = c("COG20_CATEGORY", "type"))

COG.plot <- plot(pred.COG, dodge = 0.25)

COG.plot +
  xlab(expression(bold("COG Category"))) +
  ylab(expression(bold("Model-predicted") ~ bolditalic("pN/pS")^bold("(gene)"))) +
  scale_color_manual(values=c("gray25","red3"), name = expression(bold("Gene type")), labels = c("Core genes" = "Core genes", "Accessory genes" = "Accessory genes")) +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
      legend.justification = c(0.5, 0),  
      legend.text = element_text(size = 9),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      panel.grid.major.y = element_line(),
      panel.grid.minor.y = element_line(),
      plot.title = element_blank(),
      axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = 1, vjust = 0.4),
      axis.text.y = element_text(color = "black", size = 11, angle = 0, hjust = 0, vjust = 0.5),  
      axis.title.x = element_text(color = "black", size = 13, angle = 0, hjust = 0.5, vjust = -1),
      axis.title.y = element_text(color = "black", size = 13, angle = 90, hjust = 0.5, vjust = 2)
      )
ggsave("../figures/predicted_pNpS_COG.tiff", width = 10, height = 8)
```


## Linear Model for testing plant clade
```{r}
lm.clade <- lm(log(pNpS_gene_consensus) ~ type * methylo_sp * plant_clade, data = pNpS_all_sp[pNpS_all_sp$host_species != "COCO",])
summary(lm.clade)
anova(lm.clade)
Anova(lm.clade, type = 2)
Anova(lm.clade, type = 3)

emm.clade <- emmeans(lm.clade, ~ type * methylo_sp * plant_clade)
emm.clade

pred.type.cl <- ggpredict(lm.clade, terms = c("type"))
plot(pred.type.cl)

pred.meth_sp.cl <- ggpredict(lm.clade, terms = c("methylo_sp"))
plot(pred.meth_sp.cl)

pred.clade.cl <- ggpredict(lm.clade, terms = c("plant_clade"))
plot(pred.clade.cl)

pred.all.clade <- ggpredict(lm.clade, terms = c("methylo_sp", "plant_clade", "type"), back.transform = TRUE)
plot(pred.all.clade)
```


## Linear Model for testing plant_type (form)
```{r}
lm.form <- lm(log(pNpS_gene_consensus) ~ type * methylo_sp * plant_type, data = pNpS_all_sp)
summary(lm.form)
anova(lm.form)
Anova(lm.form, type = 2)
Anova(lm.form, type = 3)

emm.form <- emmeans(lm.form, ~ type * methylo_sp * plant_type)
emm.form

pred.type.frm <- ggpredict(lm.form, terms = c("type"))
plot(pred.type.frm)

pred.meth_sp.frm <- ggpredict(lm.form, terms = c("methylo_sp"))
plot(pred.meth_sp.frm)

pred.form.frm <- ggpredict(lm.form, terms = c("plant_type"))
plot(pred.form.frm)

pred.all.form <- ggpredict(lm.form, terms = c("methylo_sp", "plant_type", "type"), back.transform = TRUE)
plot(pred.all.form)
```

################################################################################
